---
title: "TreeNetProc TWD graphing"
author: "Gabby John"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(treenetproc, `force = TRUE`)
library(zoo)
library(chron)
library(viridis)
library(dplyr)
library(ggplot2)
library(lubridate)
```

## step 0: identify helper functions

```{r}
left <- function(string, char){substr(string, 1,char)} 
right <- function (string, char){substr(string,nchar(string)-(char-1),nchar(string))}
```

## step 1: read in dendrometer data and plot raw data
R will not open your CSV file if its title starts with a number and will instead give an error message. You can change the title so it starts with a letter character instead. 
Make sure that the NA strings reader matches the NA codes in the original CSV file, in this case "NaN". 

```{r}
dend310tsfull <- read.csv("~/research/dend310/310 summer work/310tsfull.csv", na.strings="NaN", stringsAsFactors=TRUE)
head(dend310tsfull)
```


## Step 2: fix timestamps (ts) and connect with dendrometer ID number

```{r}
dend310<-dend310tsfull
datetime<-dend310$datetime
ID = '310'
ts = as.POSIXct(dend310['datetime'][,],format="%Y-%m-%d %H:%M:%S",tz="GMT") 
dend310_L0<-cbind(dend310["value"]+0,ts,series=ID)
```

## Step 3: set parameters to plot raw data

```{r}
#First grab years string
years <-left(dend310_L0[,"ts"],4)
#To plot, set number of plots with par()
par(mfrow=c(1,1))
#Then set margins 
par(mar = c(5, 5, 5, 5))
```


## Step 5: cleaning the dendrometery data

```{r}
#Level 1 processing (time align) method 1
?treenetproc::proc_L1 
dend310_L1<- proc_L1(data_L0 = dend310_L0, reso = 60 ) 
                  #input = "wide", date_format ="%Y-%m-%d %H:%M:%S") 
head(dend310_L1)
#Level 1 processing (time align) method 2

dend310_L1<- proc_L1(data_L0 = dend310_L0, reso = 60,
                  date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT")
head(dend310_L1)

#level 1 processing of temperature data
DSCMET421_mean56m <- read.csv("~/research/DSCMET421 climate data/DSCMET421_mean56m.csv", na.strings="NaN", stringsAsFactors=TRUE)
head(DSCMET421_mean56m)

temp_data_310_L0 <- cbind(DSCMET421_mean56m,series=ID)
colnames(temp_data_310_L0) <-c('ts','value','series')

#Now we can time-align temperature data with proc_L1
temp_data_310_L1 <- proc_L1(data_L0 = temp_data_310_L0, reso = 60, 
                            date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT") 
head(temp_data_310_L1)
```

Now we can move on to proc_dendro_L2, which integrates dendrometer readings with temperature, detects errors, corrects them and plots gro and twd by date.more info here: ?treenetproc::proc_dendro_L2

first plot full dataset for dendrometer. Plot will be exported to working directory as pdf

```{r}
#first plot full dataset for dendrometer. NOTE: Plot will be exported to working directory as pdf
dend310_L2 <- proc_dendro_L2(dendro_L1 = dend310_L1, temp_L1 = temp_data_310_L1, tol_out = 2, tol_jump = 3, plot = TRUE, frost_thr = 1, plot_period = "full", plot_export = TRUE, interpol = 3.5*60, frag_len = NULL, plot_name = "dend310summer_full", tz="GMT") 
graphics.off()
```

##Step 5.1: Tree water deficit (TWD) as an indicator of drought stress.

```{r}
# plot minimum daily twd against day of year

par(mfrow=c(1,1))
par(mar = c(5, 5, 5, 5))

plot(1,
     1,
     ylim=c(0,max(dend310_L2$twd,na.rm=T)),
     xlim=c(100,300),
     ylab=expression("twd ("*mu*"m)"),
     xlab="Day of year",
     col="white")

for(y in c(1:length(unique(left(dend310_L2$ts,4))))){
 
   # selected year
  sel<-dend310_L2[which(left(dend310_L2$ts,4)==unique(left(dend310_L2$ts,4))[y]),]
 
   # calc twd
  twd<-suppressWarnings(aggregate(sel$twd,list(as.Date(sel$ts)),min,na.rm=T))
  twd$doy<-as.numeric(strftime(as.Date(twd$Group.1), format = "%j"))

  # clean
  twd[which(twd$x=="Inf"),"x"]<-NA

  lines(twd$doy,twd$x,col=col_sel[y],lwd=1.5)
  twd[which(is.na(twd$x)==T),"x"]<-0

}

legend("topright",
       c(unique(left(dend310_L2$ts,4))),
       col=col_sel,
       bty="n",pch=16)

# Add growing season extent:
grow_seas_L2 <- grow_seas(dendro_L2 = dend310_L2,
                          agg_yearly=TRUE,
                          tol_seas = 0.1)


abline(v=c(mean(grow_seas_L2$gro_start),
           mean(grow_seas_L2$gro_end)),
       lty=2)
text(mean(c(mean(grow_seas_L2$gro_start),
            mean(grow_seas_L2$gro_end))),
     max(dend310_L2$twd,na.rm=T), "Growing season")
```