---
title: "Behind the scenes code that only needs to be read once"
author: "Gabby John"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load(file=here("Climate/Primet2014_2024.RData"))
load(file=here("Climate/Dscmet2016_2024.RData"))
dendrometers <- dscmet_422_all |> select(Date,DEND309_AVG_4200_0_01,Flag_DEND309_AVG_4200_0_01,)
```

#From extracting and processing VPD and AT from DSCMET
##First: load
load(file=here("Climate/Dscmet2016_2024.RData"))

#Second: check for bad flags (only need to do once) - TAKES FOREVER
unique_dscmet_strings <- unique( 
unlist(   
sapply(     
Dscmet2016_2024, function(x) x[is.character(x)])))
unique_primet_strings <- unique(  unlist(    sapply(      Primet2016_2024, function(x) x[is.character(x)])))
values_to_remove_P230rain <- c("Q", "QQ")
# Filter data frame to swap back values with NA#extracting VPD and getting bad flag strings 
Dscmet_VPD52m <- Dscmet2016_2024|>select(c(Date,VPD_MEAN_5600_0_01,Flag_VPD_MEAN_5600_0_01))
Dscmet_VPD52m$Date<-as.POSIXct(Dscmet_VPD52m$Date,format="%Y-%m-%d %H:%M:%S")unique_dscmet_VPD52m_strings <- unique( unlist(    sapply(      Dscmet_VPD52m, function(x) x[is.character(x)])))
Dscmet_VPD52m$Flag_VPD_MEAN_5600_0_01[Dscmet_VPD52m$Flag_VPD_MEAN_5600_0_01=="Q"] <- NA

##mb to kpa 
Dscmet_VPD52m$VPD_kpa<-Dscmet_VPD52m$VPD_MEAN_5600_0_01/10
save(Dscmet_VPD52m,file = here("Climate/filtered_Dscmet_VPD52m.RData"))

##same with air temp
Dscmet_AT52m <- Dscmet2016_2024|>select(c(Date,AIRTEMP_MEAN_5600_0_01,Flag_AIRTEMP_MEAN_5600_0_01,AIRTEMP_MAX_5600_0_01,Flag_AIRTEMP_MAX_5600_0_01))Dscmet_AT52m$Date<-as.POSIXct(Dscmet_AT52m$Date,format="%Y-%m-%d %H:%M:%S")unique_dscmet_AT52m_strings <- unique(  unlist(    sapply(      Dscmet_AT52m, function(x) x[is.character(x)])))
Dscmet_AT52m[Dscmet_AT52m %in% c("MI", "Q")] <- NA
save(Dscmet_AT52m,file = here("Climate/filtered_Dscmet_AT52m.RData"))

##adding hour and day columns
##VPD
Dscmet_VPD52m$hour <- format(
  Dscmet_VPD52m$Date, "%Y-%m-%d %H:00:00")
Dscmet_VPD52m$Date<-as.POSIXct(
  Dscmet_VPD52m$hour,format="%Y-%m-%d %H")

Dscmet_VPD52m$day <- format(
  Dscmet_VPD52m$Date, "%Y-%m-%d")
Dscmet_VPD52m$day<-as.POSIXct(
  Dscmet_VPD52m$day,format="%Y-%m-%d")

##AT
Dscmet_AT52m$hour <- format(
  Dscmet_AT52m$Date, "%Y-%m-%d %H:00:00")
Dscmet_AT52m$Date<-as.POSIXct(
  Dscmet_AT52m$hour,format="%Y-%m-%d %H")

Dscmet_AT52m$day <- format(
  Dscmet_AT52m$Date, "%Y-%m-%d")
Dscmet_AT52m$day<-as.POSIXct(
  Dscmet_AT52m$day,format="%Y-%m-%d")

##hourly VPD means
filtered_Dscmet_VPD52m_hourly <- Dscmet_VPD52m |> 
  group_by(hour) |> 
  summarise(VPD52m_kpa_hourly = mean(VPD_kpa,na.rm=TRUE))|>
  filter(!is.nan(VPD52m_kpa_hourly))
##daily VPD means 
filtered_Dscmet_VPD52m_daily <- Dscmet_VPD52m |> 
  group_by(day) |> 
  summarise(VPD52m_kpa_daily = mean(VPD_kpa,na.rm=TRUE))|>
  filter(!is.nan(VPD52m_kpa_daily))

##hourly max VPD mean
filtered_Dscmet_maxVPD52m_hourly <- Dscmet_VPD52m |> 
  group_by(hour) |> 
  summarise(maxVPD52m_kpa_hourly = max(VPD_kpa,na.rm=TRUE))|>
  filter(!is.nan(maxVPD52m_kpa_hourly))
##daily VPD max 
filtered_Dscmet_maxVPD52m_daily <- Dscmet_VPD52m |> 
  group_by(day) |> 
  summarise(maxVPD52m_kpa_daily = max(VPD_kpa,na.rm=TRUE))|>
  filter(!is.nan(maxVPD52m_kpa_daily))

##hourly and daily AT means 
filtered_Dscmet_AT52m_hourly <- Dscmet_AT52m |> 
  group_by(hour) |> 
  summarise(AT52m_mean_hourly = mean(AIRTEMP_MEAN_5600_0_01,na.rm=TRUE))|>
  filter(!is.nan(AT52m_mean_hourly))
##daily values 
filtered_Dscmet_AT52m_daily <- Dscmet_AT52m |> 
  group_by(day) |> 
  summarise(AT52m_mean_daily = mean(AIRTEMP_MEAN_5600_0_01,na.rm=TRUE))|>
  filter(!is.nan(AT52m_mean_daily))

##hourly and daily AT max 
filtered_Dscmet_maxAT52m_hourly <- Dscmet_AT52m |> 
  group_by(hour) |> 
  summarise(AT52m_max_hourly = max(AIRTEMP_MAX_5600_0_01,na.rm=TRUE))|>
  filter(!is.nan(AT52m_max_hourly))
##daily values 
filtered_Dscmet_maxAT52m_daily <- Dscmet_AT52m |> 
  group_by(day) |> 
  summarise(AT52m_max_daily = max(AIRTEMP_MAX_5600_0_01,na.rm=TRUE))|>
  filter(!is.nan(AT52m_max_daily))

##saving all of these
##means
save(filtered_Dscmet_VPD52m_hourly,file = here(
  "Climate/filtered_Dscmet_VPD52m_hourly.RData"))
save(filtered_Dscmet_VPD52m_daily,file = here(
  "Climate/filtered_Dscmet_VPD52m_daily.RData"))
save(filtered_Dscmet_AT52m_hourly,file = here(
  "Climate/filtered_Dscmet_AT52m_hourly.RData"))
save(filtered_Dscmet_AT52m_daily,file = here(
  "Climate/filtered_Dscmet_AT52m_daily.RData"))
##maxes
save(filtered_Dscmet_maxVPD52m_hourly,file = here(
  "Climate/filtered_Dscmet_maxVPD52m_hourly.RData"))
save(filtered_Dscmet_maxVPD52m_daily,file = here(
  "Climate/filtered_Dscmet_maxVPD52m_daily.RData"))
save(filtered_Dscmet_maxAT52m_hourly,file = here(
  "Climate/filtered_Dscmet_maxAT52m_hourly.RData"))
save(filtered_Dscmet_maxAT52m_daily,file = here(
  "Climate/filtered_Dscmet_maxAT52m_daily.RData"))




#From FES524 Final Project
data_prep524<-bind_rows(dscmet_422_all,P230_df)
data_prep_422_230<-data_prep524|>select(c(
  Site,
  Date,
  DEND309_AVG_4200_0_01,Flag_DEND309_AVG_4200_0_01,
  DEND310_AVG_150_0_01,Flag_DEND310_AVG_150_0_01,
  DEND311_AVG_150_0_01,Flag_DEND311_AVG_150_0_01,
  DEND646_AVG_150_0_01,Flag_DEND646_AVG_150_0_01,
  DEND648_AVG_150_0_01,Flag_DEND648_AVG_150_0_01,
  DEND652_AVG_150_0_01,Flag_DEND652_AVG_150_0_01,
  Date,
  yr,
  DOY,
  SOILTEMP_MEAN_0_10_01,Flag_SOILTEMP_MEAN_0_10_01,
  SOILTEMP_MEAN_0_20_02,Flag_SOILTEMP_MEAN_0_20_02,
  SOILTEMP_MEAN_0_50_03,Flag_SOILTEMP_MEAN_0_50_03,
  SOILTEMP_MEAN_0_100_04,Flag_SOILTEMP_MEAN_0_100_04, 
  SWC_MEAN_0_10_01,Flag_SWC_MEAN_0_10_01,
  SWC_MEAN_0_20_02,Flag_SWC_MEAN_0_20_02,
  SWC_MEAN_0_50_03,Flag_SWC_MEAN_0_50_03,
  SWC_MEAN_0_100_04,Flag_SWC_MEAN_0_100_04,
  PRECIP_TOT_100_0_01,Flag_PRECIP_TOT_100_0_01,
  PRECIP_ACC_100_0_01,Flag_PRECIP_ACC_100_0_01
  ))
save(data_prep_422_230,file = "data_prep_422_230.RData") #have to run this in the console for it to save to the directory 
load("Data/data_prep_422_230.RData")
#more data prep: raw to filtered and processed
#getting hourly averages - starting with adding an hour column
data_prep_422_230 <- data_prep_422_230 |>
  mutate(hour = floor_date(Date, unit = "hour"))
data_prep_422_230$year <- format(data_prep_422_230$hour, "%Y")
data_prep_422_230$year<-as.factor(data_prep_422_230$year)
data_prep_422_230<-data_prep_422_230|>
  filter(year!="2018")|>
  mutate(
    DEND860_AVG_150_0_01_NoNeg = ifelse(DEND860_AVG_150_0_01 < 0, NA,
                                        DEND860_AVG_150_0_01))
data_prep_6dendros_hourly<-data_prep_422_230|>select(c(
  Date,
  hour,
  year,
  DEND310_AVG_150_0_01,
  DEND311_AVG_150_0_01,
  DEND654_AVG_150_0_01,
  DEND655_AVG_150_0_01,
  DEND656_AVG_150_0_01,
  DEND860_AVG_150_0_01_NoNeg,
))
save(data_prep_6dendros_hourly,file = "data_prep_6dendros_hourly.RData")


```{r}
load("Data/data_prep_6dendros_hourly.RData")
data_prep_6dendros_hourly$day <- format(
  data_prep_6dendros_hourly$Date, "%Y-%m-%d")
data_prep_6dendros_hourly$day<-as.POSIXct(data_prep_6dendros_hourly$day,format="%Y-%m-%d")


#daily values 
#310 - OG DF
d310_daily <- data_prep_6dendros_hourly |> 
  group_by(day) |> 
  summarise(value = mean(DEND310_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(value))  

# 311 - OG DF
d311_daily <- data_prep_6dendros_hourly |> 
  group_by(day) |> 
  summarise(value = mean(DEND311_AVG_150_0_01, na.rm = TRUE)) |> 
  filter(!is.nan(value))

# 654 - Secondary DF
d654_daily <- data_prep_6dendros_hourly |> 
  group_by(day) |> 
  summarise(value = mean(DEND654_AVG_150_0_01, na.rm = TRUE)) |> 
  filter(!is.nan(value))

# 655 - Secondary DF
d655_daily <- data_prep_6dendros_hourly |> 
  group_by(day) |> 
  summarise(value = mean(DEND655_AVG_150_0_01, na.rm = TRUE)) |> 
  filter(!is.nan(value))

# 656 - Secondary DF
d656_daily <- data_prep_6dendros_hourly |> 
  group_by(day) |> 
  summarise(value = mean(DEND656_AVG_150_0_01, na.rm = TRUE)) |> 
  filter(!is.nan(value))

# 860 - Secondary DF - note only ~15% of original canopy remains alive after 2024
d860_daily <- data_prep_6dendros_hourly |> 
  group_by(day) |> 
  summarise(value = mean(DEND860_AVG_150_0_01_NoNeg, na.rm = TRUE)) |> 
  filter(!is.nan(value))


#Correcting jumps
jump_free_d310<-jump.locator(df=d310_daily,v=200)
jump_free_d310$ID<-as.factor("310")
jump_free_d310$AgeClass<-as.factor("OG")
#differencing
jump_free_d310 <- jump_free_d310 |> arrange(TIME)|> mutate(
    value_diff = value - lag(value))

jump_free_d311<-jump.locator(df=d311_daily,v=200)
jump_free_d311$ID<-as.factor("311")
jump_free_d311$AgeClass<-as.factor("OG")
#differencing
jump_free_d311 <- jump_free_d311 |> arrange(TIME)|> mutate(
    value_diff = value - lag(value))

jump_free_d654<-jump.locator(df=d654_daily,v=200)
jump_free_d654$ID<-as.factor("654")
jump_free_d654$AgeClass<-as.factor("Secondary")
#differencing
jump_free_d654 <- jump_free_d654 |> arrange(TIME)|> mutate(
    value_diff = value - lag(value))

jump_free_d655<-jump.locator(df=d655_daily,v=200)
jump_free_d655$ID<-as.factor("655")
jump_free_d655$AgeClass<-as.factor("Secondary")
#differencing
jump_free_d655 <- jump_free_d655 |> arrange(TIME)|> mutate(
    value_diff = value - lag(value))

jump_free_d656<-jump.locator(df=d656_daily,v=200)
jump_free_d656$ID<-as.factor("656")
jump_free_d656$AgeClass<-as.factor("Secondary")
#differencing
jump_free_d656 <- jump_free_d656 |> arrange(TIME)|> mutate(
    value_diff = value - lag(value))

jump_free_d860<-jump.locator(df=d860_daily,v=200)
jump_free_d860$ID<-as.factor("860")
jump_free_d860$AgeClass<-as.factor("Secondary")
#differencing
jump_free_d860 <- jump_free_d860 |> arrange(TIME)|> mutate(
    value_diff = value - lag(value))


data_prep_6dendros_daily_jf<-bind_rows(
  jump_free_d310,
  jump_free_d311,
  jump_free_d654,
  jump_free_d655,
  jump_free_d656,
  jump_free_d860)
```

	

AT52m_max_daily
2021-06-28
44.63
2021-06-27
42.67
2017-08-02
40.68
2020-08-15
40.11
2022-07-30
39.96
2022-07-29
39.73
2017-06-25
39.58
2017-08-03
39.13
2021-08-11
39.00
2024-07-08
38.96
2024-07-05
38.72
2024-09-05
38.59
2021-06-26
38.55

#Try #2 for making a model
DOY 178 = june 27
DOY 179 = june 28
DOY 178-28 = 150 = May 30
DOY 179+28 = 207 = July 26
```{r}
load(file=here("DendroAnalyst/daily_stats_6dendros.RData"))
#making m-d column
daily_stats_6dendros$day <- format(daily_stats_6dendros$DATE,format = "%m-%d")

#truncating df to be HD +/- 4 weeks
daily_stats_6dendros_truncated <- daily_stats_6dendros |>
  filter(
    (month(DATE) == 5 & day(DATE) >= 30) |
    (month(DATE) == 6) |
    (month(DATE) == 7 & day(DATE) <= 26))

#adding period column
daily_stats_6dendros_truncated <- daily_stats_6dendros_truncated |>
  mutate(
    period = case_when(
      (month(DATE) == 5 & day(DATE) >= 30) |
      (month(DATE) == 6 & day(DATE) < 27) ~ "before",
      (month(DATE) == 6 & day(DATE) %in% c(27, 28)) ~ "during",
      (month(DATE) == 6 & day(DATE) > 28) |
      (month(DATE) > 6) ~ "after",
      TRUE ~ NA_character_))

# Ensuring Period is chronologically ordered 
daily_stats_6dendros_truncated$period <- factor(daily_stats_6dendros_truncated$period, levels = c("before", "during", "after"))

#Days are repeating so let's do Julian days instead of DOY days
daily_stats_6dendros_truncated <- daily_stats_6dendros_truncated |>
  arrange(DATE) |>
  mutate(
    day_integer = as.numeric(as.Date(DATE) - min(as.Date(DATE)) + 1))


#updated model: july
library(emmeans)
library(nlme)

model1 <- lme(
  amplitude ~ yr*period + AgeClass,
  data = daily_stats_6dendros_truncated,
  random = ~ 1 | yr/ID, #changed from 1 | ID or else wouldn't run
  correlation = corARMA(p=1,form = ~ day_integer | yr/ID)) #/ not : to avoid incompatible formula
#is p=1 correct?

model2 <- lme(
  amplitude_signed ~ yr*period + AgeClass,
  data = daily_stats_6dendros_truncated,
  random = ~ 1 | yr/ID, #changed from 1 | ID or else wouldn't run
  correlation = corARMA(p=1,form = ~ day_integer | yr/ID)) #/ not : to avoid incompatible formula
#is p=1 correct?

model3 <- lme(
  amplitude_signed ~ yr*period + AgeClass,
  data = daily_stats_6dendros_truncated,
  random = list(~ 1 | ID), #changed from 1 | ID or else wouldn't run
  correlation = corARMA(p=1,form = ~ day_integer | ID)) #/ not : to avoid incompatible formula



summary(model1)
summary(model2)
summary(model3)

# Diagnostic plots - amp
qqnorm(model1, ~ resid(., type = "p")) #test for normality
acf(residuals(model1)) # checks correlation structure
plot(model1, which = 1)  #test for equal variance
hist(residuals(model1, type = "pearson"))

# Diagnostic plots - amp_signed
qqnorm(model2, ~ resid(., type = "p")) #test for normality
acf(residuals(model2)) # checks correlation structure
plot(model2, which = 1)  #test for equal variance
hist(residuals(model1, type = "pearson"))

# Diagnostic plots - amp_signed
qqnorm(model3, ~ resid(., type = "normalized")) #test for normality
acf(residuals(model3)) # checks correlation structure
plot(model3, which = 1)  #test for equal variance
hist(residuals(model3, type = "normalized"))

```


data_prep_6dendros_daily_jf$Year <- as.factor(data_prep_6dendros_daily_jf$year)

data_6dendros_growing_season <- data_prep_6dendros_daily_jf |>
  filter(month(TIME) >= 4 & month(TIME) <=9)
data_6dendros_growing_season

data_6dendros_growing_season <- data_6dendros_growing_season |>
  filter(!is.na(value_diff),
         !is.na(Day),
         !is.na(Year),
         !is.na(Period),
         !is.na(AgeClass),
         !is.na(ID))
data_6dendros_growing_season<-data_6dendros_growing_season|>select(c(TIME,ID,AgeClass,Period,Year,Day,value_diff))

save(data_6dendros_growing_season,file = "Data/data_6dendros_growing_season.RData")
load("Data/data_6dendros_growing_season.RData")
```