---
title: "Ecomatik_Adjustments"
format: html
editor: visual
---

## Notes

QA/QC flags for Dendrometer data value (Vi). Resolution on CR1000 3.3 um. Accuracy on CR1000: +- 0.12%. (flagging criteria, where "x" is DEND310_AVG_150_0_01: flag_valuechange(PROGID,0.1,0.1,1)="P")

Sometimes there are multiple points of addition for each adjustment line. This is the product of Gabby adjusting her adjustment values, usually due to finding another flag after a first run and simply finding it easier to add the new adjustment value rather than recalculating the original value.

IDEA: similar to the d'orangeville paper now that we have circumference data let's see if we can find the dates at which 25%, 50%, and 75% growth are done based on total circ gained in 1 year or maybe in one growing szn?

V0 : the first valid record of the dendrometer in mm

BAI: Basal Area = π × Radius²Basal Area = π × Radius²

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
library(here)
library(dplyr)
library(ggplot2)
# naming year breaks and labels for plotting later
year_breaks19_24 <- seq(as.POSIXct("2019-01-01"), as.POSIXct("2024-01-01"), by = "year")
year_labels19_24 <- paste("Jan", format(year_breaks19_24, "%Y"))
```

little difference between mean and median reductions in data

309

```{r}
D309_annotated <- read_csv(here("Data/D309_annotated.csv"))
D309_annotated_adjusted <- D309_annotated %>%
  mutate(adjusted_value = case_when(
    Flag_DEND309_AVG_4200_0_01 == "A" ~ DEND309_AVG_4200_0_01 - 21009,
    Flag_DEND309_AVG_4200_0_01 == "B" ~ DEND309_AVG_4200_0_01 + 4246,
    Flag_DEND309_AVG_4200_0_01 == "C" ~ DEND309_AVG_4200_0_01 + 4246, 
    Flag_DEND309_AVG_4200_0_01 == "D" ~ DEND309_AVG_4200_0_01 + 4211, 
    Flag_DEND309_AVG_4200_0_01 == "F" ~ DEND309_AVG_4200_0_01 - 6255, 
    Flag_DEND309_AVG_4200_0_01 == "FF" ~ DEND309_AVG_4200_0_01 - 7205,  
    Flag_DEND309_AVG_4200_0_01 == "FFF" ~ DEND309_AVG_4200_0_01 - 7718,      
    Flag_DEND309_AVG_4200_0_01 == "G" ~ DEND309_AVG_4200_0_01 + 10226 - 513,  
    Flag_DEND309_AVG_4200_0_01 == "H" ~ DEND309_AVG_4200_0_01 + 8108 - 513, 
    Flag_DEND309_AVG_4200_0_01 == "J" ~ DEND309_AVG_4200_0_01 - 7554 - 513, 
    Flag_DEND309_AVG_4200_0_01 == "K" ~ DEND309_AVG_4200_0_01 - 426 - 513,
    Flag_DEND309_AVG_4200_0_01 == "L" ~ DEND309_AVG_4200_0_01 - 454 - 513,
    Flag_DEND309_AVG_4200_0_01 == "O" ~ DEND309_AVG_4200_0_01 - 443 - 513,
    Flag_DEND309_AVG_4200_0_01 == "P" ~ DEND309_AVG_4200_0_01 - 443 - 513,
    Flag_DEND309_AVG_4200_0_01 == "PP" ~ DEND309_AVG_4200_0_01 - 432 - 513,
    Flag_DEND309_AVG_4200_0_01 == "PPP" ~ DEND309_AVG_4200_0_01 - 995 - 513,
    TRUE ~ DEND309_AVG_4200_0_01
  ))
D309_annotated_adjusted <- D309_annotated_adjusted |>
  mutate(difference = DEND309_AVG_4200_0_01 - lag(DEND309_AVG_4200_0_01))
ggplot(D309_annotated_adjusted,aes(Date,adjusted_value))+geom_point(alpha=0.7)

#separating dfs based on PLNG value 
# 2.353 until 10.49 which starts 2020-07-20 09:10:00
D309_annotated_adjusted_PLNG1 <- D309_annotated_adjusted |>  filter(DEND309_PLNG_4200_0_01 == 2.353)|>arrange(Date)
D309_annotated_adjusted_PLNG2 <- D309_annotated_adjusted |>  filter(DEND309_PLNG_4200_0_01 == 10.49)

#circumference conversion
D309_annotated_adjusted <- D309_annotated_adjusted |> arrange(Date)
D309_adjusted_circ <- dc3_to_circumference(
  df = D309_annotated_adjusted_PLNG1,
  date_column = D309_annotated_adjusted_PLNG1$Date, 
  Vi_micrometers =D309_annotated_adjusted_PLNG1$adjusted_value, 
  C0= 1586, V0 = 2.353) 

D309_adjusted_circ2 <- dc3_to_circumference(
  df = D309_annotated_adjusted_PLNG2,
  date_column = D309_annotated_adjusted_PLNG2$Date, 
  Vi_micrometers =D309_annotated_adjusted_PLNG2$adjusted_value, 
  C0= 1589.756, V0 = 10.49) 

D309_adjusted_circ3 <- dc3_to_circumference(
  df = D309_annotated_adjusted,
  date_column = D309_annotated_adjusted$Date, 
  Vi_micrometers =D309_annotated_adjusted$adjusted_value, 
  C0= 1586, V0 = 2.353) 

ggplot(D309_adjusted_circ,aes(date,circumference_mm))+geom_point(alpha=0.7)
ggplot(D309_adjusted_circ2,aes(date,circumference_mm))+geom_point(alpha=0.7)
ggplot(D309_adjusted_circ3,aes(date,circumference_mm))+geom_point(alpha=0.7)
```

# 309

```{r}
ff
```

# 310

```{r}
#loading raw adjusted data from Excel
D310_manual_flags <- read.csv(here("Data/D310_manual_flags.csv"))
D310_manual_flags$Date<-as.POSIXct(D310_manual_flags$Date,format="%Y-%m-%d %H:%M:%S")
D310_manual_flags$DEND310_AVG_150_0_01[D310_manual_flags$DEND310_AVG_150_0_01 == ""] <- NA


#plotting raw data
D310plot_raw <- ggplot(D310_manual_flags,aes(Date,DEND310_AVG_150_0_01))+geom_point(alpha=0.7)+
  labs(title = "OG DF (ID 310) Raw Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()

#converting each raw data segment to circumference based on manual flag
D310_raw_A <- D310_manual_flags |> filter(Manual_flag=="A")
D310_circ_A <- dc3_to_circumference(
  df = D310_raw_A,
  date_column = D310_raw_A$Date, 
  Vi_micrometers =D310_raw_A$DEND310_AVG_150_0_01, 
  C0= 3402, V0 = 3.462, na_method = "skip")

D310_raw_B <- D310_manual_flags |> filter(Manual_flag=="B")
D310_circ_B <- dc3_to_circumference(
  df = D310_raw_B,
  date_column = D310_raw_B$Date, 
  Vi_micrometers =D310_raw_B$DEND310_AVG_150_0_01, 
  C0= 3420.760, V0 = 4.976, na_method = "skip") #C0 from last circ value from previous conversion; V0 can automatically fill but can be input from first valid point in mm in new dataset

D310_raw_C <- D310_manual_flags |> filter(Manual_flag=="C")
D310_circ_C <- dc3_to_circumference(
  df = D310_raw_C,
  date_column = D310_raw_C$Date, 
  Vi_micrometers =D310_raw_C$DEND310_AVG_150_0_01, 
  C0= 3427.691, V0 = 5.967, na_method = "skip")

D310_raw_D <- D310_manual_flags |> filter(Manual_flag=="D")
D310_circ_D <- dc3_to_circumference(
  df = D310_raw_D,
  date_column = D310_raw_D$Date, 
  Vi_micrometers =D310_raw_D$DEND310_AVG_150_0_01, 
  C0= 3434.585, V0 = 6.219, na_method = "skip")

D310_raw_E <- D310_manual_flags |> filter(Manual_flag=="E")
D310_circ_E <- dc3_to_circumference(
  df = D310_raw_E,
  date_column = D310_raw_E$Date, 
  Vi_micrometers =D310_raw_E$DEND310_AVG_150_0_01, 
  C0= 3440.615, V0 = 2.309, na_method = "skip")

#combining all of these into one df
D310_circ_A_E <- bind_rows(D310_circ_A,D310_circ_B,D310_circ_C,D310_circ_D,D310_circ_E)
D310_circ_A_E <- D310_circ_A_E |>
  mutate(difference = circumference_mm - lag(circumference_mm))

#plotting & saving circumference 
D310plot_circ <- ggplot(D310_circ_A_E,aes(date,circumference_mm))+geom_point(alpha=0.7)+
  labs(title = "OG DF (ID 310) Stem Circumference",
    x = "Date",
    y = "Stem Circumference (mm)") +
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()+
  geom_vline(xintercept = as.POSIXct("2021-06-28"), linetype = "dashed")+
  annotate("text",as.POSIXct("2021-06-28"),y=3415,label="Onset of Heat Dome",color="black")
D310plot_circ

ggsave(here("Plots/Ecomatik_adjustments", "D310plot_raw.png"), plot = D310plot_raw)
ggsave(here("Plots/Ecomatik_adjustments", "D310plot_circ.png"), plot = D310plot_circ)
```

# 311

```{r}
D311_manual_flags <- read.csv(here("Data/D311_manual_flags.csv"))
D311_manual_flags$Date<-as.POSIXct(D311_manual_flags$Date,format="%Y-%m-%d %H:%M:%S")
D311_manual_flags$DEND311_AVG_150_0_01[D311_manual_flags$DEND311_AVG_150_0_01 == ""] <- NA

#plotting raw data
D311plot_raw <- ggplot(D311_manual_flags,aes(Date,DEND311_AVG_150_0_01))+geom_point(alpha=0.7)+
  labs(title = "OG DF (ID 311) Raw Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()

#converting each raw data segment to circumference based on manual flag
D311_raw_A <- D311_manual_flags |> filter(Manual_flag=="A")
D311_circ_A <- dc3_to_circumference(
  df = D311_raw_A,
  date_column = D311_raw_A$Date, 
  Vi_micrometers = D311_raw_A$DEND311_AVG_150_0_01, 
  C0 = 2755, V0 = 4.738, na_method = "skip")

D311_raw_B <- D311_manual_flags |> filter(Manual_flag=="B")
D311_circ_B <- dc3_to_circumference(
  df = D311_raw_B,
  date_column = D311_raw_B$Date, 
  Vi_micrometers = D311_raw_B$DEND311_AVG_150_0_01, 
  C0 = 2779.422, V0 = 0.967, na_method = "skip")

D311_raw_C <- D311_manual_flags |> filter(Manual_flag=="C")
D311_circ_C <- dc3_to_circumference(
  df = D311_raw_C,
  date_column = D311_raw_C$Date, 
  Vi_micrometers = D311_raw_C$DEND311_AVG_150_0_01, 
  C0 = 2800.949, V0 = 5.560, na_method = "skip")

D311_raw_D <- D311_manual_flags |> filter(Manual_flag=="D")
D311_circ_D <- dc3_to_circumference(
  df = D311_raw_D,
  date_column = D311_raw_D$Date, 
  Vi_micrometers = D311_raw_D$DEND311_AVG_150_0_01, 
  C0 = 2820.472, V0 = 3.359, na_method = "skip")

D311_raw_E <- D311_manual_flags |> filter(Manual_flag=="E")
D311_circ_E <- dc3_to_circumference(
  df = D311_raw_E,
  date_column = D311_raw_E$Date, 
  Vi_micrometers = D311_raw_E$DEND311_AVG_150_0_01, 
  C0 = 2838.937, V0 = 2.509, na_method = "skip")

D311_raw_F <- D311_manual_flags |> filter(Manual_flag=="F")
D311_circ_F <- dc3_to_circumference(
  df = D311_raw_F,
  date_column = D311_raw_F$Date, 
  Vi_micrometers = D311_raw_F$DEND311_AVG_150_0_01, 
  C0 = 2855.429, V0 = 4.961, na_method = "skip")

#combining all of these into one df
D311_circ_A_F <- bind_rows(
  D311_circ_A,D311_circ_B,D311_circ_C,D311_circ_D,D311_circ_E,D311_circ_F)
D311_circ_A_F <- D311_circ_A_F |>
  mutate(difference = circumference_mm - lag(circumference_mm))

#plotting and saving circumference 
D311plot_circ <- ggplot(
  D311_circ_A_F,aes(date,circumference_mm))+geom_point(alpha=0.7) +
  labs(title = "OG DF (ID 311) Circumference",
    x = "Date",
    y = "Stem Circumference (mm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()+
  geom_vline(xintercept = as.POSIXct("2021-06-28"), linetype = "dashed")+
  annotate("text",as.POSIXct("2021-06-28"),y=2780,label="Onset of Heat Dome",color="black")

ggsave(here("Plots/Ecomatik_adjustments", "D311plot_raw.png"), plot = D311plot_raw)
ggsave(here(
  "Plots/Ecomatik_adjustments", "D311plot_circ.png"),plot = 
    D311plot_circ)
D311plot_circ
```

# 654

```{r}
D654_manual_flags <- read.csv(here("Data/D654_manual_flags.csv"))
D654_manual_flags$Date<-as.POSIXct(D654_manual_flags$Date,format="%Y-%m-%d %H:%M:%S")
D654_manual_flags$DEND654_AVG_150_0_01[D654_manual_flags$DEND654_AVG_150_0_01 == ""] <- NA

#plotting raw data
D654plot_raw <- ggplot(D654_manual_flags,aes(Date,DEND654_AVG_150_0_01))+geom_point(alpha=0.7)+
  labs(title = "2nd-Growth DF (ID 654) Raw Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()

#converting each raw data segment to circumference based on manual flag
D654_raw_A <- D654_manual_flags |> filter(Manual_flag=="A")
D654_circ_A <- dc3_to_circumference(
  df = D654_raw_A,
  date_column = D654_raw_A$Date, 
  Vi_micrometers = D654_raw_A$DEND654_AVG_150_0_01, 
  C0= 1170.5, V0 = 1.522, na_method = "skip") 

D654_raw_B <- D654_manual_flags |> filter(Manual_flag=="B")
D654_circ_B <- dc3_to_circumference(
  df = D654_raw_B,
  date_column = D654_raw_B$Date, 
  Vi_micrometers = D654_raw_B$DEND654_AVG_150_0_01, 
  C0= 1197.026, V0 = 3.949, na_method = "skip") 

D654_raw_C <- D654_manual_flags |> filter(Manual_flag=="C")
D654_circ_C <- dc3_to_circumference(
  df = D654_raw_C,
  date_column = D654_raw_C$Date, 
  Vi_micrometers = D654_raw_C$DEND654_AVG_150_0_01, 
  C0= 1213.621, V0 = 2.436, na_method = "skip") 

D654_raw_D <- D654_manual_flags |> filter(Manual_flag=="D")
D654_circ_D <- dc3_to_circumference(
  df = D654_raw_D,
  date_column = D654_raw_D$Date, 
  Vi_micrometers = D654_raw_D$DEND654_AVG_150_0_01, 
  C0= 1233.814, V0 = 5.484, na_method = "skip") 

D654_raw_E <- D654_manual_flags |> filter(Manual_flag=="E")
D654_circ_E <- dc3_to_circumference(
  df = D654_raw_E,
  date_column = D654_raw_E$Date, 
  Vi_micrometers = D654_raw_E$DEND654_AVG_150_0_01, 
  C0= 1249.523, V0 = 4.517, na_method = "skip") 

#combining all of these into one df
D654_circ_A_E <- bind_rows(
  D654_circ_A,D654_circ_B,D654_circ_C,D654_circ_D,D654_circ_E)
D654_circ_A_E <- D654_circ_A_E |>
  mutate(difference = circumference_mm - lag(circumference_mm))

#plotting and saving circumference 
D654plot_circ <- ggplot(
  D654_circ_A_E,aes(date,circumference_mm))+geom_point(alpha=0.7) +
  labs(title = "2nd-Growth DF (ID 654) Circumference",
    x = "Date",
    y = "Stem Circumference (mm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed")+
  annotate("text",as.POSIXct("2021-06-27"),y=1185,label="Onset of Heat Dome",color="black")

ggsave(here("Plots/Ecomatik_adjustments", "D654plot_raw.png"), plot = D654plot_raw)
ggsave(here(
  "Plots/Ecomatik_adjustments", "D654plot_circ.png"),plot = 
    D654plot_circ)
D654plot_circ
```

# 655

```{r}
D655_manual_flags <- read.csv(here("Data/D655_manual_flags.csv"))
D655_manual_flags$Date<-as.POSIXct(D655_manual_flags$Date,format="%Y-%m-%d %H:%M:%S")
D655_manual_flags$DEND655_AVG_150_0_01[D655_manual_flags$DEND655_AVG_150_0_01 == ""] <- NA

#plotting raw data
D655plot_raw <- ggplot(D655_manual_flags,aes(Date,DEND655_AVG_150_0_01))+geom_point(alpha=0.7)+
  labs(title = "2nd-Growth DF (ID 655) Raw Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()

#converting each raw data segment to circumference based on manual flag
D655_raw_A <- D655_manual_flags |> filter(Manual_flag=="A")
D655_circ_A <- dc3_to_circumference(
  df = D655_raw_A,
  date_column = D655_raw_A$Date, 
  Vi_micrometers = D655_raw_A$DEND655_AVG_150_0_01, 
  C0 = 884.8, V0 = 3.119, na_method = "skip")

D655_raw_B <- D655_manual_flags |> filter(Manual_flag=="B")
D655_circ_B <- dc3_to_circumference(
  df = D655_raw_B,
  date_column = D655_raw_B$Date, 
  Vi_micrometers = D655_raw_B$DEND655_AVG_150_0_01, 
  C0 = 917.7963, V0 = 6.072, na_method = "skip")

D655_raw_C <- D655_manual_flags |> filter(Manual_flag=="C")
D655_circ_C <- dc3_to_circumference(
  df = D655_raw_C,
  date_column = D655_raw_C$Date, 
  Vi_micrometers = D655_raw_C$DEND655_AVG_150_0_01, 
  C0 = 937.9167, V0 = 2.169, na_method = "skip")

D655_raw_D <- D655_manual_flags |> filter(Manual_flag=="D")
D655_circ_D <- dc3_to_circumference(
  df = D655_raw_D,
  date_column = D655_raw_D$Date, 
  Vi_micrometers = D655_raw_D$DEND655_AVG_150_0_01, 
  C0 = 956.3595, V0 = 2.610, na_method = "skip")

D655_raw_E <- D655_manual_flags |> filter(Manual_flag=="E")
D655_circ_E <- dc3_to_circumference(
  df = D655_raw_E,
  date_column = D655_raw_E$Date, 
  Vi_micrometers = D655_raw_E$DEND655_AVG_150_0_01, 
  C0 = 968.9284, V0 = 2.814, na_method = "skip")

#combining all of these into one df
D655_circ_A_E <- bind_rows(
  D655_circ_A,D655_circ_B,D655_circ_C,D655_circ_D,D655_circ_E)
D655_circ_A_E <- D655_circ_A_E |>
  mutate(difference = circumference_mm - lag(circumference_mm))

#plotting and saving circumference 
D655plot_circ <- ggplot(
  D655_circ_A_E,aes(date,circumference_mm))+geom_point(alpha=0.7) +
  labs(title = "2nd-Growth DF (ID 655) Circumference",
    x = "Date",
    y = "Stem Circumference (mm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()+
  geom_vline(xintercept = as.POSIXct("2021-06-28"), linetype = "dashed")+
  annotate("text",as.POSIXct("2021-06-28"),y=900,label="Onset of Heat Dome",color="black")

ggsave(here("Plots/Ecomatik_adjustments", "D655plot_raw.png"), plot = D655plot_raw)
ggsave(here(
  "Plots/Ecomatik_adjustments", "D655plot_circ.png"),plot = 
    D655plot_circ)
D655plot_circ
```

# 656

```{r}
D656_manual_flags <- read.csv(here("Data/D656_manual_flags.csv"))
D656_manual_flags$Date<-as.POSIXct(D656_manual_flags$Date,format="%Y-%m-%d %H:%M:%S") #make sure that this format matches what your date column is in Excel (I manually adjust to the above format before loading)
D656_manual_flags$DEND310_AVG_150_0_01[D656_manual_flags$DEND656_AVG_150_0_01 == ""] <- NA

#plotting raw data
D656plot_raw <- ggplot(D656_manual_flags,aes(Date,DEND656_AVG_150_0_01))+geom_point(alpha=0.7)+
  labs(title = "2nd-Growth DF (ID 656) Raw Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()

#converting each raw data segment to circumference based on manual flag
D656_raw_A <- D656_manual_flags |> filter(Manual_flag=="A")
D656_circ_A <- dc3_to_circumference(
  df = D656_raw_A,
  date_column = D656_raw_A$Date, 
  Vi_micrometers = D656_raw_A$DEND656_AVG_150_0_01, 
  C0 = 1221.3, V0 = 3.518, na_method = "skip")

D656_raw_B <- D656_manual_flags |> filter(Manual_flag=="B")
D656_circ_B <- dc3_to_circumference(
  df = D656_raw_B,
  date_column = D656_raw_B$Date, 
  Vi_micrometers = D656_raw_B$DEND656_AVG_150_0_01, 
  C0 = 1247.355, V0 = 4.555, na_method = "skip")

D656_raw_C <- D656_manual_flags |> filter(Manual_flag=="C")#2022-02-07 15:15:00
D656_circ_C <- dc3_to_circumference(
  df = D656_raw_C,
  date_column = D656_raw_C$Date, 
  Vi_micrometers = D656_raw_C$DEND656_AVG_150_0_01, 
  C0 = 1263.201, V0 = 3.097, na_method = "skip")

D656_raw_D <- D656_manual_flags |> filter(Manual_flag=="D")#2022-12-21 15:25:00
D656_circ_D <- dc3_to_circumference(
  df = D656_raw_D,
  date_column = D656_raw_D$Date, 
  Vi_micrometers = D656_raw_D$DEND656_AVG_150_0_01, 
  C0 = 1274.993, V0 = 4.258, na_method = "skip")

D656_raw_E <- D656_manual_flags |> filter(Manual_flag=="E")#2024-02-06 16:10:00
D656_circ_E <- dc3_to_circumference(
  df = D656_raw_E,
  date_column = D656_raw_E$Date, 
  Vi_micrometers = D656_raw_E$DEND656_AVG_150_0_01, 
  C0 = 1284.104, V0 = 2.167, na_method = "skip")

#combining all of these into one df
D656_circ_A_E <- bind_rows(
  D656_circ_A,D656_circ_B,D656_circ_C,D656_circ_D,D656_circ_E)
D656_circ_A_E <- D656_circ_A_E |>
  mutate(difference = circumference_mm - lag(circumference_mm))

#plotting and saving circumference 
D656plot_circ <- ggplot(
  D656_circ_A_E,aes(date,circumference_mm))+geom_point(alpha=0.7) +
  labs(title = "2nd-Growth DF (ID 656) Circumference",
    x = "Date",
    y = "Stem Circumference (mm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed")+
  annotate("text",as.POSIXct("2021-06-27"),y=1235,label="Onset of Heat Dome",color="black")

ggsave(here("Plots/Ecomatik_adjustments", "D656plot_raw.png"), plot = D656plot_raw)
ggsave(here(
  "Plots/Ecomatik_adjustments", "D656plot_circ.png"),plot = 
    D656plot_circ)
D656plot_circ
```

# 860

```{r}
D860_manual_flags <- read.csv(here("Data/D860_manual_flags.csv"))
D860_manual_flags$Date<-as.POSIXct(D860_manual_flags$Date,format="%Y-%m-%d %H:%M:%S")

#plotting raw data
D860plot_raw <- ggplot(D860_manual_flags,aes(Date,DEND860_AVG_150_0_01))+geom_point(alpha=0.7)+
  labs(title = "2nd-Growth DF (ID 860) Raw Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()

ggsave(here(
  "Plots/Ecomatik_adjustments", "D860plot_raw.png"),plot = 
    D860plot_raw)

#converting each raw data segment to circumference based on manual flag
D860_raw_A <- D860_manual_flags |> filter(Manual_flag=="A")
D860_circ_A <- dc3_to_circumference(
  df = D860_raw_A,
  date_column = D860_raw_A$Date, 
  Vi_micrometers = D860_raw_A$DEND860_AVG_150_0_01, 
  C0 = 760.3, V0 = 2.170, na_method = "skip")
#PLNG value is 2.340 mm but I disregard PLNG value because it doesn't match first valid v1 reading  and C0 record but either method leads to similar circ at the end of this time period is similar +/- 2 mm

D860_raw_B <- D860_manual_flags |> filter(Manual_flag=="B")
D860_circ_B <- dc3_to_circumference(
  df = D860_raw_B,
  date_column = D860_raw_B$Date, 
  Vi_micrometers = D860_raw_B$DEND860_AVG_150_0_01, 
  C0 = 791.0916, V0 = 3.500, na_method = "skip")

D860_raw_C <- D860_manual_flags |> filter(Manual_flag=="C")
D860_circ_C <- dc3_to_circumference(
  df = D860_raw_C,
  date_column = D860_raw_C$Date, 
  Vi_micrometers = D860_raw_C$DEND860_AVG_150_0_01, 
  C0 = 804.0259, V0 = 4.960, na_method = "skip")

#combining all of these into one df
D860_circ_A_C <- bind_rows(
  D860_circ_A,D860_circ_B,D860_circ_C)
D860_circ_A_C <- D860_circ_A_C |>
  mutate(difference = circumference_mm - lag(circumference_mm))

#plotting and saving circumference 
D860plot_circ <- ggplot(
  D860_circ_A_C,aes(date,circumference_mm))+geom_point(alpha=0.7) +
  labs(title = "2nd-Growth DF (ID 860) Circumference",
    x = "Date",
    y = "Stem Circumference (mm)")+
scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")),
    breaks = year_breaks19_24,
    labels = year_labels19_24
  )+theme_minimal()+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed")+
  annotate("text",as.POSIXct("2021-06-27"),y=765,label="Onset of Heat Dome",color="black")

ggsave(here("Plots/Ecomatik_adjustments", "D860plot_raw.png"), plot = D860plot_raw)
ggsave(here(
  "Plots/Ecomatik_adjustments", "D860plot_circ.png"),plot = 
    D860plot_circ)
D860plot_circ
```

# Circumference to BAI function

From circumference, we need diameter which is then converted to BA using the area of a circle formula:

![{\\displaystyle BA=\\pi \\times (DBH/2)\^{2}}](https://wikimedia.org/api/rest_v1/media/math/render/svg/a6247e795021667351c32ade0f0dc890ef364b3f){alt="{\\displaystyle BA=\\pi \\times (DBH/2)^{2}}"}

If DBH was measured in cm, BA will be in cm^2^. To convert to m^2^, divide by 10,000:

BA(m2)=π×(DBH(cm)/2)210,000

```{r}
circ_to_bai <- function(df, date_col = NULL, radius_col, circumference_col) {
  
  # Check if dplyr is available
  if (!requireNamespace("dplyr", quietly = TRUE)) {
    stop("The 'dplyr' package is required. Please install it using install.packages('dplyr').")
  }
  
  # If date_col is provided, check chronological order
  if (!is.null(substitute(date_col))) {
    date_values <- df %>% dplyr::pull({{ date_col }})
    if (is.unsorted(date_values[!is.na(date_values)])) {
      warning("Date column is not in chronological order. BAI calculations may be incorrect.")
    }
  }
    
    result <- df %>%
      dplyr::mutate(
        basal_area_mm2 = pi * radius_mm^2,
        bai_mm2 = c(NA, diff(basal_area_mm2))
      )
  
  return(result)
}
```

# BAI to temporal aggregation

```{r}
hourly_bai_stats <- function(df, date_col, circumference_col, bai_col) {
  if (!requireNamespace("dplyr", quietly = TRUE)) {
    stop("Please install the 'dplyr' package.")
  }
  if (!requireNamespace("lubridate", quietly = TRUE)) {
    stop("Please install the 'lubridate' package.")
  }

  df_with_bai <- df %>%
    dplyr::mutate(
      date = {{ date_col }},
      circumference_mm = {{ circumference_col }},
      bai_mm2 = {{ bai_col }}
    ) %>%
    dplyr::filter(!is.na(bai_mm2))

  hourly_bai <- df_with_bai %>%
    dplyr::mutate(
      date_hour = lubridate::floor_date(date, "hour")
    ) %>%
    dplyr::group_by(date_hour) %>%
    dplyr::summarise(
      total_hourly_bai_mm2 = sum(bai_mm2, na.rm = TRUE),
      max_circumference_mm = max(circumference_mm, na.rm = TRUE),
      min_circumference_mm = min(circumference_mm, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    dplyr::rename(date = date_hour)

  return(hourly_bai)
}

#example usage
#hourly_df <- hourly_bai_stats(df, date, circumference_mm, bai_mm2)



daily_bai_stats <- function(df, date_col, circumference_col, basal_area_col, bai_col) {
  if (!requireNamespace("dplyr", quietly = TRUE)) {
    stop("Please install the 'dplyr' package.")
  }
  if (!requireNamespace("lubridate", quietly = TRUE)) {
    stop("Please install the 'lubridate' package.")
  }

  df_with_bai <- df %>%
    dplyr::mutate(
      date = {{ date_col }},
      circumference_mm = {{ circumference_col }},
      basal_area_mm2 = {{ basal_area_col }},
      bai_mm2 = {{ bai_col }}
    ) %>%
    dplyr::filter(!is.na(bai_mm2))

  daily_bai <- df_with_bai %>%
    dplyr::mutate(date_day = lubridate::as_date(date)) %>%
    dplyr::group_by(date_day) %>%
    dplyr::summarise(
      total_bai_mm2 = sum(bai_mm2, na.rm = TRUE),
      max_bai_mm2 = max(bai_mm2, na.rm = TRUE),
      daily_circ_amplitude_mm = max(
        circumference_mm, na.rm = TRUE) - min(
          circumference_mm, na.rm = TRUE),
      daily_bai_amplitude_mm2 = max(
        bai_mm2, na.rm = TRUE) - min(bai_mm2, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    dplyr::arrange(date_day) %>%
    dplyr::mutate(
      cumu_bai_mm2 = cumsum(total_bai_mm2),
      cumu_bai_mm2_norm = cumu_bai_mm2 / max(cumu_bai_mm2, na.rm = TRUE),
      delta_max_bai_mm2 = c(NA, diff(max_bai_mm2)),
      DOY = lubridate::yday(date_day),
      year = as.factor(lubridate::year(date_day))
    ) %>%
    dplyr::rename(date = date_day)

  return(daily_bai)
}

#example usage
#daily_df  <- daily_bai_stats(df, date, circumference_mm, basal_area_mm2, bai_mm2)
```

# Doing the conversions and plotting

## 310

```{r}
#310
D310_bai <- circ_to_bai(
  D310_circ_A_E, date, radius_mm, circumference_mm)
D310_hourly_bai_stats <- hourly_bai_stats(
  D310_bai, date, circumference_mm, bai_mm2)
D310_daily_bai_stats <- daily_bai_stats(
  D310_bai, date, circumference_mm, basal_area_mm2, bai_mm2)

# adding metadata to daily stats
D310_daily_bai_stats$date <- as.POSIXct(D310_daily_bai_stats$date)
D310_daily_bai_stats <- D310_daily_bai_stats %>%
  mutate(
    ID = "310",
    AgeClass = "OG")

ggplot(D310_bai, aes(x=date,y=bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")+ scale_y_continuous(
    limits = c(-50,250))

ggplot(D310_daily_bai_stats, aes(x=date,y=total_bai_mm2,color))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D310_daily_bai_stats, aes(
  x=date,y=delta_max_bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")+ scale_y_continuous(
    limits = c(-250,250))

ggplot(D310_daily_bai_stats, aes(
  x=date,y=cumu_bai_mm2_norm))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")
```

## 311

```{r}
#311
D311_bai <- circ_to_bai(
  D311_circ_A_F, date, radius_mm, circumference_mm)
D311_hourly_bai_stats <- hourly_bai_stats(
  D311_bai, date, circumference_mm, bai_mm2)
D311_daily_bai_stats <- daily_bai_stats(
  D311_bai, date, circumference_mm, basal_area_mm2, bai_mm2)

# adding metadata to daily stats
D311_daily_bai_stats$date <- as.POSIXct(D311_daily_bai_stats$date)
D311_daily_bai_stats <- D311_daily_bai_stats %>%
  mutate(
    ID = "311",
    AgeClass = "OG")

ggplot(D311_bai, aes(x=date,y=bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D311_daily_bai_stats, aes(x=date,y=total_bai_mm2,color))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D311_daily_bai_stats, aes(
  x=date,y=delta_max_bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D311_daily_bai_stats, aes(
  x=date,y=cumu_bai_mm2_norm))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")
```

## 654

```{r}
#654
D654_bai <- circ_to_bai(
  D654_circ_A_E, date, radius_mm, circumference_mm)
D654_hourly_bai_stats <- hourly_bai_stats(
  D654_bai, date, circumference_mm, bai_mm2)
D654_daily_bai_stats <- daily_bai_stats(
  D654_bai, date, circumference_mm, basal_area_mm2, bai_mm2)

# adding metadata to daily stats
D654_daily_bai_stats$date <- as.POSIXct(D654_daily_bai_stats$date)
D654_daily_bai_stats <- D654_daily_bai_stats %>%
  mutate(
    ID = "654",
    AgeClass = "2nd-growth")

ggplot(D654_bai, aes(x=date,y=bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D654_daily_bai_stats, aes(x=date,y=total_bai_mm2,color))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D654_daily_bai_stats, aes(
  x=date,y=delta_max_bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D654_daily_bai_stats, aes(
  x=date,y=cumu_bai_mm2_norm))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")
```

## 655

```{r}
#655
D655_bai <- circ_to_bai(
  D655_circ_A_E, date, radius_mm, circumference_mm)
D655_hourly_bai_stats <- hourly_bai_stats(
  D655_bai, date, circumference_mm, bai_mm2)
D655_daily_bai_stats <- daily_bai_stats(
  D655_bai, date, circumference_mm, basal_area_mm2, bai_mm2)

# adding metadata to daily stats
D655_daily_bai_stats$date <- as.POSIXct(D655_daily_bai_stats$date)
D655_daily_bai_stats <- D655_daily_bai_stats %>%
  mutate(
    ID = "655",
    AgeClass = "2nd-growth")

ggplot(D655_bai, aes(x=date,y=bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D655_daily_bai_stats, aes(x=date,y=total_bai_mm2,color))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D655_daily_bai_stats, aes(
  x=date,y=delta_max_bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D655_daily_bai_stats, aes(
  x=date,y=cumu_bai_mm2_norm))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")
```

## 656

```{r}
#656
D656_bai <- circ_to_bai(
  D656_circ_A_E, date, radius_mm, circumference_mm)
D656_hourly_bai_stats <- hourly_bai_stats(
  D656_bai, date, circumference_mm, bai_mm2)
D656_daily_bai_stats <- daily_bai_stats(
  D656_bai, date, circumference_mm, basal_area_mm2, bai_mm2)

# adding metadata to daily stats
D656_daily_bai_stats$date <- as.POSIXct(D656_daily_bai_stats$date)
D656_daily_bai_stats <- D656_daily_bai_stats %>%
  mutate(
    ID = "656",
    AgeClass = "2nd-growth")

ggplot(D656_bai, aes(x=date,y=bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D656_daily_bai_stats, aes(x=date,y=total_bai_mm2,color))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D656_daily_bai_stats, aes(
  x=date,y=delta_max_bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D656_daily_bai_stats, aes(
  x=date,y=cumu_bai_mm2_norm))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")
```

## 860

```{r}
#860
D860_bai <- circ_to_bai(
  D860_circ_A_C, date, radius_mm, circumference_mm)
D860_hourly_bai_stats <- hourly_bai_stats(
  D860_bai, date, circumference_mm, bai_mm2)
D860_daily_bai_stats <- daily_bai_stats(
  D860_bai, date, circumference_mm, basal_area_mm2, bai_mm2)

# adding metadata to daily stats
D860_daily_bai_stats$date <- as.POSIXct(D860_daily_bai_stats$date)
D860_daily_bai_stats <- D860_daily_bai_stats %>%
  mutate(
    ID = "860",
    AgeClass = "2nd-growth")


ggplot(D860_bai, aes(x=date,y=bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D860_daily_bai_stats, aes(x=date,y=total_bai_mm2,color))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D860_daily_bai_stats, aes(
  x=date,y=delta_max_bai_mm2))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")

ggplot(D860_daily_bai_stats, aes(
  x=date,y=cumu_bai_mm2_norm))+geom_line()+
  scale_x_datetime(
    limits = as.POSIXct(c("2019-01-01", "2024-10-01")))+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")



```

## Combining into one df

```{r}
daily_bai_stats_6dendros <- bind_rows(
  D310_daily_bai_stats,D311_daily_bai_stats,D654_daily_bai_stats,
  D655_daily_bai_stats,D656_daily_bai_stats,D860_daily_bai_stats)
daily_bai_stats_6dendros$ID <- as.factor(daily_bai_stats_6dendros$ID)
daily_bai_stats_6dendros$AgeClass <- as.factor(daily_bai_stats_6dendros$AgeClass)

ggplot(daily_bai_stats_6dendros, aes(
  x=date,y=cumu_bai_mm2_norm,color=ID))+geom_line(size=1)+
  geom_vline(xintercept = as.POSIXct("2021-06-27"), linetype = "dashed",color="red")
```

# Putting into a model

## Further sorting the combined 6 dendros df  

```{r}
library(emmeans)
library(nlme)
library(lubridate)

#eliminating 2024 for ID 860
daily_bai_stats_6dendros <- subset(daily_bai_stats_6dendros, !(ID == 860 & year == 2024))
daily_bai_stats_6dendros <- subset(daily_bai_stats_6dendros, !(year == 2018))

#truncating df to be HD +/- 4 weeks
daily_bai_stats_6dendros_truncated <- daily_bai_stats_6dendros |>
  filter(
    (month(date) == 5 & day(date) >= 30) |
    (month(date) == 6) |
    (month(date) == 7 & day(date) <= 26))

#adding period column
daily_bai_stats_6dendros_truncated <- daily_bai_stats_6dendros_truncated |>
  mutate(
    period = case_when(
      (month(date) == 5 & day(date) >= 30) |
      (month(date) == 6 & day(date) < 27) ~ "before",
      (month(date) == 6 & day(date) %in% c(27, 28)) ~ "during",
      (month(date) == 6 & day(date) > 28) |
      (month(date) > 6) ~ "after",
      TRUE ~ NA_character_))

# Ensuring Period is chronologically ordered 
daily_bai_stats_6dendros_truncated$period <- factor(daily_bai_stats_6dendros_truncated$period, levels = c("before", "during", "after"))

#Days are repeating so let's do Julian days instead of DOY days
daily_bai_stats_6dendros_truncated <- daily_bai_stats_6dendros_truncated |>
  arrange(date) |>
  mutate(
    day_integer = as.numeric(as.Date(date) - min(as.Date(date)) + 1))
```

## Making different model options 

```{r}
model_cumu_bai_norm <- lme(
  cumu_bai_mm2_norm ~ year*period + AgeClass,
  data = daily_bai_stats_6dendros_truncated,
  random = list(~ 1 | ID), 
  correlation = corARMA(p=1,form = ~ day_integer | ID)) 
#p=2 led to "Error in `coef<-.corARMA`(`*tmp*`, value = value[parMap[, i]]) : Coefficient matrix not invertible""

model_total_daily_bai<- lme(
  total_bai_mm2 ~ year*period + AgeClass,
  data = daily_bai_stats_6dendros_truncated,
  random = list(~ 1 | ID), 
  correlation = corARMA(p=2,form = ~ day_integer | ID)) 

model_delta_max_bai<- lme(
  delta_max_bai_mm2 ~ year*period + AgeClass,
  data = daily_bai_stats_6dendros_truncated,
  random = list(~ 1 | ID), 
  correlation = corARMA(p=2,form = ~ day_integer | ID)) 

AIC(model_cumu_bai_norm,model_total_daily_bai,model_delta_max_bai)
```

### Diagnoses of model based on cumu_bai_mm2_norm

```{r}
summary(model_cumu_bai_norm)
anova(model_cumu_bai_norm) #for direct year:period interaction
daily_bai_stats_6dendros_truncated$residuals_norm_CBN <- residuals(
  model_cumu_bai_norm,type="normalized")

# residuals vs year (box plot)
boxplot(residuals_norm_CBN ~ year, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs Year",
        xlab = "Year", ylab = "Normalized Residuals")

# Residuals vs period
boxplot(residuals_norm_CBN ~ period, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs Period",
        xlab = "Period", ylab = "Normalized Residuals")

# Residuals vs AgeClass
boxplot(residuals_norm_CBN ~ AgeClass, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs AgeClass",
        xlab = "AgeClass", ylab = "Normalized Residuals")

#ACF and PACF plots 
# step 1 - Get combinations of ID and year
id_year_combos <- unique(
  daily_bai_stats_6dendros_truncated[, c("ID", "year")])

# step 2 - Loop through each combination
for (i in 1:nrow(id_year_combos)) {
  this_id <- id_year_combos$ID[i]
  this_year <- id_year_combos$year[i]

  # step 3 - Subset residuals for that ID and year
  subset_df <- subset(
    daily_bai_stats_6dendros_truncated, 
    ID == this_id & year == this_year)

  # step 4 - make the plots 
  par(mfrow = c(1, 2))
  acf(subset_df$residuals_norm_CBN, main = paste("CBN ACF: ID", this_id, this_year), na.action = na.pass)
  pacf(subset_df$residuals_norm_CBN, main = paste("CBN PACF: ID", this_id, this_year), na.action = na.pass)
}
```

### Diagnoses of model based on total_bai_mm2

```{r}
summary(model_total_daily_bai)
anova(model_total_daily_bai) #for direct year:period interaction
daily_bai_stats_6dendros_truncated$residuals_norm_TDB <- residuals(
  model_total_daily_bai,type="normalized")

# residuals vs year (box plot)
boxplot(residuals_norm_TDB ~ year, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs Year",
        xlab = "Year", ylab = "Normalized Residuals")

# Residuals vs period
boxplot(residuals_norm_TDB ~ period, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs Period",
        xlab = "Period", ylab = "Normalized Residuals")

# Residuals vs AgeClass
boxplot(residuals_norm_TDB ~ AgeClass, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs AgeClass",
        xlab = "AgeClass", ylab = "Normalized Residuals")

#ACF and PACF plots 
# step 1 - Get combinations of ID and year
id_year_combos <- unique(
  daily_bai_stats_6dendros_truncated[, c("ID", "year")])

# step 2 - Loop through each combination
for (i in 1:nrow(id_year_combos)) {
  this_id <- id_year_combos$ID[i]
  this_year <- id_year_combos$year[i]

  # step 3 - Subset residuals for that ID and year
  subset_df <- subset(
    daily_bai_stats_6dendros_truncated, 
    ID == this_id & year == this_year)

  # step 4 - make the plots
  par(mfrow = c(1, 2))
  acf(subset_df$residuals_norm_TDB, main = paste("TDB ACF: ID", this_id, this_year), na.action = na.pass)
  pacf(subset_df$residuals_norm_TDB, main = paste("TDB PACF: ID", this_id, this_year), na.action = na.pass)
}
```

### Diagnoses of model based on delta_max_bai_mm2

```{r}
summary(model_delta_max_bai)
anova(model_delta_max_bai) #for direct year:period interaction
daily_bai_stats_6dendros_truncated$residuals_norm_DMB <- residuals(
  model_delta_max_bai,type="normalized")

# residuals vs year (box plot)
boxplot(residuals_norm_DMB ~ year, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs Year",
        xlab = "Year", ylab = "Normalized Residuals")

# Residuals vs period
boxplot(residuals_norm_DMB ~ period, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs Period",
        xlab = "Period", ylab = "Normalized Residuals")

# Residuals vs AgeClass
boxplot(residuals_norm_DMB ~ AgeClass, 
        data = daily_bai_stats_6dendros_truncated,
        main = "Residuals vs AgeClass",
        xlab = "AgeClass", ylab = "Normalized Residuals")

#ACF and PACF plots 
# step 1 - Get combinations of ID and year
id_year_combos <- unique(
  daily_bai_stats_6dendros_truncated[, c("ID", "year")])

# step 2 - Loop through each combination
for (i in 1:nrow(id_year_combos)) {
  this_id <- id_year_combos$ID[i]
  this_year <- id_year_combos$year[i]

  # step 3 - Subset residuals for that ID and year
  subset_df <- subset(
    daily_bai_stats_6dendros_truncated, 
    ID == this_id & year == this_year)

  # step 4 - make the plots 
  par(mfrow = c(1, 2))
  acf(subset_df$residuals_norm_DMB, main = paste("DMB ACF: ID", this_id, this_year), na.action = na.pass)
  pacf(subset_df$residuals_norm_DMB, main = paste("DMB PACF: ID", this_id, this_year), na.action = na.pass)
}
```

## Using emmeans to define contrasts for each model

### Age classes averaged

```{r}
#getting EMM for each period nested within year
emm_CBM <- emmeans(model_cumu_bai_norm, ~ period | year | AgeClass)
emm_TDB <- emmeans(model_total_daily_bai, ~ period | year | AgeClass)
emm_DMB <- emmeans(model_delta_max_bai, ~ period | year | AgeClass)

#defining contrast to match period order
contrast_CBM <- contrast(
  emm_CBM, method = list(after_vs_before = c(-1, 0, 1)), 
  by = c("year", "AgeClass"))
contrast_TDB <- contrast(
  emm_TDB, method = list(after_vs_before = c(-1, 0, 1)), 
  by = c("year", "AgeClass"))
contrast_DMB <- contrast(
  emm_DMB, method = list(after_vs_before = c(-1, 0, 1)), 
  by = c("year", "AgeClass"))

#getting summaries of the contrasts
summary(contrast_CBM)
summary(contrast_TDB)
summary(contrast_DMB)

#getting pairwise comparisons of before-after effects between years
pairs(contrast_CBM, by = c("year","AgeClass"))
pairs(contrast_TDB, by = c("year","AgeClass"))
pairs(contrast_DMB, by = c("year","AgeClass"))
```

### Separate IDs

```{r}
# Step 1: Filter to just before and after
daily_bai_stats_6dendros_truncated_BandA <- daily_bai_stats_6dendros_truncated %>%
  filter(period %in% c("before", "after"))

# Group and compute mean value per group and contrast
contrast_CBM_per_id <- daily_bai_stats_6dendros_truncated_BandA %>%
  group_by(ID, year, period) %>%
  summarise(
    mean_cumu_bai_mm2_norm = mean(cumu_bai_mm2_norm, na.rm = TRUE), 
    .groups = "drop") %>%
  pivot_wider(names_from = period, 
              values_from = mean_cumu_bai_mm2_norm) %>%
  mutate(contrast_after_minus_before = after - before)

contrast_TBM_per_id <- daily_bai_stats_6dendros_truncated_BandA %>%
  group_by(ID, year, period) %>%
  summarise(
    mean_total_bai_mm2 = mean(total_bai_mm2, na.rm = TRUE), 
    .groups = "drop") %>%
  pivot_wider(names_from = period, 
              values_from = mean_total_bai_mm2) %>%
  mutate(contrast_after_minus_before = after - before)

contrast_DMB_per_id <- daily_bai_stats_6dendros_truncated_BandA %>%
  group_by(ID, year, period) %>%
  summarise(
    mean_delta_max_bai_mm2 = mean(delta_max_bai_mm2, na.rm = TRUE), 
    .groups = "drop") %>%
  pivot_wider(names_from = period, 
              values_from = mean_delta_max_bai_mm2) %>%
  mutate(contrast_after_minus_before = after - before)

# View results
print(contrast_CBM_per_id)
print(contrast_TBM_per_id)
print(contrast_DMB_per_id)
```
