---
title: "Cleaning/Reading with TNP
author: "Gabby John"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##R script for reading and cleaning dendrometer data and exporting analysis files

# Installing for the first time

```{r}
packages <- (c("devtools","zoo","chron","dplyr","viridis", "RCurl", "DT","lubridate"))
install.packages(setdiff(packages, rownames(installed.packages())))
devtools::install_github("treenet/treenetproc")
```

```{r}
library(treenetproc, `force = TRUE`)
library(zoo)
library(chron)
library(viridis)
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r}
getwd()
```

# Example with Chaney's data

## step 0: identify helper functions

```{r}
left <- function(string, char){substr(string, 1,char)} 
right <- function (string, char){substr(string,nchar(string)-(char-1),nchar(string))}
```

## step 1: read in dendrometer data and plot raw data

```{r}
all_data_140<-read.table('data_92223140_2023_04_23_0.csv', header=F, sep=";", dec=',') 
str(all_data_140) 
dendro_data_140=all_data_140[,c(2,7)] 
colnames(dendro_data_140) <- c('datetime','value')
```

## Step 2: fix timestamps (ts) and connect with dendrometer ID number

```{r}
ID = '140'
ts = as.POSIXct(dendro_data_140['datetime'][,],format="%Y.%m.%d %H:%M",tz="GMT") 
d140_L0<-cbind(dendro_data_140["value"]+0,ts,series=ID)
```

## Step 3: set parameters to plot raw data

```{r}
#First grab years string
years <-left(d140_L0[,"ts"],4)
#To plot, set number of plots with par()
par(mfrow=c(1,1))
#Then set margins 
par(mar = c(5, 5, 5, 5))
```

## Step 4: plotting raw data for loop plots data iteratively.

```{r}
for(y in 1:length(unique(years))){
#Now we can select years
sel<-d140_L0[which(years==unique(years)[y]),]
#But we need to handle first year first
if(y==1){
   plot(difftime(as.POSIXct(sel$ts,format="%Y-%m-%d %H:%M:%S",tz="GMT"),
  as.POSIXct(paste0(unique(years)[y],"-01-01 00:00:00"),
                  format="%Y-%m-%d %H:%M:%S",tz="GMT"),
                  units = "days"),
         sel$value,
         ylab=expression("L0 ("*mu*"m)"),
         xlab="Day of year",type="l",
         col=viridis(length(unique(years)))[y],
         xlim=c(0,365),
         ylim=c(min(d140_L0$value,na.rm=T),
                max(d140_L0$value,na.rm=T)),
         main=unique(d140_L0$series))
    
    legend("bottomright",
           as.character(unique(years)[-4]),
           col=viridis(length(unique(years))),
           bty="n",lty=1)
# Now we can add other years
  }else{
    lines(difftime(as.POSIXct(sel$ts,format="%Y-%m-%d %H:%M:%S",tz="GMT"),
                   as.POSIXct(paste0(unique(years)[y],"-01-01 00:00:00"),
                              format="%Y-%m-%d %H:%M:%S",tz="GMT"),
                   units = "days"),
          sel$value,
          col=viridis(length(unique(years)))[y])}
}
```

## Step 5: cleaning the dendrometery data

```{r}
#Level 1 processing (time align) method 1
?treenetproc::proc_L1 
d140_L1<- proc_L1(data_L0 = d140_L0, reso = 60 ) 
                  #input = "wide", date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT") 
head(d140_L1)

#Level 1 processing (time align) method 2
d140_L1<- proc_L1(data_L0 = d140_L0, reso = 60,
                  date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT")
head(d140_L1)

#level 1 processing of temperature data
temp_data_140_L0 <- cbind(all_data_140[,c(4,6)],ts,series=ID)
colnames(temp_data_140_L0) <-c('value','empty','ts','series')

#Now we can time-align temperature data with proc_L1
temp_data_140_L1 <- proc_L1(data_L0 = temp_data_140_L0, reso = 60, date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT") 
head(temp_data_140_L1)
```

Now we can move on to proc_dendro_L2, which integrates dendrometer readings with temperature, detects errors, corrects them and plots gro and twd by date.more info here: ?treenetproc::proc_dendro_L2

first plot full dataset for dendrometer. Plot will be exported to working directory as pdf

```{r}
#first plot full dataset for dendrometer. NOTE: Plot will be exported to working directory as pdf
d140_L2 <- proc_dendro_L2(dendro_L1 = d140_L1, temp_L1 = temp_data_140_L1, tol_out = 2, tol_jump = 3, plot = TRUE, frost_thr = 1, plot_period = "full", plot_export = TRUE, interpol = 3.5*60, frag_len = NULL, plot_name = "Slice_butte_2022_23_140_full", tz="GMT") 
graphics.off()
```

Looking at this we can see that the tolerance for outliers and tolerance for jumps were set lower (more stringent) than default value. We can plot monthly instead and look at exported pdf for greater detail.

```{r}
d140_L2_monthly <- proc_dendro_L2(dendro_L1 = d140_L1, temp_L1 = temp_data_140_L1, tol_out = 2, tol_jump = 3, frost_thr = 1, plot = TRUE, interpol = 3.5*60, plot_period = "monthly", plot_export = TRUE, plot_name = "Slice_butte_2022_23_140_monthly", tz="GMT") 
graphics.off()
```

We can manually correct some areas where automated data cleaning was not up to snuff.

```{r}
# First off, delete data from before installation on June 22st 2022 and the first week after installation 
#info on function here: ?corr_dendro_L2

d140_L2_corr1 <- corr_dendro_L2(dendro_L1 = d140_L1, dendro_L2 = d140_L2, delete = c("2022-05-10","2022-06-29"), n_days = 2, plot = TRUE, plot_export = TRUE, tz="GMT")

#check errors in pdf and with this subset 
d140_L2_corr1[which((d140_L2_corr1$flags)!="NA"),]
```

## Step 6: plot!

```{r}
ggplot((subset(d140_L2_corr1, frost == "FALSE")), aes(x=ts))+ geom_line(aes(y=value), color = "grey70")+ geom_line(aes(y=gro_yr), color = "seagreen")+ geom_line(aes(y=twd), color = "red")
```

What if we want to visualize cumulative growth instead of growth for each year?

```{r}
#create column for separate years 
d140_L2_corr1$yr <- substr(d140_L2_corr1$ts,1,4)
#Turn these columns into separate dataframes for each year
d140_2023_growth <- d140_L2_corr1[which(d140_L2_corr1$yr == "2023"),] 
d140_2022_growth <- d140_L2_corr1[which(d140_L2_corr1$yr == "2022"),]

#for 2023, add that years growth to 2022's maximum growth of 2214 µm #for 2022, make a copy of gro_yr

d140_2023_growth$gro_tot <- d140_2023_growth$gro_yr + 2214
d140_2022_growth$gro_tot <- d140_2022_growth$gro_yr

#recombine both years data frames into one. 
d_140_L3 <- rbind(d140_2022_growth, d140_2023_growth)

#Now plot!
ggplot((subset(d_140_L3, frost == "FALSE")), aes(x=ts))+ geom_line(aes(y=value), color = "grey70")+ geom_line(aes(y=gro_tot), color = "seagreen")+ geom_line(aes(y=twd), color = "red")
```

What if we just want ro look at the growing season?

```{r}
ggplot((subset(d_140_L3, ts < as.POSIXct("2022-10-23 00:00") & ts > as.POSIXct("2022-06-29 00:00"))), aes(x=ts))+ geom_line(aes(y=value), color = "grey70")+ geom_line(aes(y=gro_tot), color = "seagreen")+ geom_line(aes(y=twd), color = "red")

#The function phase_stats() will compute rates of expansion(growth) and shrinkage (twd) and aggregate these on a daily basis. 
#More info: ?phase_stats

d140_phase <- phase_stats(d140_L2_corr1, plot_phase = TRUE, plot_export = TRUE, agg_daily = TRUE, tz = 'GMT' )

#These growth and shrink rates can be compared with temperature, VPD and soil moisture.
#We will likely restrict this analysis to the growing season.
#some preliminary analysis/exploration using this

ggplot(subset(d140_phase, doy > 180 & doy < 250), aes(x=doy))+ geom_point(aes(y = shrink_amp, colour = "Shrinkage" ))+ geom_point(aes(y = exp_amp, colour ="Expansion"))+ ylab("Amplitude")

ggplot(subset(d140_phase, doy > 180 & doy < 250), aes(x=doy))+ geom_point(aes(y = shrink_slope, colour = "Shrinkage" ))+ geom_point(aes(y = exp_slope, colour ="Expansion"))+ ylab("Slope")

ggplot(subset(d140_phase, doy > 180 & doy < 250), aes(x=shrink_amp, y = exp_amp))+ geom_point()+ ylab(" Expansion amplitude")+ xlab(" Shrinkage amplitude")

ggplot(subset(d140_phase, doy > 180 & doy < 250), aes(x=shrink_slope, y = exp_slope))+ geom_point()+ ylab(" Expansion slope")+ xlab(" Shrinkage slope")
```

## Step 6: Export

Now that data is cleaned, we can export for analysis with weather data.

```{r}
#for plotting 
write.csv(file = "dendro_140_cleaned.csv", d_140_L3)

#daily growth stats 
write.csv(file = "dendro_140_phase_stats.csv", d140_phase)
```

# 
