---
title: "Untitled"
author: "Gabby John"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(treenetproc, `force = TRUE`)
library(ggplot2)
library(lubridate)
library(tidyr)
library(scales)

left <- function(string, char){substr(string, 1,char)} 
right <- function (string, char){substr(string,nchar(string)-(char-1),nchar(string))}
```


```{r}
CEN5180ts<- read.csv("~/research/TOMST/CEN_5180_ts_DMY.csv",stringsAsFactors = TRUE)
CEN5180ts$value<-as.numeric(CEN5180ts$value)
CEN5180ts$ts<-parse_date_time(CEN5180ts$ts,orders=c("dmy","dmy HMS"))
CEN5180ts$yr<-format(CEN5180ts$ts, "%Y")
CEN5180ts$yr<-as.factor(CEN5180ts$yr)


CEN5180temp<- read.csv("~/research/TOMST/CEN_5180_temp.csv",stringsAsFactors = TRUE)
CEN5180temp$value<-as.numeric(CEN5180temp$value)
CEN5180temp$ts<-parse_date_time(CEN5180temp$ts,orders=c("dmy","dmy HMS"))
CEN5180temp$yr<-format(CEN5180temp$ts, "%Y")
CEN5180temp$yr<-as.factor(CEN5180temp$yr)
write.csv(file="CEN5180temp.csv",CEN5180temp)

CEN5180ts$series<-"CEN5180"
CEN5180temp$series<-"CEN5180"
CEN5180temp$version<-"N.A"


ggplot(CEN5180ts,aes(x=ts,y=value,color=yr))+
  geom_point(na.rm=TRUE)+
  labs(title="CEN5180 Stem Diameter Changes",x="Date",y="Stem Diameter Change
  (µm)",color="Legend") +
  guides(color=guide_legend(override.aes = list(alpha = 1, size = 5)))+
  theme_minimal()
```
Including the temperature dataset increases the thresholds for jump and outlier
detection in periods of probable frost (i.e. when temperature < `lowtemp`), 
which are indicated with a horizontal cyan line in the plot.

The increase of the thresholds for outlier detection prevented the 
classification of the values during the frost shrinkage as outliers.

```{r}
CEN5180_L2 <- proc_dendro_L2(dendro_L1 = CEN5180ts,temp_L1 = CEN5180temp, tol_out = 2, tol_jump = 3, plot = TRUE, frost_thr = 1, plot_period = "full", plot_export = TRUE, interpol = 3.5*60, frag_len = NULL, plot_name = "CEN5180", tz="GMT") 

#Frost info
par(mar = c(2.1, 4, 1, 2))
plot(data = CEN5180_L2, value ~ ts, type = "n", 
     ylab = paste0("L2 (", "\u00b5", "m)"), las = 1)
lines(data = CEN5180ts, value ~ ts, col = "grey70")
lines(data = CEN5180_L2, value ~ ts, col = "#08519c")
# add abline to indicate frost
CEN5180_L2_frost <- CEN5180_L2 %>% 
  dplyr::rename(value_L2 = value)
plot_frost_period(df = CEN5180_L2_frost)

#L2 processing 
ggplot((subset(CEN5180_L2, frost == "FALSE")), aes(x=ts))+ geom_point(aes(y=value), color = "grey70")+ geom_point(aes(y=gro_yr), color = "seagreen")+ geom_point(aes(y=twd), color = "red")+labs(title="CEN5180 dendrometer values (gray), annual growth (green), twd (red)")
```
#Phase stats and growing season
Several characteristics of shrinkage and expansion phases can be calculated
with the function `phase_stats`.

It returns a `data.frame` containing the timing, duration, amplitude
and slope of the shrinkage (`shrink`) and expansion (`exp`) phases (Table 5). 
In addition, the parameter `phase_class` identifies days on which radial change
is likely driven by transpiration (`phase_class = 1`) or temperature 
(`phase_class = -1`).

To evaluate the identification of shrinkage and expansion phases, all phases 
can be plotted by setting `plot_phase = TRUE`. The plot shows the 
maximum (filled circle) and minimum (filled triangle) of the respective phase
and reports its statistics. Empty circles and triangles show maxima or minima
of previous or subsequent phases.
```{r}
CEN5180_L2_monthly <- proc_dendro_L2(dendro_L1 = CEN5180ts, temp_L1 = CEN5180temp, tol_out = 2, tol_jump = 3, frost_thr = 1, plot = TRUE, interpol = 3.5*60, plot_period = "monthly", plot_export = TRUE, plot_name = "CEN5180_L2_monthly", tz="GMT") 

# Calculate phase statistics as many separate graphs
phase_stats(dendro_L2 = CEN5180_L2, plot_phase = TRUE, plot_export = TRUE)

# calculate phase_stats as a table
phase_stats <- phase_stats(dendro_L2 = CEN5180_L2)
knitr::kable(phase_stats[1:5, ], caption = "**Table 5**: Sample output data of 
             the function `phase_stats`.")

#GroSeas (issue: returning blank table) maybe because Values are returned starting from the second year only, since `gro_start` and `gro_end` depend on the values from the previous year.


dendro_L2 <- CEN5180_L2 %>% filter(series == "CEN5180_L2")
# Calculate growing season start and end
grow_seas(dendro_L2 = CEN5180_L2)
# calculate grow_seas
grow_seas <- grow_seas(dendro_L2 = CEN5180_L2)
knitr::kable(grow_seas, caption = "**Table 4**: Sample output data of the 
             function `grow_seas`.")
```

```{r}
#load("C:/Users/gabhr/Downloads/Primet2014_2024 (1).RData")
#Primet2014_2024$ts<-as.POSIXct(Primet2014_2024$Date,format="%Y-%m-%d %H:%M:%S",tz="GMT")
#Primet2014_2024$yr<-format(Primet2014_2024$ts,"%Y")
PrimetTOMST<-Primet2014_2024
PrimetTOMSTairtemp<-PrimetTOMST|>select(c(AIRTEMP_MEAN_450_0_01,Flag_AIRTEMP_MEAN_450_0_01,ts,yr))

PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="CQQ"] <- NA
PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="Q"] <- NA
PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="QQ"] <- NA
PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="QQVQQ"] <- NA
PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="QV"] <- NA
PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="QVCQQ"] <- NA
PrimetTOMSTairtemp$AIRTEMP_MEAN_450_0_01[PrimetTOMSTairtemp$Flag_AIRTEMP_MEAN_450_0_01=="QVQQ"] <- NA
PrimetTOMSTairtemp$DOY<-yday(PrimetTOMSTairtemp$ts)
```

#Should this be max and min instead of mean?
```{r}
custom_colors <- c("2014" = "magenta","2015" = "maroon","2016" = "red", "2017" = "orange", "2018" = "yellow", "2019" = "green", "2020" = "turquoise", "2021" = "blue", "2022" = "purple", "2023" = "pink","2024"="coral")

# Calculate the 5th and 95th percentiles for each year
percentiles <- PrimetTOMSTairtemp |>
  group_by(yr) |>
  summarize(
    temp_5th = quantile(AIRTEMP_MEAN_450_0_01, 0.05,na.rm=TRUE),
    temp_95th = quantile(AIRTEMP_MEAN_450_0_01, 0.95,na.rm=TRUE)
  )

# Merge the percentiles back into the original data
PrimetTOMSTairtempPercentiles <- PrimetTOMSTairtemp |>
  left_join(percentiles, by = "yr")

# Filter the data to include only the temperatures within the range
PrimetTOMSTairtempPercentilesOnly <- PrimetTOMSTairtempPercentiles |>
  filter(AIRTEMP_MEAN_450_0_01 <= temp_5th | AIRTEMP_MEAN_450_0_01 >= temp_95th)

# Plot the histogram
library(ggplot2)
ggplot(PrimetTOMSTairtempPercentilesOnly, aes(x = DOY, y = AIRTEMP_MEAN_450_0_01, color = factor(yr))) +
  geom_point() +
  labs(title = "Highest and Lowest 5% Primet Temperatures",
       x = "Day of Year", y = "Temperature (°C)") +   scale_color_manual(values = custom_colors) + guides(color=guide_legend(override.aes = list(alpha = 1, size = 5)))+
  theme_minimal()+  geom_vline(xintercept = 176, linetype = "dashed", color = "black")

ggplot(PrimetTOMSTairtempPercentilesOnly, aes(x = AIRTEMP_MEAN_450_0_01, color = factor(yr))) +
  geom_histogram() +
  labs(title = "Highest and Lowest 5% Primet Temperature Distribution") +   scale_color_manual(values = custom_colors) + guides(color=guide_legend(override.aes = list(alpha = 1, size = 5)))+
  theme_minimal()

```

```{r}
percentiles2 <- PrimetTOMSTairtemp |>
  group_by(yr) |>
  summarize(
    temp_1 = quantile(AIRTEMP_MEAN_450_0_01, 0.01,na.rm=TRUE),
    temp_99th = quantile(AIRTEMP_MEAN_450_0_01, 0.99,na.rm=TRUE)
  )

# Merge the percentiles back into the original data
PrimetTOMSTairtempPercentiles2 <- PrimetTOMSTairtemp |>
  left_join(percentiles2, by = "yr")

# Filter the data to include only the temperatures within the range
PrimetTOMSTairtempPercentilesOnly2 <- PrimetTOMSTairtempPercentiles2 |>
  filter(AIRTEMP_MEAN_450_0_01 <= temp_1 | AIRTEMP_MEAN_450_0_01 >= temp_99th)

# Plot the histogram
library(ggplot2)
ggplot(PrimetTOMSTairtempPercentilesOnly2, aes(x = DOY, y = AIRTEMP_MEAN_450_0_01, color = factor(yr))) +
  geom_point() +
  labs(title = "Highest and Lowest 1% Primet Temperatures",
       x = "Day of Year", y = "Temperature (°C)")+   scale_color_manual(values = custom_colors) + guides(color=guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() + geom_vline(xintercept = 176, linetype = "dashed", color = "black")

ggplot(PrimetTOMSTairtempPercentilesOnly2, aes(x = AIRTEMP_MEAN_450_0_01, color = factor(yr))) +
  geom_histogram() +
  labs(title = "Highest and Lowest 1% Primet Temperature Distribution") +   scale_color_manual(values = custom_colors) + guides(color=guide_legend(override.aes = list(alpha = 1, size = 5)))+
  theme_minimal()
```



