---
title: "ecomatik"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(viridis)
library(tidyr)
library(lubridate)
library(zoo)
library(tidyverse)
library(dendRoAnalyst)
#All relative file paths have "Ecomatik" as the working directory, e.g.,:
#setwd("C:/Users/Gabby/Desktop/GJ_OSU_Thesis/GJ_OSU_Thesis/Ecomatik")
#"GJ_OSU_Thesis" is the default WD when I first open the .RProj (at least it should be)


custom_colors <- c("2014" = "magenta","2015" = "maroon","2016" = "red", "2017" = "orange", "2018" = "yellow", "2019" = "green", "2020" = "turquoise", "2021" = "blue", "2022" = "purple", "2023" = "pink","2024"="coral")
```
#reading in each year of dendrometer values in DSCMET422
```{r}
#each CSV 1st adjuTosted in Excel to remove non-header/data rows & format ts YMD
# YYYY-MM-DD HH:MM:SS ts format

dscmet_422_a_5min_2018YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2018YMD.csv")
dscmet_422_a_5min_2019YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2019YMD.csv")
dscmet_422_a_5min_2020YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2020YMD.csv")
dscmet_422_a_5min_2021YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2021YMD.csv")
dscmet_422_a_5min_2022YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2022YMD.csv")
dscmet_422_a_5min_2023YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2023YMD.csv")
dscmet_422_a_5min_2024YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2024YMD.csv")

#combining into one df
dscmet_422_all <- bind_rows(
  dscmet_422_a_5min_2018YMD,
  dscmet_422_a_5min_2019YMD,
  dscmet_422_a_5min_2020YMD,
  dscmet_422_a_5min_2021YMD,
  dscmet_422_a_5min_2022YMD,
  dscmet_422_a_5min_2023YMD,
  dscmet_422_a_5min_2024YMD
)

#This final object should already have Date in POSIXct YYYY-MM-DD HH:MM:SS foramat

dscmet_422_all$ts<-dscmet_422_all$Date
dscmet_422_all$yr<-format(dscmet_422_all$ts, "%Y")
dscmet_422_all$yr<-as.factor(dscmet_422_all$yr)
dscmet_422_all$DOY<-yday(dscmet_422_all$ts)
write.csv(file="dscmet_422_all.csv",dscmet_422_all)
save(dscmet_422_all,file="dscmet_422_all")
#can recall this with load() without bringing in the CSV file  

unique_422_strings <- unique(unlist(sapply(dscmet_422_all, function(x) x[is.character(x)])))
```

The flags in this dataframe ended up not being problematic. They were "M" "P" "E" and "MM"
From metadata, P:  "flag_valuechange(PROGID,0.1,0.1,1)="P", manual)"
E: data value estimated by interpolation

Example code if there were problematic flags (e.g., "I" for impossible):
1- Finding where in the dataframes problem flags come in so we can fix their corresponding values (if necessary) 
values_to_remove <- c("M", "I")

2- Filter data frame
filtered_Primet2014_2024_filled <- Primet2014_2024_filled[!rowSums(sapply(Primet2014_2024_filled, function(x) x %in% values_to_remove | is.na(x))) > 0, ]

3- Filtering flags after searching in R environment object where flags show up 
Primet2014_2024_sorted_filled$SWC_MEAN_0_100_04[P230long_filled$Flag_SWC_MEAN_0_100_04=="Q"] <- NA

#Plotting AVG dendro value for 309
Ecomatik website says the DC series measures "high-precision and continuous measurement of changes in stem circumference"
```{r}
dscmet_422_all$Dend309 <- dscmet_422_all$DEND309_AVG_4200_0_01 

ggplot(dscmet_422_all ,aes(x=DOY,y=Dend309,color=factor(yr)))+
  geom_point(size=1,alpha=0.5)+labs(title="HJA Douglas-fir (ID 309) Stem Circumference Changes",x="Day of Year",y="Stem Circumference Change (µm)",color="Legend") +   scale_color_manual(values = custom_colors) + guides(color=guide_legend(override.aes = list(alpha = 1, size = 5)))+theme_minimal()
```


