---
title: "ecomatik"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(viridis)
library(tidyr)
library(lubridate)
library(zoo)
library(tidyverse)
library(dendRoAnalyst)
#All relative file paths have "Ecomatik" as the working directory, e.g.,:
#setwd("C:/Users/Gabby/Desktop/GJ_OSU_Thesis/GJ_OSU_Thesis/Ecomatik")
#"GJ_OSU_Thesis" is the default WD when I first open the .RProj (at least it should be)


custom_colors <- c("2014" = "magenta","2015" = "maroon","2016" = "red", "2017" = "orange", "2018" = "yellow", "2019" = "green", "2020" = "turquoise", "2021" = "blue", "2022" = "purple", "2023" = "pink","2024"="coral")
```
#Summary of what we're working with
April 1 = DOY 91
June 26 (Heat Dome onset) = DOY 177
Oct 1 = DOY 274

Dendro summary (653 is in dscmet but not usable): 
309	Doug-fir (DT)
310	Doug-fir
311	Doug-fir
646	Western hem
648	Western hem
!!!651 (no core) western red cedar
652	Western hem
!!!654 (no core...might have actually and mislabeled as 645)
655	Doug-fir
656	Doug-fir
860	Doug-fir

309 10 and 11 are old growth PSME
860 654 655 and 656 are second growth PSME
646 648 and 652 TSHE and 651 THPL are old growth


#reading in each year of dendrometer values in DSCMET422
```{r}
#each CSV 1st adjuTosted in Excel to remove non-header/data rows & format ts YMD
# YYYY-MM-DD HH:MM:SS ts format

dscmet_422_a_5min_2018YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2018YMD.csv")
dscmet_422_a_5min_2019YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2019YMD.csv")
dscmet_422_a_5min_2020YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2020YMD.csv")
dscmet_422_a_5min_2021YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2021YMD.csv")
dscmet_422_a_5min_2022YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2022YMD.csv")
dscmet_422_a_5min_2023YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2023YMD.csv")
dscmet_422_a_5min_2024YMD <- read_csv("raw_data_DSCMET422/dscmet_422_a_5min_2024YMD.csv")

#combining into one df
dscmet_422_all <- bind_rows(
  dscmet_422_a_5min_2018YMD,
  dscmet_422_a_5min_2019YMD,
  dscmet_422_a_5min_2020YMD,
  dscmet_422_a_5min_2021YMD,
  dscmet_422_a_5min_2022YMD,
  dscmet_422_a_5min_2023YMD,
  dscmet_422_a_5min_2024YMD
)

#This final object should already have Date in POSIXct YYYY-MM-DD HH:MM:SS foramat

dscmet_422_all$ts<-dscmet_422_all$Date
dscmet_422_all$yr<-format(dscmet_422_all$ts, "%Y")
dscmet_422_all$yr<-as.factor(dscmet_422_all$yr)
dscmet_422_all$DOY<-yday(dscmet_422_all$ts)
#write.csv(file="dscmet_422_all.csv",dscmet_422_all)
save(dscmet_422_all,file=here("DendroAnalyst/dscmet_422_all.RData"))
#can recall this with load(dscmet_422_all.Rdata) without bringing in the CSV file  

unique_422_strings <- unique(unlist(sapply(dscmet_422_all, function(x) x[is.character(x)])))
```

The flags in this dataframe ended up not being problematic. They were "M" "P" "E" and "MM"
From metadata, P:  "flag_valuechange(PROGID,0.1,0.1,1)="P", manual)"
E: data value estimated by interpolation

Example code if there were problematic flags (e.g., "I" for impossible):
1- Finding where in the dataframes problem flags come in so we can fix their corresponding values (if necessary) 
values_to_remove <- c("M", "I")

2- Filter data frame
filtered_Primet2014_2024_filled <- Primet2014_2024_filled[!rowSums(sapply(Primet2014_2024_filled, function(x) x %in% values_to_remove | is.na(x))) > 0, ]

3- Filtering flags after searching in R environment object where flags show up 
Primet2014_2024_sorted_filled$SWC_MEAN_0_100_04[P230long_filled$Flag_SWC_MEAN_0_100_04=="Q"] <- NA

#309 Plotting raw AVG dendro value for 309
Ecomatik website says the DC series measures "high-precision and continuous measurement of changes in stem circumference"
April 1 = DOY 91
June 26 (Heat Dome onset) = DOY 177
Oct 1 = DOY 274
```{r}
dscmet_422_all$Dend309 <- dscmet_422_all$DEND309_AVG_4200_0_01 

#All days
raw309all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend309,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 309) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw309all
ggsave("raw309all.png", plot = raw309all)


#Apr1-Oct1
raw309gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend309,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 309) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(-2000,17500)
raw309gro
ggsave("raw309gro.png", plot = raw309gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#310 Plotting raw AVG dendro value for 310
```{r}
dscmet_422_all$Dend310 <- dscmet_422_all$DEND310_AVG_150_0_01 

#All days
raw310all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend310,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 310) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw310all
ggsave("raw310all.png", plot = raw310all)


#Apr1-Oct1
raw310gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend310,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 310) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)
raw310gro
ggsave("raw310gro.png", plot = raw310gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#311 Plotting raw AVG dendro value for 311
```{r}
dscmet_422_all$Dend311 <- dscmet_422_all$DEND311_AVG_150_0_01 

#All days
raw311all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend311,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 311) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw311all
ggsave("raw311all.png", plot = raw311all)


#Apr1-Oct1
raw311gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend311,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 311) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)#+ylim(-2000,17500)
raw311gro
ggsave("raw311gro.png", plot = raw311gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#646 Plotting raw AVG dendro value for 646
```{r}
dscmet_422_all$Dend646 <- dscmet_422_all$DEND646_AVG_150_0_01 
wide646<-dscmet_422_all|>
  select(c(Dend646,yr,ts,DOY))|>
  pivot_wider(names_from = yr, values_from = Dend646)
#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide646gro would normalize to DOY 91

#recombine both years data frames into one. 
wide646gro<-wide646
growth2018<-wide646gro|>select(c(ts,DOY,'2018'))
growth2019<-wide646gro|>select(c(ts,DOY,'2019'))
growth2020<-wide646gro|>select(c(ts,DOY,'2020'))
growth2021<-wide646gro|>select(c(ts,DOY,'2021'))
growth2022<-wide646gro|>select(c(ts,DOY,'2022'))
growth2023<-wide646gro|>select(c(ts,DOY,'2023'))
growth2024<-wide646gro|>select(c(ts,DOY,'2024'))

growth2018$`2018`<-growth2018$`2018` 
growth2019$`2019`<-growth2019$`2019`- 3596
growth2020$`2020`<-growth2020$`2020`- 8364
growth2021$`2021`<-growth2021$`2021`- 14003
growth2022$`2022`<-growth2022$`2022` - 5760
growth2023$`2023`<-growth2023$`2023`- 7088
growth2024$`2024`<-growth2024$`2024`- 11400

wide646gro<-bind_rows(growth2018,growth2019,growth2020,growth2021,growth2022,growth2023,growth2024)
long646grow<- wide646gro |>   pivot_longer(
    cols = starts_with("20"),
    names_to = "yr",  
    values_to = "annual_growth")  


ggplot(long646grow, aes(x=DOY,y=annual_growth,color=factor(yr))) + geom_point(
  size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 646) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(-1000,6000)
```


#All days
raw646all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend646,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 646) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw646all
ggsave("raw646all.png", plot = raw646all)


#Apr1-Oct1
raw646gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend646,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 646) Stem Circumference Changes: A1-O1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(0,21000)
raw646gro
ggsave("raw646gro.png", plot = raw646gro)

#normalized A1-O1


norm646gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend646,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 646) Stem Circumference Changes: A1-O1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(0,21000)
norm646gro
ggsave("norm646gro.png", plot = norm646gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```


#648 Plotting raw AVG dendro value for 648
```{r}
dscmet_422_all$Dend648 <- dscmet_422_all$DEND648_AVG_150_0_01 

#All days
raw648all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend648,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 648) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw648all
ggsave("raw648all.png", plot = raw648all)


#Apr1-Oct1
raw648gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend648,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 648) Stem Circumference Changes: A1-O1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(4000,12500)
raw648gro
ggsave("raw648gro.png", plot = raw648gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#651 Plotting raw AVG dendro value for 651
```{r}
dscmet_422_all$Dend651 <- dscmet_422_all$DEND651_AVG_150_0_01 

#All days
raw651all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend651,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western red cedar (ID 651) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw651all
ggsave("raw651all.png", plot = raw651all)


#Apr1-Oct1
raw651gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend651,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western red cedar (ID 651) Stem Circumference Changes: A1-O1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(0,12500)
raw651gro
ggsave("raw651gro.png", plot = raw651gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")


```

#652 Plotting raw AVG dendro value for 652
```{r}
dscmet_422_all$Dend652 <- dscmet_422_all$DEND652_AVG_150_0_01 

#All days
raw652all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend652,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 652) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )
raw652all
ggsave("raw652all.png", plot = raw652all)


#Apr1-Oct1
raw652gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend652,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA western hemlock (ID 652) Stem Circumference Changes: A1-O1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(5000,14000)
raw652gro
ggsave("raw652gro.png", plot = raw652gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#654 Plotting raw AVG dendro value for 654
```{r}
dscmet_422_all$Dend654 <- dscmet_422_all$DEND654_AVG_150_0_01 

#All days
raw654all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend654,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 654) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+ylim(0,17500)
#ylim here because early 2018 has garbage numbers that are big outliers
raw654all
ggsave("raw654all.png", plot = raw654all)


#Apr1-Oct1
raw654gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend654,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 654) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(0,17500)
raw654gro
ggsave("raw654gro.png", plot = raw654gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#655 Plotting raw AVG dendro value for 655
```{r}
dscmet_422_all$Dend655 <- dscmet_422_all$DEND655_AVG_150_0_01 

#All days
raw655all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend655,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 655) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+ylim(0,22500)
raw655all
ggsave("raw655all.png", plot = raw655all)


#Apr1-Oct1
raw655gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend655,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 655) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(1500,22500)
raw655gro
ggsave("raw655gro.png", plot = raw655gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#656 Plotting raw AVG dendro value for 656
```{r}
dscmet_422_all$Dend656 <- dscmet_422_all$DEND656_AVG_150_0_01 

#All days
raw656all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend656,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 656) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+ylim(0,20000)
raw656all
ggsave("raw656all.png", plot = raw656all)


#Apr1-Oct1
raw656gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend656,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 656) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(0,20000)
raw656gro
ggsave("raw656gro.png", plot = raw656gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```

#860 Plotting raw AVG dendro value for 860
```{r}
dend860<-dscmet_422_all$Dend860
dscmet_422_all$Dend860 <- dscmet_422_all$DEND860_AVG_150_0_01 

#All days
raw860all <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend860,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 860) Stem Circumference Changes",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+ylim(0,20000)
raw860all
ggsave("raw860all.png", plot = raw860all)


#Apr1-Oct1
raw860gro <- ggplot(
  dscmet_422_all,aes(x=DOY,y=Dend860,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="HJA Douglas-fir (ID 860) Stem Circumference Changes: April 1 - Oct 1",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal(
                )+xlim(91,274)+ylim(0,20000)
raw860gro
ggsave("raw860gro.png", plot = raw860gro)

#+annotate("text",x=170,y=-2000,label="Onset of Heat Dome",color="black")
```























