---
title: "dendroanalyst"
format: pdf
editor: visual
---

# Code to run at launch

```{r}
#install.packages("dendRoAnalyst")
library(ggplot2)
library(dplyr)
library(readr)
library(viridis)
library(tidyr)
library(lubridate)
library(here)
library(zoo)
library(tidyverse)
library(dendRoAnalyst)
custom_colors <- c("2014" = "magenta","2015" = "maroon","2016" = "red", "2017" = "orange", "2018" = "yellow", "2019" = "green", "2020" = "lightblue", "2021" = "blue", "2022" = "purple", "2023" = "pink","2024"="darkgreen")

#keep getting errors with logical and character vectors mixing with remarks= so let's patch the code locally with getAnywhere(daily.data)
daily.data.patched<-function (df, TreeNum) 
{
    if (TreeNum > ncol(df) - 1) {
        stop("The tree number is not valid for provided dataset.")
    }
    TIME <- DATE <- dm <- time <- Max <- Min <- Time_max <- Time_min <- NULL
    tn <- TreeNum + 1
    if (!inherits(df[[1]], "Date") && !inherits(df[[1]], "POSIXct")) {
        df[[1]] <- ymd_hms(df[[1]])
    }
    df <- tibble(df) %>% select(c(1, all_of(tn)))
    df <- df %>% rename(TIME = 1, dm = 2) %>% mutate(DATE = as.Date(TIME), 
        time = format(TIME, "%H:%M:%S"))
    result <- df %>% group_by(DATE) %>% summarize(Min = min(dm, 
        na.rm = T), Time_min = time[which.min(dm)], Max = max(dm, 
        na.rm = T), Time_max = time[which.max(dm)], mean = mean(dm, 
        na.rm = T), median = median(dm, na.rm = T), amplitude = Max - 
        Min, Remarks = ifelse(isTRUE(Time_max > Time_min), "*", ""))
    return(result)
}

#From example DA code in vignette:
read.den
 pkgname <- "dendRoAnalyst"
 source(file.path(R.home("share"), "R", "examples-header.R"))
 options(warn = 1)
 options(pager = "console")
 library('dendRoAnalyst')
 base::assign(".oldSearch", base::search(), pos = 'CheckExEnv')
 base::assign(".old_wd", base::getwd(), pos = 'CheckExEnv')
 flush(stderr()); flush(stdout())
```

# Function Information

Summaries in the right column in bold indicate extra relevant functions. Example code for all these are in separate code chunks in their corresponding headings which you can see in the outline to the right.

+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [clim.twd](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/clim.twd.html)                                                                                                                                                                                                                                                                                             | **Calculating relative growth change during no-rain periods. Requires dendrometer (e.g., nepa17) and rainfall (e.g., rain17) data.**       |
|                                                                                                                                                                                                                                                                                                                                                                              |                                                                                                                                            |
| Description from ?clim.twd:                                                                                                                                                                                                                                                                                                                                                  | E.g.,\                                                                                                                                     |
|                                                                                                                                                                                                                                                                                                                                                                              | relative_dry_growth\<-clim.twd(df=gf_nepa17, Clim=ktm_rain17, dailyValue='max', showPlot=TRUE) \[1\]\                                      |
| This function calculates the number and the location of climatically adverse periods within a climate time series. The user can define a duration and threshold of these conditions. The function also provides the relative radial/circumferencial change during each adverse period for the original or normalized data. See Raffelsbauer et al., (2019) for more details. | Output: "The adverse climatic periods are listed below:" start_doy, end_doy, Id,                                                           |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [daily.data](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/daily.data.html)                                                                                                                                                                                                                                                                                         | Calculation of **daily statistics for dendrometer data**                                                                                   |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dendro.resample](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dendro.resample.html)                                                                                                                                                                                                                                                                               | Resampling temporal resolution of dendrometer and climate data                                                                             |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dendro.truncate](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dendro.truncate.html)                                                                                                                                                                                                                                                                               | Truncation of the dendrometer data                                                                                                         |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.detrend](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.detrend.html)                                                                                                                                                                                                                                                                                         | Detrend the dendrometer data (**should we?**)                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.fit.gompertz](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.fit.gompertz.html)                                                                                                                                                                                                                                                                               | Fitting gompertz function on annual dendrometer data                                                                                       |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.na.interpolation](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.na.interpolation.html)                                                                                                                                                                                                                                                                       | Detection and **interpolation of missing values** in dendrometer data.                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [gf_nepa17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/gf_nepa17.html)                                                                                                                                                                                                                                                                                           | Dendrometer data of Kathmandu for 2017 with gap filled                                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [i.jump.locator](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/i.jump.locator.html)                                                                                                                                                                                                                                                                                 | Removing artefacts due to manual adjustments of dendrometers interactively                                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [jump.locator](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/jump.locator.html)                                                                                                                                                                                                                                                                                     | **Removing artefacts due to manual adjustments of dendrometers automatically for more than one dendrometers**                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [ktm_rain17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/ktm_rain17.html)                                                                                                                                                                                                                                                                                         | Daily rainfall data of Kathmandu for 2017.                                                                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [mean_detrended.dm](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/mean_detrended.dm.html)                                                                                                                                                                                                                                                                           | Calculate mean of detrended dendrometer data.                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [mov.cor.dm](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/mov.cor.dm.html)                                                                                                                                                                                                                                                                                         | **Calculating running correlation between dendrometer data and daily climate.**                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [nepa](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa.html)                                                                                                                                                                                                                                                                                                     | Dendrometer data from Kathmandu                                                                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [nepa17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa17.html)                                                                                                                                                                                                                                                                                                 | Dendrometer data of Kathmandu for 2017                                                                                                     |
|                                                                                                                                                                                                                                                                                                                                                                              |                                                                                                                                            |
|                                                                                                                                                                                                                                                                                                                                                                              | from the paper: "The data was derived using an automatic logging band dendrometer (DRL26, EMS Brno) recording hourly circumference change" |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [pa2](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa2.html)                                                                                                                                                                                                                                                                                                     | Dendrometer data from Kathmandu version 2                                                                                                  |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [network.interpolation](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/network.interpolation.html)                                                                                                                                                                                                                                                                   | Interpolation of NA values using the dendrometer network                                                                                   |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [phase.sc](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/phase.sc.html)                                                                                                                                                                                                                                                                                             | **Application of the stem-cycle approach to calculate different phases, their duration and to plot them.**                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [phase.zg](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/phase.zg.html)                                                                                                                                                                                                                                                                                             | **Application of the zero-growth approach to calculate different phases, their duration and to plot them.**                                |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_mov.cor](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_mov.cor.html)                                                                                                                                                                                                                                                                                     | **Plotting moving correlation with climate.**                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_SC_output](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_SC_output.html)                                                                                                                                                                                                                                                                                 | **Plotting output of stem cycle approach (i.e., output of phase.sc function).**                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_ZG_output](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_ZG_output.html)                                                                                                                                                                                                                                                                                 | **Plotting output of zero-growth approach (i.e., output of phase.zg function).**                                                           |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [read.dendrometer](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/read.dendrometer.html)                                                                                                                                                                                                                                                                             | **Reading dendrometer data and automatically converts datetime format to what is needed.**                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [twd.maxima](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/twd.maxima.html)                                                                                                                                                                                                                                                                                         | **Locating the maxima of TWD periods**                                                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+

-   The ‘phase.sc’ and ‘phase.zg’ only calculate the phases and phase statistics.

-   There are two separate functions added to plot the output of ‘phase.sc’ and ‘phase.zg’.

-   A new function, ‘dm.detrend’ has been added to detrend the dendrometer data using ‘gam’, ‘first difference’ and ‘gompertz’ function.

-   A new function ‘dm.fit.gompertz’ is introduced to fit Gompertz function. It also provide option to calculate the growth rate applying first derivative of the Gompertz function.

-   The function ‘spline.interpolation’ is now ‘dm.na.interpolation’, with two methods: ‘spline’ and ‘seasonal’ interpolation. The seasonal interpolation used ARIMA model to detect the daily pattern for three weeks (the week of gap occurrence + one week earlier and later) and interpolates the missing values.

-   A new function ‘read.dendrometer’ is included to read the dendrometer data from TXT, CSV and EXCEL files.

-   A new function ‘mean.detrended.dm’ calculates the averaged detrended dendrometer data with possibility getting first order autocorrelation free averaged series.

-   A new function ‘mov.cor.dm’ is added for calculating climate-growth relationships.

# Loading in and prepping data

Example data for DendRoAnalyst has hourly dendrometer data and daily rainfall data, so we will be condensing our data down to match.

```{r}
#raw dendrometer data
load(here(file="DendroAnalyst/dscmet_422_all.RData")) #calling saved R object from ecomatik.rmd
#checking bad flags - no bad ones found so this code won't be rerun 
#unique_422_strings <- unique(
#  unlist(
#    sapply(
#      dscmet_422_all, function(x) x[is.character(x)])))

#getting hourly averages - starting with adding an hour column
dscmet_422_all$hour <- format(
  dscmet_422_all$Date, "%Y-%m-%d %H:00:00")
dscmet_422_all$Date<-as.POSIXct(dscmet_422_all$hour,format="%Y-%m-%d %H")
#getting daily averages - starting with adding a day column
dscmet_422_all$day <- format(
  dscmet_422_all$Date, "%Y-%m-%d")
dscmet_422_all$day<-as.POSIXct(dscmet_422_all$day,format="%Y-%m-%d")

#309 - OG DF - need to check if these values are accurate
d309_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND309_AVG_4200_0_01,na.rm=TRUE))|>
  filter(!is.nan(value))
colnames(d309_hourly)[colnames(d309_hourly) == 'hour'] <- 'Date'

d309_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(dailygrowth = mean(DEND309_AVG_4200_0_01,na.rm=TRUE))
colnames(d309_daily)[colnames(d309_daily) == 'day'] <- 'Date'

#310 - OG DF
d310_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg310 = mean(DEND310_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg310))  

#311 - OG DF
d311_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg311 = mean(DEND311_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg311)) 

#654 - Secondary 

646	Western hem
648	Western hem
!!!651 (no core) western red cedar
652	Western hem
!!!654 (no core...might have actually and mislabeled as 645)
655	Doug-fir
656	Doug-fir
860	Doug-fir

860 654 655 and 656 are second growth PSME
646 648 and 652 TSHE
#


#rainfall data (from climate comparisons updated.rmd)
load(file=here("P230.RData"))

#checking for bad flags
P230_rain<-P230_df|>select(c(Date,PRECIP_TOT_100_0_01,Flag_PRECIP_TOT_100_0_01))
unique_rain_strings <- unique(
  unlist(
    sapply(
      P230_rain, function(x) x[is.character(x)])))
values_to_remove <- c("Q", "QQ")

# Filter data frame to swap back values with NA
filtered_P230_rain <- P230_rain[!rowSums(sapply(P230_rain, function(x) x %in% values_to_remove | is.na(x))) > 0, ]
#checking that it worked 
unique_filtered_rain_strings <- unique(
  unlist(
    sapply(
      filtered_P230_rain, function(x) x[is.character(x)])))

#daily rain totals
filtered_P230_rain$day <- format(
  filtered_P230_rain$Date, "%Y-%m-%d")
filtered_P230_rain_daily <- filtered_P230_rain |> 
  group_by(day) |> 
  summarise(value = sum(PRECIP_TOT_100_0_01))
filtered_P230_rain_daily$IDs<-"precip_sum"
colnames(filtered_P230_rain_daily)[colnames(filtered_P230_rain_daily) == 'day'] <- 'Date'

#hourly totals 
filtered_P230_rain$hour <- format(
  filtered_P230_rain$Date, "%Y-%m-%d %H:00:00")
filtered_P230_rain_hourly <- filtered_P230_rain |> 
  group_by(hour) |> 
  summarise(precip_sum = sum(PRECIP_TOT_100_0_01,na.rm = TRUE))
filtered_P230_rain_hourly$time<-as.POSIXct(filtered_P230_rain_hourly$hour,format="%Y-%m-%d %H:00:00")
filtered_P230_rain_hourly<-filtered_P230_rain_hourly|>select(c(time,precip_sum))
filtered_P230_rain_hourly$IDs<-"precip_sum"
```

# Adjusting temporal resolution via dendro.resample

```         
dendro.resample(df, by, value) 
```

### Explanation of options

+----------------------------------+--------------------------------------------------------------------------------------------+
| `df`                             | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`.  |
+----------------------------------+--------------------------------------------------------------------------------------------+
| `by`                             | either *H, D, W* or *M* to resample data into hourly, daily, weekly or monthly resolution. |
+----------------------------------+--------------------------------------------------------------------------------------------+
| `value`                          | either *max, min*, *mean* or *sum* for the resampling value.                               |
+----------------------------------+--------------------------------------------------------------------------------------------+

```{r}
# To resample monthly with maximum value
resample_M<-dendro.resample(df=nepa17[,1:2], by='M', value='max')
head(resample_M,10)
#example output
#         Time       T2
#1  2017-01-01 63.90855
#2  2017-02-01 64.88705
```

# Extracting specific values with dendro.truncate

```         
dendro.truncate(df, CalYear, DOY) 
```

### Explanation of options

+----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                             | dataframe with the first column named date and time in the format `yyyy-mm-dd HH:MM:SS`.                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
+----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `CalYear`                        | desired year of calculation.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
+----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `DOY`                            | If we provide an array instead of a single value for `CalYear` and a single value for `DOY`, it truncates data from the `DOY` of the first `CalYear` to the same `DOY` of the second `CalYear`. Conversely, if we provide one value for `CalYear` and an array of two elements for `DOY` truncates the data form first `DOY` to second `DOY` within the same `CalYear`. Finally, if we provide an array with two values for both `DOY` and `CalYear`, it truncates data from the first `DOY` of the first `CalYear` to the second `DOY` of second `CalYear`. |
+----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#Example extracting data from doy 20 to 50 in 2017.
#trunc1<-dendro.truncate(df=nepa, CalYear=2017, DOY=c(20,50))
#head(trunc1,10)
#                 Time       T2       T3
#1  2017-01-20 00:00:00 64.92424 48.87772
#...10 2017-01-20 09:00:00 64.96733 48.88450
```

# Interpolating NA values using dendrometer network via network.interpolation

```         
network.interpolation(df, referenceDF, niMethod) 
```

### Explanation of options

+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                             | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and dendrometer data in the second column and onward. The data gaps must be filled with `NA` using the gap.interpolation function. |
+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `referenceDF`                    | dataframe with other dendrometers to be used as reference for the interpolation. The more dendrometers are included, the more robust will be the interpolation.                                                             |
+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `niMethod`                       | string, either *'linear'* or *'proportional'* for interpolation method.                                                                                                                                                     |
+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#Example from vignette
#df1<-gf_nepa17
#> # Creating an artificial reference dataset.
#> df2<-cbind(gf_nepa17,gf_nepa17[,2:3],gf_nepa17[,2:3])
#> # Creating gaps in dataset by replacing some of the reading with NA in dataset.
#> df1[40:50,3]<-NA
#> # Using proportional interpolation method.
#> df1_NI<-network.interpolation(df=df1, referenceDF=df2, niMethod='proportional')
#Warning in norm.inter(t, adj.alpha) :
#  extreme order statistics used as endpoints
#[1] "Interpolation in T3"
#         Time Interpolated_value    LCI95    UCI95
#1  1483369200           48.31525 48.30600 48.32141
#2  1483372800           48.31522 48.31331 48.31585
#> head(df1_NI,10)
#                  Time       T2       T3
#1  2017-01-01 00:00:00 62.26459 48.16801
#...10 2017-01-01 09:00:00 62.46552 48.21498
```

# Calculating difference phase using two approaches

```         
phase.sc(df, TreeNum, smoothing = NULL) 
```

### Explanation of options for stem cycle approach

three phases in this approach: 1) Shrinkage, when the dendrometer reading is less than previous reading, 2) Expansion, when current reading is more than previous reading and 3) Increment, when current reading is higher than the previous maximum. Output can be plotted with code like this:

**plot_SC_output( SC_output, DOY, Year, cols = c("#fee8c8", "#fdbb84", "#e34a33"), phNames = c("Shrinkage", "Expansion", "Increment") )**

+----------------------------------+----------------------------------------------------------------------------------------------------------------+
| `DOY`                            | array with initial and final day for plotting. E.g. *c(a,b)*, where a = initial date and b = final date.       |
+----------------------------------+----------------------------------------------------------------------------------------------------------------+
| `cols`                           | array of three elements: colour for each phases.                                                               |
+----------------------------------+----------------------------------------------------------------------------------------------------------------+
| `phNames`                        | array with three elements for three different phases. Default is **"Shrinkage", "Expansion" and "Increment"**. |
+----------------------------------+----------------------------------------------------------------------------------------------------------------+

### Inputs for phase.sc:

+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                             | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`. It should contain data with constant temporal resolution for best results.                                                                                                                    |
+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`                        | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                                                                                                                                       |
+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `smoothing`                      | numerical value from 1 to 12 which indicates the length of the smoothing spline, i.e. 1 = 1 hour and 12 = 12 hours. Default is `NULL` for no smoothing.The function [`smooth.Pspline`](http://127.0.0.1:17265/help/library/pspline/help/smooth.Pspline) is used for smoothing the data. |
+----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Outputs a list of two dataframes. The first dataframe *SC_cycle* with cyclic phases along with various statistics and the second dataframe *SC_phase* with assigned phases for each data point.The dataframe *SC_cycle* contains the beginning, end, duration, magnitude and rate of each phase. The dataframe *SC_phase* contains time and corresponding phases during that time.

```         
phase.zg(df, TreeNum) 
```

### Explanation of options for zero-growth approach

+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                             | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`. It should contain data with constant temporal resolution for best results. |
+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`                        | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                    |
+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Outputs a list of two dataframes. The first dataframe *ZG_cycle* contains the cyclic phases along with various statistics and the second dataframe *ZG_phase* with assigned phases for each data point.

Output can be plotted with th following code:

```         
plot_ZG_output(ZG_output, DOY, Year) 
```

+----------------------------------+----------------------------------------------------------------------------------------------------------+
| `ZG_output`                      | list the output of `phase.zg` function.                                                                  |
+----------------------------------+----------------------------------------------------------------------------------------------------------+
| `DOY`                            | array with initial and final day for plotting. E.g. *c(a,b)*, where a = initial date and b = final date. |
+----------------------------------+----------------------------------------------------------------------------------------------------------+
| `Year`                           | array for indicating year for plotting.                                                                  |
+----------------------------------+----------------------------------------------------------------------------------------------------------+

Two subplots with GRO and TWD respectively.

```{r}
#SC Example
sc.phase<-phase.sc(df=d309_hourly, TreeNum=1, smoothing=12, outputplot=TRUE, days=c(150,160))

#ZG Example
zg.phase<-phase.zg(df=gf_nepa17, TreeNum=1, outputplot=TRUE, days=c(150,160))
```

```{r}
#raw
sc_d309<-phase.sc(df=d309_hourly,TreeNum = 1,smoothing = 1)
sc_d309_plot<-plot_SC_output(sc_d309,DOY = c(91,274),Year = 2021,cols = c("red","green","blue"))

#jump-free
sc_d309_jf<-phase.sc(df=jump_free_d309_noNA,TreeNum = 1,smoothing = 1)
sc_d309_jf_plot<-plot_SC_output(sc_d309_jf,DOY = c(91,274),Year = 2021,cols = c("red","green","blue"))
gridExtra::grid.arrange(sc_d309_plot,sc_d309_jf_plot)

#zero-growth approach: looks very strange
zg_309<-phase.zg(df=d309_hourly,TreeNum = 1)
plot_ZG_output(zg_309,DOY = c(91,274),Year = 2020)
```

# Interpolating Missing Values via dm.na.interpolation

```         
dm.na.interpolation(df, resolution, fill = FALSE, method = "spline") 
```

### Explanation of options

+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                             | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and following columns with dendrometer data for the same temporal resolution and time period. |
+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `resolution`                     | integer, indicating the resolution of dendrometer data in **minutes**.                                                                                                                 |
+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `fill`                           | logical, if `TRUE` it fills the `NA` values using spline interpolation. Default is `FALSE`.                                                                                            |
+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `method`                         | string, 'spline' for the spline interpolation or 'seasonal' for the seasonal interpolation                                                                                             |
+----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
sc_d309<-
```

Locating maxima TWD Periods with twd.maxima

```{r}
# library(dendRoAnalyst)
#> data(gf_nepa17)
#> df1=gf_nepa17[2500:3500,]
#> twd_max<-twd.maxima(df=df1, TreeNum=2, showPlot=FALSE)
#> head(twd_max,10)
#             start.time            end.time                Time         TWD
#20  2017-04-15 11:00:00 2017-04-16 04:00:00 2017-04-15 22:00:00 0.028245289
#43  2017-04-16 08:00:00 2017-04-18 06:00:00 2017-04-16 21:00:00 0.070221431
#    duration_from_start twd.number
#20                  540          1
#43                  660          2
```

# Plotting amplitude (max-min) from daily.data.patched

See "Code to run at launch" section for details on daily.data vs daily.data.patched

```         
daily.data(df, TreeNum) 
```

This function calculates various statistics of dendrometer data on a daily basis. The daily statistics includes the daily maximum and minimum with their corresponding times and daily amplitude (difference between daily maximum and minimum). See King et al. (2013) for details. Output is a dataframe with the daily statistics of the dendrometer data and Remarks = "\*" if Time_max \> Time_min otherwise "".

```{r}
# 309
daily_stats309<-daily.data.patched(df=d309_hourly, TreeNum=1)
daily_stats309$yr <- format(
  daily_stats309$DATE, "%Y")
daily_stats309$yr<-as.factor(daily_stats309$yr)
daily_stats309$DOY<-yday(daily_stats309$DATE)

raw309amp <- ggplot(
  daily_stats309,aes(x=DOY,y=amplitude,color=factor(yr)))+geom_point()+labs(
      title="HJA Douglas-fir (ID 309) Daily Net Circumference Change",
      x="Day of Year",y="Net Daily Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal()+ylim(0,900)+xlim(91,274)
raw309amp



# 310
daily_stats310<-daily.data.patched(df=d310_hourly, TreeNum=1)
daily_stats310$yr <- format(
  daily_stats310$DATE, "%Y")
daily_stats310$yr<-as.factor(daily_stats310$yr)
daily_stats310$DOY<-yday(daily_stats310$DATE)

raw310amp <- ggplot(
  daily_stats310,aes(x=DOY,y=amplitude,color=factor(yr)))+geom_point()+labs(
      title="HJA Douglas-fir (ID 310) Daily Net Circumference Change",
      x="Day of Year",y="Net Daily Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal()+ylim(0,550)+xlim(91,274)
raw310amp

# 311
daily_stats311<-daily.data.patched(df=d311_hourly, TreeNum=1)
daily_stats311$yr <- format(
  daily_stats311$DATE, "%Y")
daily_stats311$yr<-as.factor(daily_stats311$yr)
daily_stats311$DOY<-yday(daily_stats311$DATE)

raw311amp <- ggplot(
  daily_stats311,aes(x=DOY,y=amplitude,color=factor(yr)))+geom_point()+labs(
      title="HJA Douglas-fir (ID 311) Daily Net Circumference Change",
      x="Day of Year",y="Net Daily Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal()+ylim(0,625)+xlim(91,274)
raw311amp
```

# 1Relative Dry Growth During Climate Adverse events with clim.twd

```         
clim.twd(df,   Clim,   dailyValue = "max",   thresholdClim = "<10",   thresholdDays = ">5",   showPlot = TRUE ) 
```

```{r}
library(dendRoAnalyst)
data(gf_nepa17)
data(ktm_rain17)
relative_dry_growth<-clim.twd(df=gf_nepa17, Clim=ktm_rain17, dailyValue='max', showPlot=TRUE)
1

```

### Explanation of options

+----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `Clim`                           | dataframe with the first column containing `Date in yyyy-mm-dd` and second column containing corresponding climate data.                                                                                                                                |
+----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `dailyValue`                     | either *'max', 'min', 'mean'*, or *'sum'* for selecting the daily resampled value. Default is *'max'*. See [`dendro.resample`](http://127.0.0.1:17265/help/library/dendRoAnalyst/help/dendro.resample) for details.                                     |
+----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `thresholdClim`                  | string, the theshold for the respective climatic parameter. E.g. if climatic data is precipitation then days, where precipitation is below or equal to this value, are considered as adverse climate. Dafault is '\<10'.                                |
+----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `thresholdDays`                  | string, the minimum number of consecutive adverse days to be considered for analysis. For example, `thresholdDays`=2 means the relative radial/circumferential change is calculated for adverse periods lasting for more than 2 days. Default is '\>5'. |
+----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#raw309
 library(dendRoAnalyst)
relative_dry_growth309<-clim.twd(
  df=d309_daily, 
  Clim=filtered_P230_rain_daily, 
  dailyValue='mean', 
  thresholdClim = "<2.5",
  thresholdDays = ">10",
  showPlot=TRUE)

#jump-free309
relative_dry_growth_jf309<-clim.twd(
  df=jump_free_d309, 
  Clim=filtered_P230_rain_daily, 
  dailyValue='mean', 
  thresholdClim = "<2.5",
  thresholdDays = ">10",
  showPlot=TRUE)
```

# Removing artefacts from manual adjustments with jump.locator

```         
i.jump.locator(df, TreeNum, v) 
```

### Explanation of options for similar function i.jump.locator (only difference is that it adds TreeNum whereas jump.locator doesn't)

output is a new df that locates and adjusts the artefacts

+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                             | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and the dendrometer data in following columns.                                                        |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`                        | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                                              |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `v`                              | numerical value which is considered as artefact. E.g. `v`=1 implies that if the difference to the consecutive data point is more than 1 or less than -1, it will be considered as an artefact. |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

#### Examples from vignette:

```{r}
jump_free_nepa<-jump.locator(df=nepa,v=1)
```

#### Using our own data:

First look at the raw data outputs and see how the jumps are to help set the "v=" value. When you run the function it makes a table with the number of rows equal to the number of removed values and the columns equal to observation number, timestamp of the jump, and the value of the jump. If you name the jump locator object, that will add an object to your environment of the adjusted df with the jump values removed.

```{r}
jump_free_d309<-
jump.locator(df=d309_hourly,v=200) 
#this takes out 41 values and makes a table with the jump value & time and makes a df with the removed values 

jump_free_d310<-jump.locator(df=d310_hourly,v=200)

jump_free_d311<-jump.locator(df=d311_hourly,v=200)

```

#### Plotting new data to see the difference:

Plots with all dates on the x axis

```{r}
vline_HD <- as.POSIXct("2021-06-25 00:00")
vline_oct1<-as.POSIXct(c("2018-10-01 00:00","2019-10-01 00:00","2020-10-01 00:00","2021-10-01 00:00","2022-10-01 00:00","2023-10-01 00:00","2024-10-01 00:00"))
vline_apr1<-as.POSIXct(c("2018-04-01 00:00","2019-04-01 00:00","2020-04-01 00:00","2021-04-01 00:00","2022-04-01 00:00","2023-04-01 00:00","2024-04-01 00:00"))

ggplot(
  jump_free_d309,aes(x=TIME,y=value))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 309) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+  geom_vline(xintercept = as.numeric(vline_oct1), linetype = "dashed", color = "green")+  geom_vline(xintercept = as.numeric(vline_apr1), linetype = "dashed", color = "purple")

ggplot(
  jump_free_d310,aes(x=TIME,y=hourly_avg310))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 310) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+  geom_vline(xintercept = as.numeric(vline_oct1), linetype = "dashed", color = "green")+  geom_vline(xintercept = as.numeric(vline_apr1), linetype = "dashed", color = "purple")

ggplot(
  jump_free_d311,aes(x=TIME,y=hourly_avg311))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 311) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+  geom_vline(xintercept = as.numeric(vline_oct1), linetype = "dashed", color = "green")+  geom_vline(xintercept = as.numeric(vline_apr1), linetype = "dashed", color = "purple")
```

Raw plots were made in "ecomatik.rmd"

```{r}
#Apr1-Oct1 D309
jump_free_d309$yr<-format(jump_free_d309$TIME, "%Y")
jump_free_d309$yr<-as.factor(jump_free_d309$yr)
jump_free_d309$DOY<-yday(jump_free_d309$TIME)
ggplot(
  jump_free_d309,aes(x=DOY,y=value,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 309) Adjusted Stem Circumference Change",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal()+ylim(-2000,25500)#+xlim(91,274)
```

# Modeling annual growth with a gompertz function

```         
dm.fit.gompertz(df, CalYear, TreeNum, f_derivative = F) 
gomp_fitted<-dm.fit.gompertz(df=gf_nepa17, TreeNum = 1, CalYear=2017)
```

### Explanation of options

+----------------------------------+------------------------------------------------------------------------------------------------------------------------+
| `CalYear`                        | numeric for year of calculation. If df has more than one year, assigning CalYear truncates the data of only that year. |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------+
| `f_derivative`                   | logical if yes returns first derivative of gompertz curve.                                                             |
+----------------------------------+------------------------------------------------------------------------------------------------------------------------+

```{r}
jump_free_d309_noNA<-jump_free_d309|>filter(!is.na(jump_free_d309$value))
gompertz309<-dm.fit.gompertz(df=jump_free_d309_noNA,TreeNum=1,CalYear = 2021)
ggplot(gompertz309,aes(x=TIME,y=value))+geom_point()
```

# Calculating running correlation between dendrometer and climate data.

Calculates running correlation between dendrometer series and provided climate variables. Users can choose methods such as pearson, kendall and spearman, see `?cor.test` for further information.

```         
mov.cor.dm(df, Clim, TreeNum, win_size, cor_method = "pearson") 
out_corr<-mov.cor.dm(df=gf_nepa17, Clim=ktm_rain17, TreeNum=1, win_size=21) 
```

### Explanation of options

+----------------------------------+------------------------------------------------------------------------------------------------------------------+
| `win_size`                       | numerical, the running days windows of which the correlation is to be calculated. The minimum value is 18.       |
+----------------------------------+------------------------------------------------------------------------------------------------------------------+
| `cor_method`                     | string, method to be applied during correlation calculation. One of the following: pearson, kendall and spearman |
+----------------------------------+------------------------------------------------------------------------------------------------------------------+

### Output can be plotted with the following:

plot_mov.cor(mov.cor.output, sig.only = T, ci = 0.95, clim_vars = "all")\
Example code:\
out_corr\<-mov.cor.dm(df=gf_nepa17, Clim=ktm_rain17, TreeNum=1, win_size=21) plot_mov.cor(mov.cor.output=out_corr, sig.only=TRUE, ci=0.95)

### Arguments

+----------------------------------+----------------------------------------------------------------+
| `mov.cor.output`                 | list the output of `mov.cor.dm` function.                      |
+----------------------------------+----------------------------------------------------------------+
| `sig.only`                       | logical TRUE to plot only significant correlation.             |
+----------------------------------+----------------------------------------------------------------+
| `ci`                             | numeric confidence interval                                    |
+----------------------------------+----------------------------------------------------------------+
| `clim_vars`                      | array of climate variables or "all" for all climate variables. |
+----------------------------------+----------------------------------------------------------------+

```{r}

```
