---
title: "dendroanalyst"
format: pdf
editor: visual
---

# Code to run at launch

```{r}
#install.packages("dendRoAnalyst")
library(ggplot2)
library(dplyr)
library(readr)
library(viridis)
library(tidyr)
library(lubridate)
library(here)
library(zoo)
library(tidyverse)
library(dendRoAnalyst)
custom_colors <- c("2019" = "springgreen4", "2020" = "springgreen3", "2021" = "red", "2022" = "steelblue1", "2023" = "steelblue3","2024"="steelblue4")

#atomspheric drought vs soil drought in 2021 vs other years to tease out effects of drought vs tree experience
```

# Function Information

Summaries in the right column in bold indicate extra relevant functions. Example code for all these are in separate code chunks in their corresponding headings which you can see in the outline to the right.

+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [clim.twd](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/clim.twd.html)                                                                                                                                                                                                                                                                                             | **Calculating relative growth change during no-rain periods. Requires dendrometer (e.g., nepa17) and rainfall (e.g., rain17) data.**       |
|                                                                                                                                                                                                                                                                                                                                                                              |                                                                                                                                            |
| Description from ?clim.twd:                                                                                                                                                                                                                                                                                                                                                  | E.g.,\                                                                                                                                     |
|                                                                                                                                                                                                                                                                                                                                                                              | relative_dry_growth\<-clim.twd(df=gf_nepa17, Clim=ktm_rain17, dailyValue='max', showPlot=TRUE) \[1\]\                                      |
| This function calculates the number and the location of climatically adverse periods within a climate time series. The user can define a duration and threshold of these conditions. The function also provides the relative radial/circumferencial change during each adverse period for the original or normalized data. See Raffelsbauer et al., (2019) for more details. | Output: "The adverse climatic periods are listed below:" start_doy, end_doy, Id,                                                           |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [daily.data](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/daily.data.html)                                                                                                                                                                                                                                                                                         | Calculation of **daily statistics for dendrometer data**                                                                                   |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dendro.resample](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dendro.resample.html)                                                                                                                                                                                                                                                                               | Resampling temporal resolution of dendrometer and climate data                                                                             |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dendro.truncate](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dendro.truncate.html)                                                                                                                                                                                                                                                                               | Truncation of the dendrometer data                                                                                                         |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.detrend](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.detrend.html)                                                                                                                                                                                                                                                                                         | Detrend the dendrometer data (**should we?**)                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.fit.gompertz](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.fit.gompertz.html)                                                                                                                                                                                                                                                                               | Fitting gompertz function on annual dendrometer data                                                                                       |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.na.interpolation](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.na.interpolation.html)                                                                                                                                                                                                                                                                       | Detection and **interpolation of missing values** in dendrometer data.                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [gf_nepa17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/gf_nepa17.html)                                                                                                                                                                                                                                                                                           | Dendrometer data of Kathmandu for 2017 with gap filled                                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [i.jump.locator](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/i.jump.locator.html)                                                                                                                                                                                                                                                                                 | Removing artefacts due to manual adjustments of dendrometers interactively                                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [jump.locator](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/jump.locator.html)                                                                                                                                                                                                                                                                                     | **Removing artefacts due to manual adjustments of dendrometers automatically for more than one dendrometers**                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [ktm_rain17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/ktm_rain17.html)                                                                                                                                                                                                                                                                                         | Daily rainfall data of Kathmandu for 2017.                                                                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [mean_detrended.dm](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/mean_detrended.dm.html)                                                                                                                                                                                                                                                                           | Calculate mean of detrended dendrometer data.                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [mov.cor.dm](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/mov.cor.dm.html)                                                                                                                                                                                                                                                                                         | **Calculating running correlation between dendrometer data and daily climate.**                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [nepa](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa.html)                                                                                                                                                                                                                                                                                                     | Dendrometer data from Kathmandu                                                                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [nepa17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa17.html)                                                                                                                                                                                                                                                                                                 | Dendrometer data of Kathmandu for 2017                                                                                                     |
|                                                                                                                                                                                                                                                                                                                                                                              |                                                                                                                                            |
|                                                                                                                                                                                                                                                                                                                                                                              | from the paper: "The data was derived using an automatic logging band dendrometer (DRL26, EMS Brno) recording hourly circumference change" |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [pa2](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa2.html)                                                                                                                                                                                                                                                                                                     | Dendrometer data from Kathmandu version 2                                                                                                  |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [network.interpolation](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/network.interpolation.html)                                                                                                                                                                                                                                                                   | Interpolation of NA values using the dendrometer network                                                                                   |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [phase.sc](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/phase.sc.html)                                                                                                                                                                                                                                                                                             | **Application of the stem-cycle approach to calculate different phases, their duration and to plot them.**                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [phase.zg](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/phase.zg.html)                                                                                                                                                                                                                                                                                             | **Application of the zero-growth approach to calculate different phases, their duration and to plot them.**                                |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_mov.cor](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_mov.cor.html)                                                                                                                                                                                                                                                                                     | **Plotting moving correlation with climate.**                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_SC_output](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_SC_output.html)                                                                                                                                                                                                                                                                                 | **Plotting output of stem cycle approach (i.e., output of phase.sc function).**                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_ZG_output](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_ZG_output.html)                                                                                                                                                                                                                                                                                 | **Plotting output of zero-growth approach (i.e., output of phase.zg function).**                                                           |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [read.dendrometer](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/read.dendrometer.html)                                                                                                                                                                                                                                                                             | **Reading dendrometer data and automatically converts datetime format to what is needed.**                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [twd.maxima](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/twd.maxima.html)                                                                                                                                                                                                                                                                                         | **Locating the maxima of TWD periods**                                                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+

-   files.

-   A new function ‘mean.detrended.dm’ calculates the averaged detrended dendrometer data with possibility getting first order autocorrelation free averaged series.

# Step 0: Dendro data: loading in and prepping

Example data for DendRoAnalyst has hourly dendrometer data and daily rainfall data, so we will be condensing our data down to match.

Initial circumferences (mm)

```{r}
#raw dendrometer data
load(here(file="DendroAnalyst/dscmet_422_all.RData")) #calling saved R object from ecomatik.rmd
writexl::write_xlsx(
  dscmet_422_all,
  here("Data", "dscmet_422_all.xlsx")
)


dscmet_422_all[] <- lapply(dscmet_422_all, function(x) {
  if (is.numeric(x)) x[is.nan(x)] <- NA
  x
})
#checking bad flags - no bad ones found so this code won't be rerun 
unique_422_strings <- unique(
  unlist(
    sapply(
      dscmet_422_all, function(x) x[is.character(x)])))

#making df for each dendro to get a better look at things for circumference conversion
d309_data <- dscmet_422_all |> 
  select(c(Date,RecNum,DEND309_AVG_4200_0_01,Flag_DEND309_AVG_4200_0_01,DEND309_CIRC_4200_0_01,Flag_DEND309_CIRC_4200_0_01,DEND309_PLNG_4200_0_01,Flag_DEND309_PLNG_4200_0_01))


d310_data <-dscmet_422_all |> 
  select(c(Date,RecNum,DEND310_AVG_150_0_01,Flag_DEND310_AVG_150_0_01,DEND310_CIRC_150_0_01,Flag_DEND310_CIRC_150_0_01,DEND310_PLNG_150_0_01,Flag_DEND310_PLNG_150_0_01))

d311_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND311_AVG_150_0_01, Flag_DEND311_AVG_150_0_01, DEND311_CIRC_150_0_01, Flag_DEND311_CIRC_150_0_01, DEND311_PLNG_150_0_01, Flag_DEND311_PLNG_150_0_01))

d646_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND646_AVG_150_0_01, Flag_DEND646_AVG_150_0_01, DEND646_CIRC_150_0_01, Flag_DEND646_CIRC_150_0_01, DEND646_PLNG_150_0_01, Flag_DEND646_PLNG_150_0_01))

d648_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND648_AVG_150_0_01, Flag_DEND648_AVG_150_0_01, DEND648_CIRC_150_0_01, Flag_DEND648_CIRC_150_0_01, DEND648_PLNG_150_0_01, Flag_DEND648_PLNG_150_0_01))|>
  filter(DEND648_PLNG_150_0_01!="NaN")|>
  filter(DEND648_AVG_150_0_01!="NaN")

d651_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND651_AVG_150_0_01, Flag_DEND651_AVG_150_0_01, DEND651_CIRC_150_0_01, Flag_DEND651_CIRC_150_0_01, DEND651_PLNG_150_0_01, Flag_DEND651_PLNG_150_0_01))|>
  filter(DEND651_PLNG_150_0_01!="NaN")|>
  filter(DEND651_AVG_150_0_01!="NaN")

d652_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND652_AVG_150_0_01, Flag_DEND652_AVG_150_0_01, DEND652_CIRC_150_0_01, Flag_DEND652_CIRC_150_0_01, DEND652_PLNG_150_0_01, Flag_DEND652_PLNG_150_0_01))|>
  filter(DEND652_PLNG_150_0_01!="NaN")|>
  filter(DEND652_AVG_150_0_01!="NaN")

d654_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND654_AVG_150_0_01, Flag_DEND654_AVG_150_0_01, DEND654_CIRC_150_0_01, Flag_DEND654_CIRC_150_0_01, DEND654_PLNG_150_0_01, Flag_DEND654_PLNG_150_0_01))|>
  filter(DEND654_PLNG_150_0_01!="NaN")|>
  filter(DEND654_AVG_150_0_01!="NaN")

d655_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND655_AVG_150_0_01, Flag_DEND655_AVG_150_0_01, DEND655_CIRC_150_0_01, Flag_DEND655_CIRC_150_0_01, DEND655_PLNG_150_0_01, Flag_DEND655_PLNG_150_0_01))|>
  filter(DEND655_PLNG_150_0_01!="NaN")|>
  filter(DEND655_AVG_150_0_01!="NaN")

d656_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND656_AVG_150_0_01, Flag_DEND656_AVG_150_0_01, DEND656_CIRC_150_0_01, Flag_DEND656_CIRC_150_0_01, DEND656_PLNG_150_0_01, Flag_DEND656_PLNG_150_0_01))|>
  filter(DEND656_PLNG_150_0_01!="NaN")|>
  filter(DEND656_AVG_150_0_01!="NaN")

d860_data <- dscmet_422_all |>
  select(c(Date, RecNum, DEND860_AVG_150_0_01, Flag_DEND860_AVG_150_0_01, DEND860_CIRC_150_0_01, Flag_DEND860_CIRC_150_0_01, DEND860_PLNG_150_0_01, Flag_DEND860_PLNG_150_0_01))|>
  filter(DEND860_PLNG_150_0_01!="NaN")|>
  filter(DEND860_AVG_150_0_01!="NaN")


#convert raw data to circumference
d309_circ <- dc3_to_circumference(
  Vi_micrometers=dscmet_422_all$DEND309_AVG_4200_0_01, C0= 1586, V0 = 2.353)

d310_circ <- dc3_to_circumference(
  Vi_micrometers=dscmet_422_all$DEND310_AVG_150_0_01, C0= 3402, V0 = 3.462)

d311_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND311_AVG_150_0_01, C0 = 2755, V0 = 4.738)
# option is 2755 & 4.738 or C0 = 2743, V0 = 0


d646_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND646_AVG_150_0_01, C0 = 1835, V0 = 2.968)

d648_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND648_AVG_150_0_01, C0 = 2013.8#, V0 = 100
  )

d651_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND651_AVG_150_0_01, C0 = 1940.5#,V0 = 100
  )

d652_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND652_AVG_150_0_01, C0 = 1402.3#, V0 = 100
  )

d654_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND654_AVG_150_0_01, C0 = 1170.5#, V0 = 1522
  )

d655_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND655_AVG_150_0_01, C0 = 884.8, V0 = 3119)

d656_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND656_AVG_150_0_01, C0 = 1221.3, V0 = 3518)

d860_circ <- dc3_to_circumference(
  Vi_micrometers = dscmet_422_all$DEND860_AVG_150_0_01, C0 = 760.3, V0 = 2340)
```

## Daily and Hourly Means and Identifyinhg threshhold for Step 2

```{r}
#getting hourly averages - starting with adding an hour column
dscmet_422_all$hour <- format(
  dscmet_422_all$Date, "%Y-%m-%d %H:00:00")
dscmet_422_all$Date<-as.POSIXct(dscmet_422_all$hour,format="%Y-%m-%d %H")
#getting daily averages - starting with adding a day column
dscmet_422_all$day <- format(
  dscmet_422_all$Date, "%Y-%m-%d")
dscmet_422_all$day<-as.POSIXct(dscmet_422_all$day,format="%Y-%m-%d")

#309 - OG DF 
#hourly
d309_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND309_AVG_4200_0_01)) |> filter(hour >= as.POSIXct("2019-01-01"))

#daily values 
d309_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND309_AVG_4200_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
d309_daily_diff <- d309_daily |> arrange(day)|> mutate(
    value_diff = value - lag(value))
#creating seprate DFs to remove NA values
d309_hourly_noNA <- d309_hourly |> filter(!is.nan(value))
d309_daily_noNA <- d309_daily |> filter(!is.nan(value))


#310 - OG DF
#hourly
d310_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND310_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d310_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND310_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d310_hourly_noNA <- d310_hourly |> filter(!is.nan(value))
d310_daily_noNA <- d310_daily |> filter(!is.nan(value))

#311 - OG DF
#hourly
d311_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND311_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d311_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND311_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d311_hourly_noNA <- d311_hourly |> filter(!is.nan(value))
d311_daily_noNA <- d311_daily |> filter(!is.nan(value))

#654 - Secondary DF
#hourly
d654_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND654_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d654_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND654_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d654_hourly_noNA <- d654_hourly |> filter(!is.nan(value))
d654_daily_noNA <- d654_daily |> filter(!is.nan(value))

#655 - Secondary DF
#hourly
d655_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND655_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d655_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND655_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d655_hourly_noNA <- d655_hourly |> filter(!is.nan(value))
d655_daily_noNA <- d655_daily |> filter(!is.nan(value))

#656 - Secondary DF
#hourly
d656_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND656_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d656_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND656_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d656_hourly_noNA <- d656_hourly |> filter(!is.nan(value))
d656_daily_noNA <- d656_daily |> filter(!is.nan(value))

#860 - Secondary DF - note only ~15% of original canopy remains alive
#hourly
d860_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND860_AVG_150_0_01))|>
  filter(hour > as.POSIXct("2019-05-28") & hour < as.POSIXct("2024-01-01"))
#daily
d860_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND860_AVG_150_0_01))|>
  filter(day > as.POSIXct("2019-05-28") & day < as.POSIXct("2024-01-01"))
#creating seprate DFs to remove NA values
d860_hourly_noNA <- d860_hourly |> filter(!is.nan(value))
d860_daily_noNA <- d860_daily |> filter(!is.nan(value))

#646 OG WH
#hourly
d646_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND646_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d646_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND646_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d646_hourly_noNA <- d646_hourly |> filter(!is.nan(value))
d646_daily_noNA <- d646_daily |> filter(!is.nan(value))

#648 OG WH
d648_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND648_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d648_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND648_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d648_hourly_noNA <- d648_hourly |> filter(!is.nan(value))
d648_daily_noNA <- d648_daily |> filter(!is.nan(value))

#651 OG WRC
#hourly
d651_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND651_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d651_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND651_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01")) 
#creating seprate DFs to remove NA values
d651_hourly_noNA <- d651_hourly |> filter(!is.nan(value))
d651_daily_noNA <- d651_daily |> filter(!is.nan(value))

#652 OG WH
#hourly
d652_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(value = mean(DEND652_AVG_150_0_01))|>
  filter(hour >= as.POSIXct("2019-01-01"))
#daily
d652_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(value = mean(DEND652_AVG_150_0_01))|>
  filter(day >= as.POSIXct("2019-01-01"))
#creating seprate DFs to remove NA values
d652_hourly_noNA <- d652_hourly |> filter(!is.nan(value))
d652_daily_noNA <- d652_daily |> filter(!is.nan(value))

#combining into one DF 

```

## Getting Median and STD Differences

```{r}
tree_ids <- c(309, 310, 311, 646, 648, 651, 652, 654, 655, 656, 860)
# Create lists of hourly and daily dataframes
hourly_dfs <- setNames(lapply(tree_ids, function(id) get(paste0("d", id, "_hourly"))), tree_ids)
daily_dfs  <- setNames(lapply(tree_ids, function(id) get(paste0("d", id, "_daily"))), tree_ids)
#Loop function for getting needed threshold values 
calculate_diff_stats <- function(df, tree_id, timescale, time_col) {
  df_diff <- df %>%
    arrange(.data[[time_col]]) %>%
    mutate(value_diff = value - lag(value)) %>%
    filter(!is.na(value_diff))
  
  median_diff <- median(df_diff$value_diff)
  mad_diff    <- mad(df_diff$value_diff)
  threshold_value <- median_diff + 2 * mad_diff
  
  tibble(
    tree_id = tree_id,
    timescale = timescale,
    median_diff = median_diff,
    mad_diff = mad_diff,
    threshold_value = threshold_value
  )
}


#applying this function to our data
hourly_summaries <- purrr::imap_dfr(hourly_dfs, ~calculate_diff_stats(.x, tree_id = .y, timescale = "hourly", time_col = "hour"))
daily_summaries  <- purrr::imap_dfr(daily_dfs,  ~calculate_diff_stats(.x, tree_id = .y, timescale = "daily", time_col = "day"))

combined_summary <- bind_rows(hourly_summaries, daily_summaries)
```

309 - 1586 (july 12 2018)\
310 - 3402 (july 12 2018)\
311 - 2743 (july 13 2018)\
646 - (july 13 2018)1835\
648 = 2013.8\
651 = 1940.5\
652 = 1402.3\
654 = 1170.5\
655 = 884.8\
656 = 1221.3\
860 = 760.3

```         
```

```{r}

```

# Step 0: Climate data

### Please visit "GJ_OSU_Thesis/Behind the scenes code that only needs to be read once.Rmd" for code on eliminating values with bad flags and restructuring/summarizing data

These climate variables include hourly and daily mean VPD & air temp; hourly and daily max mean VPD; and hourly and daily max air temp. HDI code is in RWL.qmd

```{r}
#means
load(here(file="Climate/filtered_Dscmet_VPD52m_hourly.RData"))
load(here(file="Climate/filtered_Dscmet_VPD52m_daily.RData"))
load(here(file="Climate/filtered_Dscmet_AT52m_hourly.RData"))
load(here(file="Climate/filtered_Dscmet_AT52m_daily.RData"))
load(here(file="EL RWL/Dscmet_HDI_hourly_mean.RData"))
load(here(file="EL RWL/Dscmet_HDI_daily_mean.RData"))
#maxes
load(here(file="Climate/filtered_Dscmet_maxVPD52m_hourly.RData"))
load(here(file="Climate/filtered_Dscmet_maxVPD52m_daily.RData"))
load(here(file="Climate/filtered_Dscmet_maxAT52m_hourly.RData"))
load(here(file="Climate/filtered_Dscmet_maxAT52m_daily.RData"))
load(here(file="EL RWL/Dscmet_HDI_hourly_max.RData"))
load(here(file="EL RWL/Dscmet_HDI_daily_max.RData"))


```

# Step 1: NA Interpolation

Skipped for now because the paper says that due to increasing uncertainty, this step is not recommended for gaps larger than 24 hours

## Interpolation option 1: dm.na.interpolation

```         
dm.na.interpolation(df, resolution, fill = FALSE, method = "spline") 
```

### Explanation of options

+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                              | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and following columns with dendrometer data for the same temporal resolution and time period. |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `resolution`                      | integer, indicating the resolution of dendrometer data in **minutes**.                                                                                                                 |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `fill`                            | logical, if `TRUE` it fills the `NA` values using spline interpolation. Default is `FALSE`.                                                                                            |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `method`                          | string, 'spline' for the spline interpolation or 'seasonal' for the seasonal interpolation                                                                                             |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Locating maxima TWD Periods with twd.maxima

`{r} # library(dendRoAnalyst) #> data(gf_nepa17) #> df1=gf_nepa17[2500:3500,] #> twd_max<-twd.maxima(df=df1, TreeNum=2, showPlot=FALSE) #> head(twd_max,10) #             start.time            end.time                Time         TWD #20  2017-04-15 11:00:00 2017-04-16 04:00:00 2017-04-15 22:00:00 0.028245289 #43  2017-04-16 08:00:00 2017-04-18 06:00:00 2017-04-16 21:00:00 0.070221431 #    duration_from_start twd.number #20                  540          1 #43                  660          2}`

## Interpolation option 2: using dendrometer network via network.interpolation

```         
network.interpolation(df, referenceDF, niMethod) 
```

### Explanation of options

+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                              | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and dendrometer data in the second column and onward. The data gaps must be filled with `NA` using the gap.interpolation function. |
+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `referenceDF`                     | dataframe with other dendrometers to be used as reference for the interpolation. The more dendrometers are included, the more robust will be the interpolation.                                                             |
+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `niMethod`                        | string, either *'linear'* or *'proportional'* for interpolation method.                                                                                                                                                     |
+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

`{r} #Example from vignette #df1<-gf_nepa17 #> # Creating an artificial reference dataset. #> df2<-cbind(gf_nepa17,gf_nepa17[,2:3],gf_nepa17[,2:3]) #> # Creating gaps in dataset by replacing some of the reading with NA in dataset. #> df1[40:50,3]<-NA #> # Using proportional interpolation method. #> df1_NI<-network.interpolation(df=df1, referenceDF=df2, niMethod='proportional') #Warning in norm.inter(t, adj.alpha) : #  extreme order statistics used as endpoints #[1] "Interpolation in T3" #         Time Interpolated_value    LCI95    UCI95 #1  1483369200           48.31525 48.30600 48.32141 #2  1483372800           48.31522 48.31331 48.31585 #> head(df1_NI,10) #                  Time       T2       T3 #1  2017-01-01 00:00:00 62.26459 48.16801 #...10 2017-01-01 09:00:00 62.46552 48.21498}`

# Step 2: Remove artefacts from manual adjustments with jump.locator

```         
i.jump.locator(df, TreeNum, v) 
```

### Explanation of options for similar function i.jump.locator (only difference is that it adds TreeNum whereas jump.locator doesn't)

output is a new df that locates and adjusts the artefacts

+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                              | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and the dendrometer data in following columns.                                                        |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`                         | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                                              |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `v`                               | numerical value which is considered as artefact. E.g. `v`=1 implies that if the difference to the consecutive data point is more than 1 or less than -1, it will be considered as an artefact. |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

#### Examples from vignette:

`{r} jump_free_nepa<-jump.locator(df=nepa,v=1)}`

#### Using our own data:

First look at the raw data outputs and see how the jumps are to help set the "v=" value. When you run the function it makes a table with the number of rows equal to the number of removed values and the columns equal to observation number, timestamp of the jump, and the value of the jump. If you name the jump locator object, that will add an object to your environment of the adjusted df with the jump values removed. 1.5MAD+med

```{r}
#309
jf_d309_hourly<-jump.locator(df=d309_hourly,v=10.99885) 
jf_d309_daily<-jump.locator(df=d309_daily,v=72.800966) 
jf_d309_hourly_noNA <-jump.locator(df=d309_hourly_noNA,v=10.99885) 
jf_d309_daily_noNA <-jump.locator(df=d309_daily_noNA,v=72.800966) 

#310
jf_d310_hourly<-jump.locator(df=d310_hourly,v=1.463942)
jf_d310_daily<-jump.locator(df=d310_daily,v=61.102741)
jf_d310_hourly_noNA <-jump.locator(df=d310_hourly_noNA,v=1.463942) 
jf_d310_daily_noNA <-jump.locator(df=d310_daily_noNA,v=61.102741)

# 311
jf_d311_hourly <- jump.locator(df = d311_hourly, v = 2.4739)
jf_d311_daily <- jump.locator(df = d311_daily, v = 95.132753)
jf_d311_hourly_noNA <- jump.locator(df = d311_hourly_noNA, v = 2.4739)
jf_d311_daily_noNA <- jump.locator(df = d311_daily_noNA, v = 95.132753)

# 646
jf_d646_hourly <- jump.locator(df = d646_hourly, v = 3.771175)
jf_d646_daily <- jump.locator(df = d646_daily, v = 47.968028)
jf_d646_hourly_noNA <- jump.locator(df = d646_hourly_noNA, v = 3.771175)
jf_d646_daily_noNA <- jump.locator(df = d646_daily_noNA, v = 47.968028)

# 648
jf_d648_hourly <- jump.locator(df = d648_hourly, v = 6.9217)
jf_d648_daily <- jump.locator(df = d648_daily, v = 60.450549)
jf_d648_hourly_noNA <- jump.locator(df = d648_hourly_noNA, v = 6.9217)
jf_d648_daily_noNA <- jump.locator(df = d648_daily_noNA, v = 60.450549)

# 651
jf_d651_hourly <- jump.locator(df = d651_hourly, v = 13.574742)
jf_d651_daily <- jump.locator(df = d651_daily, v = 92.584206)
jf_d651_hourly_noNA <- jump.locator(df = d651_hourly_noNA, v = 13.574742)
jf_d651_daily_noNA <- jump.locator(df = d651_daily_noNA, v = 92.584206)

# 652
jf_d652_hourly <- jump.locator(df = d652_hourly, v = 2.205242)
jf_d652_daily <- jump.locator(df = d652_daily, v = 50.242755)
jf_d652_hourly_noNA <- jump.locator(df = d652_hourly_noNA, v = 2.205242)
jf_d652_daily_noNA <- jump.locator(df = d652_daily_noNA, v = 50.242755)

# 654
jf_d654_hourly <- jump.locator(df = d654_hourly, v = 2.575892)
jf_d654_daily <- jump.locator(df = d654_daily, v = 60.339541)
jf_d654_hourly_noNA <- jump.locator(df = d654_hourly_noNA, v = 2.575892)
jf_d654_daily_noNA <- jump.locator(df = d654_daily_noNA, v = 60.339541)

# 655
jf_d655_hourly <- jump.locator(df = d655_hourly, v = 6.55105)
jf_d655_daily <- jump.locator(df = d655_daily, v = 72.565134)
jf_d655_hourly_noNA <- jump.locator(df = d655_hourly_noNA, v = 6.55105)
jf_d655_daily_noNA <- jump.locator(df = d655_daily_noNA, v = 72.565134)

# 656
jf_d656_hourly <- jump.locator(df = d656_hourly, v = 2.946542)
jf_d656_daily <- jump.locator(df = d656_daily, v = 51.728442)
jf_d656_hourly_noNA <- jump.locator(df = d656_hourly_noNA, v = 2.946542)
jf_d656_daily_noNA <- jump.locator(df = d656_daily_noNA, v = 51.728442)

# 860
jf_d860_hourly <- jump.locator(df = d860_hourly, v = 2.575892)
jf_d860_daily <- jump.locator(df = d860_daily, v = 47.159791)
jf_d860_hourly_noNA <- jump.locator(df = d860_hourly_noNA, v = 2.575892)
jf_d860_daily_noNA <- jump.locator(df = d860_daily_noNA, v = 47.159791)

```

##### Plotting new data to see the difference: One curve per dendro: raw vs jump-free

```{r}
#309 hourly
d309_hourly$date<-as.Date(d309_hourly$hour)
jf_d309_hourly_plot <- ggplot(
  jf_d309_hourly,aes(x=TIME,y=value))+geom_point()+labs(
      title="OG DF (ID 309) Hourly Adjusted Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d309_hourly_plot <- ggplot(
  d309_hourly,aes(x=date,y=value))+geom_point()+labs(
      title="OG DF (ID 309) Raw Hourly Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()+
  scale_x_date(
    limits = as.Date(c("2019-01-01", "2024-12-31")),
    date_breaks = "45 days",
    date_labels = "%y %b") + theme(
      axis.text.x = element_text(angle = 45, hjust = 1,size=8.5))
print(raw_d309_hourly_plot)
gridExtra::grid.arrange(raw_d309_hourly_plot,jf_d309_hourly_plot)

#309 daily 
jf_d309_daily_plot <- ggplot(
  jf_d309_daily,aes(x=TIME,y=value))+geom_point()+labs(
      title="OG DF (ID 309) Daily Adjusted Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d309_daily_plot <- ggplot(
  d309_daily,aes(x=day,y=value))+geom_point()+labs(
      title="OG DF (ID 309) Raw Daily Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d309_daily_plot,jf_d309_daily_plot)


# 310 hourly
jf_d310_hourly_plot <- ggplot(
  jf_d310_hourly, aes(x = TIME, y = value)) +
  geom_point() +
  labs(
    title = "OG DF (ID 310) Hourly Adjusted Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)"
  ) +
  theme_minimal()

raw_d310_hourly_plot <- ggplot(
  d310_hourly, aes(x = hour, y = value)) +
  geom_point() +
  labs(
    title = "OG DF (ID 310) Raw Hourly Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)"
  ) +
  theme_minimal()

gridExtra::grid.arrange(raw_d310_hourly_plot, jf_d310_hourly_plot)

# 310 daily
jf_d310_daily_plot <- ggplot(
  jf_d310_daily, aes(x = TIME, y = value)) +
  geom_point() +
  labs(
    title = "OG DF (ID 310) Daily Adjusted Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)"
  ) +
  theme_minimal()

raw_d310_daily_plot <- ggplot(
  d310_daily, aes(x = day, y = value)) +
  geom_point() +
  labs(
    title = "OG DF (ID 310) Raw Daily Dendrometer Value",
    x = "Date",
    y = "Dendrometer Value (µm)"
  ) +
  theme_minimal()
gridExtra::grid.arrange(raw_d310_daily_plot, jf_d310_daily_plot)


#311 hourly
jf_d311_hourly_plot <- ggplot(
  jf_d311_hourly,aes(x=TIME,y=value))+geom_point()+labs(
      title="OG DF (ID 311) Hourly Adjusted Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d311_hourly_plot <- ggplot(
  d311_hourly,aes(x=hour,y=value))+geom_point()+labs(
      title="OG DF (ID 311) Raw Hourly Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d311_hourly_plot,jf_d311_hourly_plot)

#311 daily
jf_d311_daily_plot <- ggplot(
  jf_d311_daily,aes(x=TIME,y=value))+geom_point()+labs(
      title="OG DF (ID 311) Daily Adjusted Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d311_daily_plot <- ggplot(
  d311_daily,aes(x=day,y=value))+geom_point()+labs(
      title="OG DF (ID 311) Raw Daily Dendrometer Value",
      x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d311_daily_plot,jf_d311_daily_plot)

#646 hourly
jf_d646_hourly_plot <- ggplot(jf_d646_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 646) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d646_hourly_plot <- ggplot(d646_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 646) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d646_hourly_plot,jf_d646_hourly_plot)

#646 daily
jf_d646_daily_plot <- ggplot(jf_d646_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 646) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d646_daily_plot <- ggplot(d646_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 646) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d646_daily_plot,jf_d646_daily_plot)

#648 hourly
jf_d648_hourly_plot <- ggplot(jf_d648_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 648) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d648_hourly_plot <- ggplot(d648_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 648) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d648_hourly_plot,jf_d648_hourly_plot)

#648 daily
jf_d648_daily_plot <- ggplot(jf_d648_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 648) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d648_daily_plot <- ggplot(d648_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 648) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d648_daily_plot,jf_d648_daily_plot)

#651 hourly
jf_d651_hourly_plot <- ggplot(jf_d651_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 651) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d651_hourly_plot <- ggplot(d651_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 651) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d651_hourly_plot,jf_d651_hourly_plot)

#651 daily
jf_d651_daily_plot <- ggplot(jf_d651_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 651) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d651_daily_plot <- ggplot(d651_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 651) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d651_daily_plot,jf_d651_daily_plot)

#652 hourly
jf_d652_hourly_plot <- ggplot(jf_d652_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 652) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d652_hourly_plot <- ggplot(d652_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 652) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d652_hourly_plot,jf_d652_hourly_plot)

#652 daily
jf_d652_daily_plot <- ggplot(jf_d652_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 652) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d652_daily_plot <- ggplot(d652_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 652) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d652_daily_plot,jf_d652_daily_plot)

#654 hourly
jf_d654_hourly_plot <- ggplot(jf_d654_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 654) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d654_hourly_plot <- ggplot(d654_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 654) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d654_hourly_plot,jf_d654_hourly_plot)

#654 daily
jf_d654_daily_plot <- ggplot(jf_d654_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 654) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d654_daily_plot <- ggplot(d654_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 654) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d654_daily_plot,jf_d654_daily_plot)

#655 hourly
jf_d655_hourly_plot <- ggplot(jf_d655_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 655) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d655_hourly_plot <- ggplot(d655_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 655) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d655_hourly_plot,jf_d655_hourly_plot)

#655 daily
jf_d655_daily_plot <- ggplot(jf_d655_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 655) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d655_daily_plot <- ggplot(d655_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 655) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d655_daily_plot,jf_d655_daily_plot)

#656 hourly
jf_d656_hourly_plot <- ggplot(jf_d656_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 656) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d656_hourly_plot <- ggplot(d656_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 656) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d656_hourly_plot,jf_d656_hourly_plot)

#656 daily
jf_d656_daily_plot <- ggplot(jf_d656_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 656) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d656_daily_plot <- ggplot(d656_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 656) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d656_daily_plot,jf_d656_daily_plot)

#860 hourly
jf_d860_hourly_plot <- ggplot(jf_d860_hourly,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 860) Hourly Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d860_hourly_plot <- ggplot(d860_hourly,aes(x=hour,y=value))+geom_point()+labs(
  title="OG DF (ID 860) Raw Hourly Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d860_hourly_plot,jf_d860_hourly_plot)

#860 daily
jf_d860_daily_plot <- ggplot(jf_d860_daily,aes(x=TIME,y=value))+geom_point()+labs(
  title="OG DF (ID 860) Daily Adjusted Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
raw_d860_daily_plot <- ggplot(d860_daily,aes(x=day,y=value))+geom_point()+labs(
  title="OG DF (ID 860) Raw Daily Dendrometer Value",x="Date",y="Dendrometer Value (µm)") +theme_minimal()
gridExtra::grid.arrange(raw_d860_daily_plot,jf_d860_daily_plot)

#viewing plots
print(raw_d309_hourly_plot)
```

# Step 2.1 (optional): Adjusting temporal resolution via dendro.resample

```         
dendro.resample(df, by, value) 
```

### Explanation of options

+-----------------------------------+--------------------------------------------------------------------------------------------+
| `df`                              | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`.  |
+-----------------------------------+--------------------------------------------------------------------------------------------+
| `by`                              | either *H, D, W* or *M* to resample data into hourly, daily, weekly or monthly resolution. |
+-----------------------------------+--------------------------------------------------------------------------------------------+
| `value`                           | either *max, min*, *mean* or *sum* for the resampling value.                               |
+-----------------------------------+--------------------------------------------------------------------------------------------+

```{r}
# To resample monthly with maximum value
resample_M<-dendro.resample(df=nepa17[,1:2], by='M', value='max')
head(resample_M,10)
#example output
#         Time       T2
#1  2017-01-01 63.90855
#2  2017-02-01 64.88705
```

# Step 2.2 (optional): Extracting specific values with dendro.truncate

```         
dendro.truncate(df, CalYear, DOY) 
```

### Explanation of options

+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                              | dataframe with the first column named date and time in the format `yyyy-mm-dd HH:MM:SS`.                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `CalYear`                         | desired year of calculation.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `DOY`                             | If we provide an array instead of a single value for `CalYear` and a single value for `DOY`, it truncates data from the `DOY` of the first `CalYear` to the same `DOY` of the second `CalYear`. Conversely, if we provide one value for `CalYear` and an array of two elements for `DOY` truncates the data form first `DOY` to second `DOY` within the same `CalYear`. Finally, if we provide an array with two values for both `DOY` and `CalYear`, it truncates data from the first `DOY` of the first `CalYear` to the second `DOY` of second `CalYear`. |
+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#Example extracting data from doy 20 to 50 in 2017.
#trunc1<-dendro.truncate(df=nepa, CalYear=2017, DOY=c(20,50))
#head(trunc1,10)
#                 Time       T2       T3
#1  2017-01-20 00:00:00 64.92424 48.87772
#...10 2017-01-20 09:00:00 64.96733 48.88450
```

# Step 3: Analyses without Climate Data

## Phase differentiation using two approaches

```         
phase.sc(df, TreeNum, smoothing = NULL) 
```

### Explanation of options for stem cycle approach

three phases in this approach: 1) Shrinkage, when the dendrometer reading is less than previous reading, 2) Expansion, when current reading is more than previous reading and 3) Increment, when current reading is higher than the previous maximum. Output can be plotted with code like this:

**plot_SC_output( SC_output, DOY, Year, cols = c("#fee8c8", "#fdbb84", "#e34a33"), phNames = c("Shrinkage", "Expansion", "Increment") )**

+-----------------------------------+----------------------------------------------------------------------------------------------------------------+
| `DOY`                             | array with initial and final day for plotting. E.g. *c(a,b)*, where a = initial date and b = final date.       |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------+
| `cols`                            | array of three elements: colour for each phases.                                                               |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------+
| `phNames`                         | array with three elements for three different phases. Default is **"Shrinkage", "Expansion" and "Increment"**. |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------+

#### Inputs for phase.sc:

+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                              | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`. It should contain data with constant temporal resolution for best results.                                                                                                                    |
+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`                         | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                                                                                                                                       |
+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `smoothing`                       | numerical value from 1 to 12 which indicates the length of the smoothing spline, i.e. 1 = 1 hour and 12 = 12 hours. Default is `NULL` for no smoothing.The function [`smooth.Pspline`](http://127.0.0.1:17265/help/library/pspline/help/smooth.Pspline) is used for smoothing the data. |
+-----------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Outputs a list of two dataframes. The first dataframe *SC_cycle* with cyclic phases along with various statistics and the second dataframe *SC_phase* with assigned phases for each data point.The dataframe *SC_cycle* contains the beginning, end, duration, magnitude and rate of each phase. The dataframe *SC_phase* contains time and corresponding phases during that time.

```         
phase.zg(df, TreeNum) 
```

### Explanation of options for zero-growth approach

+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`                              | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`. It should contain data with constant temporal resolution for best results. |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`                         | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                    |
+-----------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Outputs a list of two dataframes. The first dataframe *ZG_cycle* contains the cyclic phases along with various statistics and the second dataframe *ZG_phase* with assigned phases for each data point.

#### Output can be plotted with th following code:

```         
plot_ZG_output(ZG_output, DOY, Year) 
```

+-----------------------------------+----------------------------------------------------------------------------------------------------------+
| `ZG_output`                       | list the output of `phase.zg` function.                                                                  |
+-----------------------------------+----------------------------------------------------------------------------------------------------------+
| `DOY`                             | array with initial and final day for plotting. E.g. *c(a,b)*, where a = initial date and b = final date. |
+-----------------------------------+----------------------------------------------------------------------------------------------------------+
| `Year`                            | array for indicating year for plotting.                                                                  |
+-----------------------------------+----------------------------------------------------------------------------------------------------------+

Two subplots with GRO and TWD respectively.

```{r}
#SC Example
sc_phase<-phase.sc(df=nepa17, TreeNum=1, smoothing=1)
plot_SC_output<-plot_SC_output(sc_phase,DOY = c(164,193),Year = 2017,cols = c("red","green","blue"))
plot_SC_output

#ZG Example
zg_phase<-phase.zg(df=nepa17, TreeNum=1)
plot_ZG_output<-plot_ZG_output(zg_phase,DOY = c(164,193),Year = 2017)
plot_ZG_output
```

### Using Our Data - hourly jump-free: data with NAs removed: HD +/- 2 weeks

```{r}
#309
sc_jf_d309_hourly_noNA<-phase.sc(
  df=jf_d309_hourly_noNA, TreeNum=1, smoothing=1)
sc_jf_d309_hourly_noNA_plot<-plot_SC_output(
  sc_jf_d309_hourly_noNA,DOY = c(164,193),Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d309_hourly_noNA_plot

zg_jf_d309_hourly_noNA<-phase.zg(
  df=jf_d309_hourly_noNA, TreeNum=1)
zg_jf_d309_hourly_noNA_plot<-plot_ZG_output(
  zg_jf_d309_hourly_noNA,DOY = c(164,193),Year = 2021)
zg_jf_d309_hourly_noNA_plot



#310
sc_jf_d310_hourly_noNA <- phase.sc(
  df = jf_d310_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d310_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d310_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d310_hourly_noNA_plot

zg_jf_d310_hourly_noNA <- phase.zg(
  df = jf_d310_hourly_noNA, TreeNum = 1)
zg_jf_d310_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d310_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d310_hourly_noNA_plot



#311
sc_jf_d311_hourly_noNA <- phase.sc(
  df = jf_d311_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d311_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d311_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d311_hourly_noNA_plot

zg_jf_d311_hourly_noNA <- phase.zg(
  df = jf_d311_hourly_noNA, TreeNum = 1)
zg_jf_d311_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d311_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d311_hourly_noNA_plot



#646
sc_jf_d646_hourly_noNA <- phase.sc(
  df = jf_d646_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d646_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d646_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d646_hourly_noNA_plot

zg_jf_d646_hourly_noNA <- phase.zg(
  df = jf_d646_hourly_noNA, TreeNum = 1)
zg_jf_d646_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d646_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d646_hourly_noNA_plot



#648
sc_jf_d648_hourly_noNA <- phase.sc(
  df = jf_d648_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d648_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d648_hourly_noNA, DOY = c(164,193), Year = 2023,
  cols = c("red","green","blue"))
sc_jf_d648_hourly_noNA_plot

zg_jf_d648_hourly_noNA <- phase.zg(
  df = jf_d648_hourly_noNA, TreeNum = 1)
zg_jf_d648_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d648_hourly_noNA, DOY = c(164,193), Year = 2023)
zg_jf_d648_hourly_noNA_plot



#651
sc_jf_d651_hourly_noNA <- phase.sc(
  df = jf_d651_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d651_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d651_hourly_noNA, DOY = c(164,193), Year = 2023,
  cols = c("red","green","blue"))
sc_jf_d651_hourly_noNA_plot

zg_jf_d651_hourly_noNA <- phase.zg(
  df = jf_d651_hourly_noNA, TreeNum = 1)
zg_jf_d651_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d651_hourly_noNA, DOY = c(164,193), Year = 2023)
zg_jf_d651_hourly_noNA_plot



#652
sc_jf_d652_hourly_noNA <- phase.sc(
  df = jf_d652_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d652_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d652_hourly_noNA, DOY = c(164,193), Year = 2023,
  cols = c("red","green","blue"))
sc_jf_d652_hourly_noNA_plot

zg_jf_d652_hourly_noNA <- phase.zg(
  df = jf_d652_hourly_noNA, TreeNum = 1)
zg_jf_d652_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d652_hourly_noNA, DOY = c(164,193), Year = 2023)
zg_jf_d652_hourly_noNA_plot



#654
sc_jf_d654_hourly_noNA <- phase.sc(
  df = jf_d654_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d654_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d654_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d654_hourly_noNA_plot

zg_jf_d654_hourly_noNA <- phase.zg(
  df = jf_d654_hourly_noNA, TreeNum = 1)
zg_jf_d654_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d654_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d654_hourly_noNA_plot



#655
sc_jf_d655_hourly_noNA <- phase.sc(
  df = jf_d655_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d655_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d655_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d655_hourly_noNA_plot

zg_jf_d655_hourly_noNA <- phase.zg(
  df = jf_d655_hourly_noNA, TreeNum = 1)
zg_jf_d655_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d655_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d655_hourly_noNA_plot



#656
sc_jf_d656_hourly_noNA <- phase.sc(
  df = jf_d656_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d656_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d656_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d656_hourly_noNA_plot

zg_jf_d656_hourly_noNA <- phase.zg(
  df = jf_d656_hourly_noNA, TreeNum = 1)
zg_jf_d656_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d656_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d656_hourly_noNA_plot



#860
sc_jf_d860_hourly_noNA <- phase.sc(
  df = jf_d860_hourly_noNA, TreeNum = 1, smoothing = 1)
sc_jf_d860_hourly_noNA_plot <- plot_SC_output(
  sc_jf_d860_hourly_noNA, DOY = c(164,193), Year = 2021,
  cols = c("red","green","blue"))
sc_jf_d860_hourly_noNA_plot

zg_jf_d860_hourly_noNA <- phase.zg(
  df = jf_d860_hourly_noNA, TreeNum = 1)
zg_jf_d860_hourly_noNA_plot <- plot_ZG_output(
  zg_jf_d860_hourly_noNA, DOY = c(164,193), Year = 2021)
zg_jf_d860_hourly_noNA_plot

#saving plots
png(here(""))

 ggsave(filename = here(""),    plot = plot_list[[id]],
  )}
```

## Daily stats: Calculating and making amplitude plots (max-min) from daily.data

```         
daily.data(df, TreeNum) 
```

This function calculates various statistics of dendrometer data on a daily basis. The daily statistics includes the daily maximum and minimum with their corresponding times and daily amplitude (difference between daily maximum and minimum). See King et al. (2013) for details. Output is a dataframe with the daily statistics of the dendrometer data and Remarks = "\*" if Time_max \> Time_min otherwise "".

```{r}
# 309
daily_stats309<-daily.data(df=jf_d309_hourly_noNA, TreeNum=1)
daily_stats309$yr <- format(
  daily_stats309$DATE, "%Y")
daily_stats309$yr<-as.factor(daily_stats309$yr)
daily_stats309$DOY<-yday(daily_stats309$DATE)
daily_stats309$ID<-as.factor("309")
daily_stats309$AgeClass<-as.factor("OG")
daily_stats309$Species<-as.factor("DF")
#not sure if this is needed but here is the code to make amplitude_signed
daily_stats309$amplitude_signed <- ifelse(
  daily_stats309$Remarks == "*",
  abs(daily_stats309$amplitude),     
  -abs(daily_stats309$amplitude))
save(daily_stats309,file = here(
  "DendroAnalyst/daily_stats309.RData"))

jf_309amp <- ggplot(
  daily_stats309,aes(x=DOY,y=amplitude,color=factor(yr)))+geom_point()+labs(
      title="HJA Douglas-fir (ID 309) Daily Circumference Amplitude",
      x="Day of Year",y="Daily Stem Circumference Amplitude (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 178, linetype = "dashed", color = "black")+theme_minimal() + ylim(0,400) +xlim(150,202)
jf_309amp
ggsave(filename = )


# 310
daily_stats310 <- daily.data(df = jf_d310_hourly_noNA, TreeNum = 1)
daily_stats310$yr <- format(daily_stats310$DATE, "%Y")
daily_stats310$yr <- as.factor(daily_stats310$yr)
daily_stats310$DOY <- yday(daily_stats310$DATE)
daily_stats310$amplitude_signed <- ifelse(
  daily_stats310$Remarks == "*",
  abs(daily_stats310$amplitude),
  -abs(daily_stats310$amplitude)
)
daily_stats310$ID<-as.factor("310")
daily_stats310$AgeClass<-as.factor("OG")
daily_stats310$Species<-as.factor("DF")
save(daily_stats310,file = here(
  "DendroAnalyst/daily_stats310.RData"))

jf_310amp <- ggplot(
  daily_stats310, aes(x = DOY, y = amplitude, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 310) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,400) +
  xlim(150, 202)
jf_310amp

# 311
daily_stats311 <- daily.data(df = jf_d311_hourly_noNA, TreeNum = 1)
daily_stats311$yr <- format(daily_stats311$DATE, "%Y")
daily_stats311$yr <- as.factor(daily_stats311$yr)
daily_stats311$DOY <- yday(daily_stats311$DATE)
daily_stats311$amplitude_signed <- ifelse(
  daily_stats311$Remarks == "*",
  abs(daily_stats311$amplitude),
  -abs(daily_stats311$amplitude)
)
daily_stats311$ID<-as.factor("311")
daily_stats311$AgeClass<-as.factor("OG")
daily_stats311$Species<-as.factor("DF")
save(daily_stats311,file = here(
  "DendroAnalyst/daily_stats311.RData"))

jf_311amp <- ggplot(
  daily_stats311, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 311) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(-300,300) +
  xlim(150, 202)
jf_311amp

# 646
daily_stats646 <- daily.data(df = jf_d646_hourly_noNA, TreeNum = 1)
daily_stats646$yr <- format(daily_stats646$DATE, "%Y")
daily_stats646$yr <- as.factor(daily_stats646$yr)
daily_stats646$DOY <- yday(daily_stats646$DATE)
daily_stats646$amplitude_signed <- ifelse(
  daily_stats646$Remarks == "*",
  abs(daily_stats646$amplitude),
  -abs(daily_stats646$amplitude)
)
daily_stats646$ID<-as.factor("646")
daily_stats646$AgeClass<-as.factor("OG")
daily_stats646$Species<-as.factor("WH")
save(daily_stats646,file = here(
  "DendroAnalyst/daily_stats646.RData"))


jf_646amp <- ggplot(
  daily_stats646, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 646) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,400) +
  xlim(150, 202)
jf_646amp


# 648
daily_stats648 <- daily.data(df = jf_d648_hourly_noNA, TreeNum = 1)
daily_stats648$yr <- format(daily_stats648$DATE, "%Y")
daily_stats648$yr <- as.factor(daily_stats648$yr)
daily_stats648$DOY <- yday(daily_stats648$DATE)
daily_stats648$amplitude_signed <- ifelse(
  daily_stats648$Remarks == "*",
  abs(daily_stats648$amplitude),
  -abs(daily_stats648$amplitude))
daily_stats648$ID<-as.factor("648")
daily_stats648$AgeClass<-as.factor("OG")
daily_stats648$Species<-as.factor("WH")
save(daily_stats648,file = here(
  "DendroAnalyst/daily_stats648.RData"))

jf_648amp <- ggplot(
  daily_stats648, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 648) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,400) +
  xlim(150, 202)
jf_648amp

# 651
daily_stats651 <- daily.data(df = jf_d651_hourly_noNA, TreeNum = 1)
daily_stats651$yr <- format(daily_stats651$DATE, "%Y")
daily_stats651$yr <- as.factor(daily_stats651$yr)
daily_stats651$DOY <- yday(daily_stats651$DATE)
daily_stats651$amplitude_signed <- ifelse(
  daily_stats651$Remarks == "*",
  abs(daily_stats651$amplitude),
  -abs(daily_stats651$amplitude))
daily_stats651$ID<-as.factor("651")
daily_stats651$AgeClass<-as.factor("OG")
daily_stats651$Species<-as.factor("WRC")
save(daily_stats651,file = here(
  "DendroAnalyst/daily_stats651.RData"))


jf_651amp <- ggplot(
  daily_stats651, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 651) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,400) +
  xlim(150, 202)
jf_651amp

# 652
daily_stats652 <- daily.data(df = jf_d652_hourly_noNA, TreeNum = 1)
daily_stats652$yr <- format(daily_stats652$DATE, "%Y")
daily_stats652$yr <- as.factor(daily_stats652$yr)
daily_stats652$DOY <- yday(daily_stats652$DATE)
daily_stats652$amplitude_signed <- ifelse(
  daily_stats652$Remarks == "*",
  abs(daily_stats652$amplitude),
  -abs(daily_stats652$amplitude))
daily_stats652$ID<-as.factor("652")
daily_stats652$AgeClass<-as.factor("OG")
daily_stats652$Species<-as.factor("WH")
save(daily_stats652,file = here(
  "DendroAnalyst/daily_stats652.RData"))

jf_652amp <- ggplot(
  daily_stats652, aes(x = DOY, y = amplitude, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 652) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,400) +
  xlim(150, 202)
jf_652amp

# 654
daily_stats654 <- daily.data(df = jf_d654_hourly_noNA, TreeNum = 1)
daily_stats654$yr <- format(daily_stats654$DATE, "%Y")
daily_stats654$yr <- as.factor(daily_stats654$yr)
daily_stats654$DOY <- yday(daily_stats654$DATE)
daily_stats654$amplitude_signed <- ifelse(
  daily_stats654$Remarks == "*",
  abs(daily_stats654$amplitude),
  -abs(daily_stats654$amplitude))
daily_stats654$ID<-as.factor("654")
daily_stats654$AgeClass<-as.factor("Secondary")
daily_stats654$Species<-as.factor("DF")
save(daily_stats654,file = here(
  "DendroAnalyst/daily_stats654.RData"))

jf_654amp <- ggplot(
  daily_stats654, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 654) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,750) +
  xlim(150, 202)
jf_654amp

# 655
daily_stats655 <- daily.data(df = jf_d655_hourly_noNA, TreeNum = 1)
daily_stats655$yr <- format(daily_stats655$DATE, "%Y")
daily_stats655$yr <- as.factor(daily_stats655$yr)
daily_stats655$DOY <- yday(daily_stats655$DATE)
daily_stats655$amplitude_signed <- ifelse(
  daily_stats655$Remarks == "*",
  abs(daily_stats655$amplitude),
  -abs(daily_stats655$amplitude))
daily_stats655$ID<-as.factor("655")
daily_stats655$AgeClass<-as.factor("Secondary")
daily_stats655$Species<-as.factor("DF")
save(daily_stats655,file = here(
  "DendroAnalyst/daily_stats655.RData"))

jf_655amp <- ggplot(
  daily_stats655, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 655) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(-400,400) +
  xlim(150, 202)
jf_655amp

# 656
daily_stats656 <- daily.data(df = jf_d656_hourly_noNA, TreeNum = 1)
daily_stats656$yr <- format(daily_stats656$DATE, "%Y")
daily_stats656$yr <- as.factor(daily_stats656$yr)
daily_stats656$DOY <- yday(daily_stats656$DATE)
daily_stats656$amplitude_signed <- ifelse(
  daily_stats656$Remarks == "*",
  abs(daily_stats656$amplitude),
  -abs(daily_stats656$amplitude))
daily_stats656$ID<-as.factor("656")
daily_stats656$AgeClass<-as.factor("Secondary")
daily_stats656$Species<-as.factor("DF")
save(daily_stats656,file = here(
  "DendroAnalyst/daily_stats656.RData"))

jf_656amp <- ggplot(
  daily_stats656, aes(x = DOY, y = amplitude, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 656) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(0,400) +
  xlim(150, 202)
jf_656amp

# 860
daily_stats860 <- daily.data(df = jf_d860_hourly_noNA, TreeNum = 1)
daily_stats860$yr <- format(daily_stats860$DATE, "%Y")
daily_stats860$yr <- as.factor(daily_stats860$yr)
daily_stats860$DOY <- yday(daily_stats860$DATE)
daily_stats860$amplitude_signed <- ifelse(
  daily_stats860$Remarks == "*",
  abs(daily_stats860$amplitude),
  -abs(daily_stats860$amplitude))
daily_stats860$ID<-as.factor("860")
daily_stats860$AgeClass<-as.factor("Secondary")
daily_stats860$Species<-as.factor("DF")
save(daily_stats860,file = here(
  "DendroAnalyst/daily_stats860.RData"))

jf_860amp <- ggplot(
  daily_stats860, aes(x = DOY, y = amplitude_signed, color = factor(yr))) +
  geom_point() +
  labs(
    title = "HJA Douglas-fir (ID 860) Daily Circumference Amplitude",
    x = "Day of Year",
    y = "Daily Stem Circumference Amplitude (µm)",
    color = "Legend"
  ) +
  scale_color_manual(values = custom_colors) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
  theme_minimal() +
  ylim(-400,400) +
  xlim(150, 202)
jf_860amp

daily_stats_6dendros<-bind_rows(
  daily_stats310,
  daily_stats311,
  daily_stats654,
  daily_stats655,
  daily_stats656,
  daily_stats860)

daily_stats_11dendros<-bind_rows(
  daily_stats309,
  daily_stats310,
  daily_stats311,
  daily_stats646,
  daily_stats648,
  daily_stats651,
  daily_stats652,
  daily_stats654,
  daily_stats655,
  daily_stats656,
  daily_stats860)

save(daily_stats_6dendros,file=here("DendroAnalyst/daily_stats_6dendros.RData"))
save(daily_stats_11dendros,file=here("DendroAnalyst/daily_stats_11dendros.RData"))
```

# Modeling annual growth with a gompertz function

```         
dm.fit.gompertz(df, CalYear, TreeNum, f_derivative = F)  gomp_fitted<-dm.fit.gompertz(df=gf_nepa17, TreeNum = 1, CalYear=2017)
```

### Explanation of options

+-----------------------------------+------------------------------------------------------------------------------------------------------------------------+
| `CalYear`                         | numeric for year of calculation. If df has more than one year, assigning CalYear truncates the data of only that year. |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------+
| `f_derivative`                    | logical if yes returns first derivative of gompertz curve.                                                             |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------------+

`{r} jump_free_d309_noNA<-jump_free_d309|>filter(!is.na(jump_free_d309$value)) gompertz309<-dm.fit.gompertz(df=jump_free_d309_noNA,TreeNum=1,CalYear = 2021) ggplot(gompertz309,aes(x=TIME,y=value))+geom_point()}`

# Step 4: Analyses with Climate Data

```{r}
#Apr1-Oct1 D309
jump_free_d309$yr<-format(jump_free_d309$TIME, "%Y")
jump_free_d309$yr<-as.factor(jump_free_d309$yr)
jump_free_d309$DOY<-yday(jump_free_d309$TIME)
ggplot(
  jump_free_d309,aes(x=DOY,y=value,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 309) Adjusted Stem Circumference Change",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal()+ylim(-2000,25500)#+xlim(91,274)
```

## Running correlation between dendrometer and climate data

Calculates running correlation between dendrometer series and provided climate variables. Users can choose methods such as pearson, kendall and spearman, see `?cor.test` for further information.

```         
mov.cor.dm(df, Clim, TreeNum, win_size, cor_method = "pearson") 
out_corr<-mov.cor.dm(df=gf_nepa17, Clim=ktm_rain17, TreeNum=1, win_size=21) 
```

### Explanation of options

+-----------------------------------+------------------------------------------------------------------------------------------------------------------+
| `win_size`                        | numerical, the running days windows of which the correlation is to be calculated. The minimum value is 18.       |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------+
| `cor_method`                      | string, method to be applied during correlation calculation. One of the following: pearson, kendall and spearman |
+-----------------------------------+------------------------------------------------------------------------------------------------------------------+

### Output can be plotted with the following:

plot_mov.cor(mov.cor.output, sig.only = T, ci = 0.95, clim_vars = "all")\
Example code:\
out_corr\<-mov.cor.dm(df=gf_nepa17, Clim=ktm_rain17, TreeNum=1, win_size=21) plot_mov.cor(mov.cor.output=out_corr, sig.only=TRUE, ci=0.95)

+-----------------------------------+----------------------------------------------------------------+
| `mov.cor.output`                  | list the output of `mov.cor.dm` function.                      |
+-----------------------------------+----------------------------------------------------------------+
| `sig.only`                        | logical TRUE to plot only significant correlation.             |
+-----------------------------------+----------------------------------------------------------------+
| `ci`                              | numeric confidence interval                                    |
+-----------------------------------+----------------------------------------------------------------+
| `clim_vars`                       | array of climate variables or "all" for all climate variables. |
+-----------------------------------+----------------------------------------------------------------+

### Moving correlation of climate

```{r}
#consider hourly?
names(filtered_Dscmet_AT52m_daily)[names(filtered_Dscmet_AT52m_daily) == "day"] <- "TIME"
names(filtered_Dscmet_VPD52m_daily)[names(filtered_Dscmet_VPD52m_daily) == "day"] <- "TIME"

filtered_Dscmet_AT52m_daily$TIME <- as.Date(filtered_Dscmet_AT52m_daily$TIME)
filtered_Dscmet_VPD52m_daily$TIME <- as.Date(filtered_Dscmet_VPD52m_daily$TIME)

#309 
jf_d309_daily$TIME <- as.Date(
  jf_d309_daily$TIME)
#daily mean AT
jf_d309_meanAT_daily_cor <- mov.cor.dm(
  jf_d309_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d309_VPD_daily_cor <- mov.cor.dm(
  jf_d309_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily max HDI
# Flag long NA sequences
r <- rle::rle(is.na(Dscmet_HDI_daily_max$max_HDI))
na_runs <- which(r$values & r$lengths >= 10)  # example: NA runs ≥ 10 days
na_idx <- unlist(mapply(seq, 
                        cumsum(c(1, head(r$lengths, -1)))[na_runs], 
                        cumsum(r$lengths)[na_runs]))

# Set those to NA again to avoid interpolation there
Dscmet_HDI_daily_max$max_HDI[na_idx] <- NA

jf_d309_max_HDI_daily_cor <- mov.cor.dm(
  jf_d309_daily, 
  Dscmet_HDI_daily_max, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#310 
jf_d310_daily$TIME <- as.Date(
  jf_d310_daily$TIME)
#daily mean AT
jf_d310_meanAT_daily_cor <- mov.cor.dm(
  jf_d310_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d310_VPD_daily_cor <- mov.cor.dm(
  jf_d310_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#311 
jf_d311_daily$TIME <- as.Date(
  jf_d311_daily$TIME)
#daily mean AT
jf_d311_meanAT_daily_cor <- mov.cor.dm(
  jf_d311_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d311_VPD_daily_cor <- mov.cor.dm(
  jf_d311_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#646 
jf_d646_daily$TIME <- as.Date(
  jf_d646_daily$TIME)
#daily mean AT
jf_d646_meanAT_daily_cor <- mov.cor.dm(
  jf_d646_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d646_VPD_daily_cor <- mov.cor.dm(
  jf_d646_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#648 - "error: not enough finite observations"
jf_d648_daily$TIME <- as.Date(
  jf_d648_daily$TIME)
#daily mean AT - 
#jf_d648_meanAT_daily_cor <- mov.cor.dm(
#  jf_d648_daily, 
#  filtered_Dscmet_AT52m_daily, 
#  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
#jf_d648_VPD_daily_cor <- mov.cor.dm(
#  jf_d648_daily, 
#  filtered_Dscmet_VPD52m_daily, 
#  TreeNum=1, win_size=18, cor_method = "pearson")

#651 - "error: not enough finite observations"
#jf_d651_daily$TIME <- as.Date(
#  jf_d651_daily$TIME)
#daily mean AT
#jf_d651_meanAT_daily_cor <- mov.cor.dm(
#  jf_d651_daily, 
#  filtered_Dscmet_AT52m_daily, 
#  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
#jf_d651_VPD_daily_cor <- mov.cor.dm(
#  jf_d651_daily, 
#  filtered_Dscmet_VPD52m_daily, 
#  TreeNum=1, win_size=18, cor_method = "pearson")

#652  - "error: not enough finite observations"
#jf_d652_daily$TIME <- as.Date(
#  jf_d652_daily$TIME)
#daily mean AT
#jf_d652_meanAT_daily_cor <- mov.cor.dm(
#  jf_d652_daily, 
#  filtered_Dscmet_AT52m_daily, 
#  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
#jf_d652_VPD_daily_cor <- mov.cor.dm(
#  jf_d652_daily, 
#  filtered_Dscmet_VPD52m_daily, 
#  TreeNum=1, win_size=18, cor_method = "pearson")


#654 
jf_d654_daily$TIME <- as.Date(
  jf_d654_daily$TIME)
#daily mean AT
jf_d654_meanAT_daily_cor <- mov.cor.dm(
  jf_d654_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d654_VPD_daily_cor <- mov.cor.dm(
  jf_d654_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#655 
jf_d655_daily$TIME <- as.Date(
  jf_d655_daily$TIME)
#daily mean AT
jf_d655_meanAT_daily_cor <- mov.cor.dm(
  jf_d655_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d310_VPD_daily_cor <- mov.cor.dm(
  jf_d310_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#656 
jf_d656_daily$TIME <- as.Date(
  jf_d656_daily$TIME)
#daily mean AT
jf_d656_meanAT_daily_cor <- mov.cor.dm(
  jf_d656_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d656_VPD_daily_cor <- mov.cor.dm(
  jf_d656_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

#860 
jf_d860_daily$TIME <- as.Date(
  jf_d860_daily$TIME)
#daily mean AT
jf_d860_meanAT_daily_cor <- mov.cor.dm(
  jf_d860_daily, 
  filtered_Dscmet_AT52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")
#daily mean VPD
jf_d860_VPD_daily_cor <- mov.cor.dm(
  jf_d860_daily, 
  filtered_Dscmet_VPD52m_daily, 
  TreeNum=1, win_size=18, cor_method = "pearson")

plot_mov.cor(jf_d860_VPD_daily_cor)
```

```{r}
start_time <- as.POSIXct("2019-01-01 08:00:00")
end_time   <- as.POSIXct("2024-09-30 17:00:00")  # adjust this to your actual range

mov.cor.dm.hourly <- function (df, Clim, TreeNum, win_size, cor_method = "pearson") {
    if (!(cor_method %in% c("pearson", "kendall", "spearman"))) 
        if (!inherits(df[[1]], "Date") && !inherits(df[[1]], 
            "POSIXct")) {
            df[[1]] <- ymd_hms(df[[1]])
        }
    if (!inherits(Clim[[1]], "Date") && !inherits(Clim[[1]], 
        "POSIXct")) {
        Clim[[1]] <- ymd(Clim[[1]])
    }
    TIME <- NULL
    df <- dendro.resample(df = df, by = "H", value = "mean") %>% 
        mutate(TIME = as.Date(TIME))
    Clim <- as_tibble(Clim) %>% rename(TIME = 1)
    x <- intersect(Clim$TIME, df$TIME)
    if (length(x) == 0) {
        stop("'df' and 'Clim' must have time overlap.")
    }
    else {
        if ((length(x) == nrow(df)) & (length(x) == nrow(Clim))) {
            df <- df
            Clim <- Clim
        }
        else {
            warning("The temporal coverage of 'df' and 'Clim' is not same. The output data will be shortened to the common time period.")
            # 🔥 Fix is here: we skip or fix this line
            # x <- as.Date(x, origin = "1970-01-01", tz = "UTC")
            x <- as.Date(x)
            df <- df %>% filter(TIME %in% x)
            Clim <- Clim %>% filter(TIME %in% x)
        }
    }
    wl <- ifelse(win_size%%2 == 0, win_size + 1, win_size)
    dm_col <- pull(df, TreeNum + 1)
    n_row <- nrow(Clim)
    wl2 <- (wl - 1)
    wl3 <- wl2/2
    out <- lapply(Clim[-1], function(cls) {
        estimate <- as.data.frame(matrix(ncol = 3, nrow = n_row))
        estimate[, 1] <- as.Date(x)
        for (n in seq(wl, n_row)) {
            val_1 <- cls[(n - wl2):n]
            val_2 <- dm_col[(n - wl2):n]
            a = cor.test(val_1, val_2, method = cor_method)
            estimate[n - wl3, 2] <- a$estimate
            estimate[n - wl3, 3] <- a$p.value
        }
        colnames(estimate) = c("TIME", "corr", "p_val")
        return(as_tibble(estimate))
    })
    class(out) <- "mov.cor"
    return(out)
}









#310 daily
names(filtered_Dscmet_maxVPD52m_daily)[names(filtered_Dscmet_maxVPD52m_daily) == "day"] <- "TIME"
names(filtered_Dscmet_maxVPD52m_hourly)[names(filtered_Dscmet_maxVPD52m_hourly) == "hour"] <- "TIME"

filtered_Dscmet_maxVPD52m_daily$TIME <- as.POSIXct(filtered_Dscmet_maxVPD52m_daily$TIME, format = "%Y-%m-%d %H:%M:%S",tz="America/Los_Angeles")
filtered_Dscmet_maxVPD52m_hourly$TIME <- as.POSIXct(filtered_Dscmet_maxVPD52m_hourly$TIME, format = "%Y-%m-%d %H:%M:%S",tz="America/Los_Angeles")

jump_free_d310_daily$TIME <- as.POSIXct(jump_free_d310_daily$TIME, format = "%Y-%m-%d %H:%M:%S",tz="America/Los_Angeles")
jump_free_d310$TIME <- as.POSIXct(jump_free_d310$TIME, format = "%Y-%m-%d %H:%M:%S",tz="America/Los_Angeles")

filtered_Dscmet_maxVPD52m_daily <- filtered_Dscmet_maxVPD52m_daily |>
  mutate(across(where(is.numeric), ~na_if(., -Inf)))

filtered_Dscmet_maxVPD52m_hourly <- filtered_Dscmet_maxVPD52m_hourly |>
  mutate(across(where(is.numeric), ~na_if(., -Inf)))

# Filter df and Clim so only rows where Clim's value is NOT NA
common_times_daily <- filtered_Dscmet_maxVPD52m_daily %>%
  filter(!is.na(maxVPD52m_kpa_daily)) %>%
  select(TIME)

common_times <- filtered_Dscmet_maxVPD52m_hourly %>%
  filter(!is.na(maxVPD52m_kpa_hourly)) %>%
  select(TIME)

# Join with df to get matching times with non-NA Clim values
jump_free_d310daily_filtered <- jump_free_d310_daily %>%
  inner_join(common_times, by = "TIME")

jump_free_d310_filtered <- jump_free_d310 %>%
  inner_join(common_times, by = "TIME")

filtered_Clim_daily <- filtered_Dscmet_maxVPD52m_daily %>%
  filter(!is.na(maxVPD52m_kpa_daily))

filtered_Clim_hourly <- filtered_Dscmet_maxVPD52m_hourly %>%
  filter(!is.na(maxVPD52m_kpa_hourly))

valid_times_daily <- jump_free_d310daily_filtered %>%
  filter(!is.na(value)) %>%
  inner_join(filtered_Clim_daily %>% filter(!is.na(maxVPD52m_kpa_daily)), by = "TIME")

valid_times_hourly <- jump_free_d310_filtered %>%
  filter(!is.na(value)) %>%
  inner_join(filtered_Clim_hourly %>% filter(!is.na(maxVPD52m_kpa_hourly)), by = "TIME")


# 2. Combine and resample to daily (averaging hourly data)
jump_free_d310_daily <- valid_times %>%
  mutate(DATE = as.Date(TIME)) %>%
  group_by(DATE) %>%
  summarize(value = mean(value, na.rm = TRUE)) %>%
  rename(TIME = DATE)

filtered_Clim_daily <- valid_times %>%
  mutate(DATE = as.Date(TIME)) %>%
  group_by(DATE) %>%
  summarize(maxVPD52m_kpa_daily = mean(maxVPD52m_kpa_daily, na.rm = TRUE)) %>%
  rename(TIME = DATE)

jump_free_d310_hourly <- valid_times %>%
  mutate(DATE = as.Date(TIME)) %>%
  group_by(DATE) %>%
  summarize(value = mean(value, na.rm = TRUE)) %>%
  rename(TIME = DATE)

filtered_Clim_hourly <- valid_times_hourly %>%
  mutate(DATE = as.Date(TIME)) %>%
  group_by(DATE) %>%
  summarize(maxVPD52m_kpa_hourly = mean(maxVPD52m_kpa_hourly, na.rm = TRUE)) %>%
  rename(TIME = DATE)

# 3. Join on shared dates to ensure perfect alignment
common_dates_daily <- intersect(jump_free_d310_daily$TIME, filtered_Clim_daily$TIME)
common_dates_hourly <- intersect(jump_free_d310_hourly$TIME, filtered_Clim_hourly$TIME)

jump_free_d310daily_final <- jump_free_d310_daily %>% filter(
  TIME %in% common_dates_daily)
filtered_Clim_hourly_final <- filtered_Clim_hourly %>% filter(TIME %in% common_dates_hourly)


#trying chatgpt suggestion on removing flat values 
# Step 1: Compute rolling SD
roll_sd_clim_daily <- rollapply(
  filtered_Dscmet_maxVPD52m_daily$maxVPD52m_kpa_daily,
  width = 18,
  FUN = sd,
  fill = NA,
  align = "center"
)

roll_sd_clim_hourly <- rollapply(
  filtered_Dscmet_maxVPD52m_daily$maxVPD52m_kpa_hourly,
  width = 18,
  FUN = sd,
  fill = NA,
  align = "center"
)

# Step 2: Find indices of all flat windows
flat_window_centers_daily <- which(roll_sd_clim == 0)
flat_window_centers_hourly <- which(roll_sd_clim == 0)

# Step 3: Build full list of dates in any flat window (±8 days from each center)
flat_window_indices_daily <- unlist(
  lapply(flat_window_centers_daily, function(i) (i-8):(i+9)))
flat_window_indices_daily <- unique(
  flat_window_indices_daily[flat_window_indices_daily > 0 & flat_window_indices_daily <= nrow(filtered_Dscmet_maxVPD52m_daily)])

flat_window_indices_hourly <- unlist(
  lapply(flat_window_centers_hourly, function(i) (i-8):(i+9)))
flat_window_indices_hourly <- unique(
  flat_window_indices_hourly[flat_window_indices_hourly > 0 & flat_window_indices_hourly <= nrow(filtered_Dscmet_maxVPD52m_hourly)])

# Step 4: Get all corresponding TIME values
flat_window_dates_daily <- filtered_Dscmet_maxVPD52m_daily$TIME[flat_window_indices_daily]

flat_window_dates_hourly <- filtered_Dscmet_maxVPD52m_hourly$TIME[flat_window_indices_hourly]

# Step 5: Clean both datasets synchronously
clim_clean_daily <- filtered_Dscmet_maxVPD52m_daily %>%
  filter(!TIME %in% flat_window_dates_daily)

df_clean_daily <- jump_free_d310_daily %>%
  filter(TIME %in% clim_clean_daily$TIME)

clim_clean_hourly <- filtered_Dscmet_maxVPD52m_hourly %>%
  filter(!TIME %in% flat_window_dates_hourly)

df_clean_hourly <- jump_free_d310_hourly %>%
  filter(TIME %in% clim_clean_hourly$TIME)


cor310_daily <- mov.cor.dm(
  df = df_clean_daily,
  Clim = clim_clean_daily,
  TreeNum = 1,
  win_size = 18
)

cor310_hourly <- mov.cor.dm.hourly(
  df = jump_free_d310_hourly,
  Clim = filtered_Dscmet_maxVPD52m_hourly,
  TreeNum = 1,
  win_size = 18
)

#plot_mov.cor(H310, sig.only = T, ci = 0.95, clim_vars = "all")
```

## Relative Dry Growth During Climate Adverse events with clim.twd

```         
clim.twd(df,   Clim,   dailyValue = "max",   thresholdClim = "<10",   thresholdDays = ">5",   showPlot = TRUE ) 
```

`{r} #309   #310  #daily with max mean vpd d310_RelGrowth_max3<-clim.twd(   df=jf_d310_daily_noNA, Clim=filtered_Dscmet_maxVPD52m_daily,    dailyValue='max', thresholdClim = ">=3", thresholdDays = ">0", showPlot=TRUE) #daily with max AT d310_RelGrowth_max35<-clim.twd(   df=jf_d310_daily_noNA, Clim=filtered_Dscmet_maxAT52m_daily,    dailyValue='max', thresholdClim = ">=35", thresholdDays = ">1", showPlot=TRUE)}`

### Explanation of options

+-----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `Clim`                            | dataframe with the first column containing `Date in yyyy-mm-dd` and second column containing corresponding climate data.                                                                                                                                |
+-----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `dailyValue`                      | either *'max', 'min', 'mean'*, or *'sum'* for selecting the daily resampled value. Default is *'max'*. See [`dendro.resample`](http://127.0.0.1:17265/help/library/dendRoAnalyst/help/dendro.resample) for details.                                     |
+-----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `thresholdClim`                   | string, the theshold for the respective climatic parameter. E.g. if climatic data is precipitation then days, where precipitation is below or equal to this value, are considered as adverse climate. Default is '\<10'.                                |
+-----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `thresholdDays`                   | string, the minimum number of consecutive adverse days to be considered for analysis. For example, `thresholdDays`=2 means the relative radial/circumferential change is calculated for adverse periods lasting for more than 2 days. Default is '\>5'. |
+-----------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
names(filtered_Dscmet_maxVPD52m_daily)[names(filtered_Dscmet_maxVPD52m_daily) == "day"] <- "TIME"
filtered_Dscmet_maxVPD52m_daily$TIME<-as.Date(
  filtered_Dscmet_maxVPD52m_daily$TIME)

names(filtered_Dscmet_maxAT52m_daily)[names(filtered_Dscmet_maxAT52m_daily) == "day"] <- "TIME"
filtered_Dscmet_maxAT52m_daily$TIME<-as.Date(
  filtered_Dscmet_maxAT52m_daily$TIME)

names(Dscmet_HDI_daily_max)[names(Dscmet_HDI_daily_max) == "day"] <- "TIME"
Dscmet_HDI_daily_max$TIME<-as.Date(
  Dscmet_HDI_daily_max$TIME)

#309
#max HDI daily
#rel_dry_d309_maxHDI<-clim.twd(
#  df=jf_d309_daily_noNA,  
#  Clim=Dscmet_HDI_daily_max,    
#  dailyValue='max',    
#  thresholdClim = ">3",   thresholdDays = ">1",   showPlot=TRUE) 

#max VPD
rel_dry_d309_maxVPD<-clim.twd(
  df=jf_d309_daily_noNA,  
  Clim=filtered_Dscmet_maxVPD52m_daily,    
  dailyValue='max',    
  thresholdClim = ">3",   thresholdDays = ">1",   showPlot=TRUE) 

#max AT 
rel_dry_d309_maxAT<-clim.twd(
  df=jf_d309_daily_noNA,  
  Clim=filtered_Dscmet_maxAT52m_daily,    
  dailyValue='max',    
  thresholdClim = ">=30",   thresholdDays = ">3",   showPlot=TRUE) 
# from these threshholds, HD = period 10. Each year has these periods per year:
# 2019:-(); 2020:-(); 2021:-(); 2022:-(); 2023:-(); 2024:-()
#max VPD 
rel_dry_d310_maxVPD<-clim.twd(
  df=jf_d310_daily_noNA,  
  Clim=filtered_Dscmet_maxVPD52m_daily,    
  dailyValue='max',    
  thresholdClim = ">3",   thresholdDays = ">3",   showPlot=TRUE) 
# from these threshholds, HD = period 10. Each year has these periods per year:
# 2019:1-4(4); 2020:5-8(4); 2021:9-17(9); 2022:18-23(6); 2023:24-29(6); 2024:30-33(4)

#max AT 
rel_dry_d310_maxAT<-clim.twd(
  df=jf_d310_daily_noNA,  
  Clim=filtered_Dscmet_maxAT52m_daily,    
  dailyValue='max',    
  thresholdClim = ">25",   thresholdDays = ">5",   showPlot=TRUE) 
# from these threshholds, HD = period 10. Each year has these periods per year:
# 2019:-(); 2020:-(); 2021:-(); 2022:-(); 2023:-(); 2024:-()


#jump-free309 relative_dry_growth_jf309<-clim.twd(   df=jump_free_d309,    Clim=filtered_P230_rain_daily,    dailyValue='mean',    thresholdClim = "<2.5",   thresholdDays = ">10",   showPlot=TRUE)
```

# Step 5: Make Other Plots

## Daily Max from Step 3.2

DOY 178 = june 27 DOY 179 = june 28\
DOY 178-28 = 150 = May 30\
DOY 179+28 = 207 = July 26

Great loop for making multiple plots with each year starting at 0

```{r}
load(here("DendroAnalyst/daily_stats_11dendros.RData"))
#truncating df to be HD +/- 4 weeks
daily_stats_11dendros_trunc <- daily_stats_11dendros |>
  filter(
    (month(DATE) == 5 & day(DATE) >= 30) |
    (month(DATE) == 6) |
    (month(DATE) == 7 & day(DATE) <= 26))


daily_stats_11dendros_trunc_normalized <- daily_stats_11dendros_trunc |>
  arrange(ID, yr, DOY) |>
  group_by(ID, yr) |>
  mutate(Max_rel = Max - first(Max)) |>
  ungroup()

dendro_ids <- unique(daily_stats_11dendros_trunc_normalized$ID)
plot_list <- list()
for (id in dendro_ids) {
dendro_ids_subset <- daily_stats_11dendros_trunc_normalized |>
    filter(ID == id)

plot_list[[id]] <- ggplot(
  dendro_ids_subset, aes(x = DOY, y = Max_rel, color = yr)) +
    geom_line(size = 2,alpha=0.65) +
    labs(
      title = paste(
      "HJA DT Daily Max Dendrometer Value (relative to DOY 150) ID:", id),
      x = "Day of Year",
      y = "Relative Daily Maximum Dendrometer Value (µm)",
      color = "Year"
    ) +
    scale_color_manual(values = custom_colors) +
    geom_vline(xintercept = 178, linetype = "dashed", color = "black") +
    theme_minimal() +
    theme(legend.position = "bottom")
}

plot_list[["309"]]  
plot_list[["310"]]  
plot_list[["311"]]  
plot_list[["646"]]  
plot_list[["648"]]  
plot_list[["651"]]  
plot_list[["652"]]  
plot_list[["654"]]  
plot_list[["655"]]  
plot_list[["656"]]  
plot_list[["860"]]  

#saving plots
plot_folder <- here("Rel_Daily_Max/")
for (id in dendro_ids) {
  plot_file <- paste0(plot_folder, "daily_max_", id, ".png")
  ggsave(
    filename = plot_file,
    plot = plot_list[[id]],
  )}
```

```{r}

# DC3 Ecomatik Band Dendrometer Conversion Function
# Converts raw dendrometer output to tree circumference in micrometers
# Based on Excel implementation that differs from manual equation

convert_dendrometer_to_micrometers <- function(Vi_micrometers, C0, V0 = NULL, use_excel_method = TRUE) {
  # Parameters:
  # Vi_micrometers: Vector of dendrometer data records in micrometers (raw data)
  # C0: Initial circumference of tree at dendrometer height in mm
  # V0: First valid record of dendrometer in mm (optional, defaults to first Vi value converted to mm)
  # use_excel_method: If TRUE, uses Excel formula; if FALSE, uses manual equation
  
  # Input validation
  if (missing(Vi_micrometers) || missing(C0)) {
    stop("Both Vi_micrometers (dendrometer readings) and C0 (initial circumference) are required")
  }
  
  if (length(Vi_micrometers) == 0) {
    stop("Vi_micrometers must contain at least one measurement")
  }
  
  if (C0 <= 0) {
    stop("C0 (initial circumference) must be positive")
  }
  
  # Convert micrometers to mm for calculations
  Vi_mm <- Vi_micrometers / 1000
  
  # Set V0 to first measurement if not provided
  if (is.null(V0)) {
    V0 <- Vi_mm[1]
    message("V0 not provided, using first measurement: ", V0, " mm")
  }
  
  # Calculate initial radius R0 from initial circumference
  R0 <- C0 / (pi * 2)  # R0 = C0 / (π * 2)
  
  # Constants
  Dini <- 128  # Initial diameter constant from Excel
  
  # Calculate Cwire using Excel formula: C0 + 2*sqrt((Dini+R0-V0)^2-R0^2) - 2*R0*ACOS(R0/(Dini+R0-V0))
  sqrt_term_cwire <- sqrt((Dini + R0 - V0)^2 - R0^2)
  acos_term_cwire <- acos(R0 / (Dini + R0 - V0))
  Cwire <- C0 + 2 * sqrt_term_cwire - 2 * R0 * acos_term_cwire
  
  # Initialize output vectors
  Ri <- numeric(length(Vi_mm))
  Ci <- numeric(length(Vi_mm))
  
  if (use_excel_method) {
    # Excel method: Ri (mm) = (Cwire - 2*sqrt((Dini+Ri_prev-Vi)^2-Ri_prev^2) + 2*Ri_prev*ACOS(Ri_prev/(Dini+Ri_prev-Vi)))/pi/2
    # Where Cwire = G6, Dini = G7 (128), Ri_prev = previous D value
    
    for (i in seq_along(Vi_mm)) {
      if (i == 1) {
        # For first measurement, use R0 as previous radius (D4 = R0)
        Ri_prev <- R0
      } else {
        # For subsequent measurements, use previous calculated radius (D4 = previous Ri)
        Ri_prev <- Ri[i-1]
      }
      
      # Excel formula: (Cwire - 2*sqrt((Dini+Ri_prev-Vi)^2-Ri_prev^2) + 2*Ri_prev*ACOS(Ri_prev/(Dini+Ri_prev-Vi)))/π/2
      # where Ri_prev = D4 (previous row's Ri value)
      sqrt_term <- sqrt((Dini + Ri_prev - Vi_mm[i])^2 - Ri_prev^2)
      acos_term <- acos(Ri_prev / (Dini + Ri_prev - Vi_mm[i]))
      
      Ri[i] <- (Cwire - 2 * sqrt_term + 2 * Ri_prev * acos_term) / pi / 2
      Ci[i] <- Ri[i] * pi * 2
    }
    
  } else {
    # Manual equation method (original implementation)
    for (i in seq_along(Vi_mm)) {
      if (i == 1) {
        Ri_prev <- R0
      } else {
        Ri_prev <- Ci[i-1] / (pi * 2)
      }
      
      # First term: R0 component
      term1_arccos <- acos(R0 / (R0 + 128 - V0))
      term1_sqrt <- sqrt((R0 + 128 - V0)^2 - R0^2)
      term1 <- R0 * term1_arccos + term1_sqrt
      
      # Second term: Ri-1 component  
      term2_arccos <- acos(Ri_prev / (Ri_prev + 128 - Vi_mm[i]))
      term2_sqrt <- sqrt((Ri_prev + 128 - Vi_mm[i])^2 - Ri_prev^2)
      term2 <- Ri_prev * term2_arccos - term2_sqrt
      
      # Calculate circumference Ci
      Ci[i] <- 2 * (pi * R0 - term1 + term2)
      Ri[i] <- Ci[i] / (pi * 2)
    }
  }
  
  # Convert to micrometers
  Ci_micrometers <- Ci * 1000
  
  # Create result data frame
  result <- data.frame(
    measurement = seq_along(Vi_micrometers),
    Vi_micrometers = Vi_micrometers,
    Vi_mm = Vi_mm,
    Ri_mm = Ri,
    circumference_mm = Ci,
    circumference_micrometers = Ci_micrometers,
    change_from_initial_micrometers = (Ci - C0) * 1000
  )
  
  return(result)
}

# Example usage function
example_usage <- function() {
  # Example data matching your Excel setup
  C0 <- 790.00000  # Initial circumference in mm
  Vi_micrometers <- c(6014.50, 6100, 6200, 6300, 6250)  # Raw dendrometer readings in micrometers
  V0 <- 6.01450  # First valid record in mm (Vi_micrometers[1]/1000)
  
  # Convert using Excel method
  result_excel <- convert_dendrometer_to_micrometers(Vi_micrometers, C0, V0, use_excel_method = TRUE)
  
  # Convert using manual method
  result_manual <- convert_dendrometer_to_micrometers(Vi_micrometers, C0, V0, use_excel_method = FALSE)
  
  cat("Excel Method Results:\n")
  print(result_excel)
  
  cat("\n\nManual Method Results:\n")
  print(result_manual)
  
  # Compare the methods
  cat("\n\nComparison of methods (circumference in mm):\n")
  comparison <- data.frame(
    measurement = result_excel$measurement,
    Vi_micrometers = result_excel$Vi_micrometers,
    Excel_method = result_excel$circumference_mm,
    Manual_method = result_manual$circumference_mm,
    Difference = result_excel$circumference_mm - result_manual$circumference_mm
  )
  print(comparison)
}

# Test with your specific values
test_your_values <- function() {
  # Your specific values
  C0 <- 790.00000  # G3 - Initial circumference
  R0 <- 125.73251  # G4 - Initial radius (should equal C0/(pi*2))
  V0 <- 6.01450    # G5 - First valid reading in mm
  Dini <- 128.00000 # G7 - Initial diameter constant
  
  # Calculate R0 to verify
  R0_calculated <- C0 / (pi * 2)
  
  cat("=== SETUP VALUES ===\n")
  cat("C0 (G3 - initial circumference):", C0, "mm\n")
  cat("R0 (G4 - initial radius):", R0, "mm\n")
  cat("R0 calculated from C0:", R0_calculated, "mm\n")
  cat("Match:", abs(R0 - R0_calculated) < 0.00001, "\n")
  cat("V0 (G5 - first valid reading):", V0, "mm\n")
  cat("Dini (G7 - diameter constant):", Dini, "mm\n\n")
  
  # Calculate Cwire using Excel formula
  # Cwire = C0 + 2*sqrt((Dini+R0-V0)^2-R0^2) - 2*R0*ACOS(R0/(Dini+R0-V0))
  sqrt_term_cwire <- sqrt((Dini + R0 - V0)^2 - R0^2)
  acos_term_cwire <- acos(R0 / (Dini + R0 - V0))
  Cwire_calculated <- C0 + 2 * sqrt_term_cwire - 2 * R0 * acos_term_cwire
  
  cat("=== CWIRE CALCULATION ===\n")
  cat("Cwire formula components:\n")
  cat("  Dini + R0 - V0 =", Dini + R0 - V0, "\n")
  cat("  sqrt((Dini+R0-V0)^2-R0^2) =", sqrt_term_cwire, "\n")
  cat("  ACOS(R0/(Dini+R0-V0)) =", acos_term_cwire, "radians\n")
  cat("Cwire calculated:", Cwire_calculated, "mm\n")
  cat("Cwire from Excel (G6):", 955.74301, "mm\n")
  cat("Match:", abs(Cwire_calculated - 955.74301) < 0.00001, "\n\n")
  
  # Test with some example readings
  Vi_micrometers <- c(6014.50, 6020.00, 6025.50, 6030.00, 6027.25)
  
  # Convert using the function
  result <- convert_dendrometer_to_micrometers(Vi_micrometers, C0, V0, use_excel_method = TRUE)
  
  cat("=== FUNCTION RESULTS ===\n")
  print(result)
  
  # Manual verification of first measurement
  cat("\n=== MANUAL VERIFICATION OF FIRST MEASUREMENT ===\n")
  Vi_mm_first <- Vi_micrometers[1] / 1000
  Ri_prev <- R0  # For first calculation
  
  cat("Vi (first measurement):", Vi_mm_first, "mm\n")
  cat("Should equal V0:", V0, "mm\n")
  cat("Match:", abs(Vi_mm_first - V0) < 0.00001, "\n")
  
  # Apply Excel formula for first measurement
  denominator <- Dini + Ri_prev - Vi_mm_first
  sqrt_term <- sqrt(denominator^2 - Ri_prev^2)
  acos_term <- acos(Ri_prev / denominator)
  
  Ri_calculated <- (Cwire_calculated - 2 * sqrt_term + 2 * Ri_prev * acos_term) / pi / 2
  Ci_calculated <- Ri_calculated * pi * 2
  
  cat("\nExcel formula calculation:\n")
  cat("Ri calculated:", Ri_calculated, "mm\n")
  cat("Ci calculated:", Ci_calculated, "mm\n")
  cat("Change from initial:", Ci_calculated - C0, "mm\n")
  cat("Change in micrometers:", (Ci_calculated - C0) * 1000, "μm\n")
  
  return(result)
}

# Run test with your actual data - corrected version
run_your_data_corrected <- function() {
  # Your specific values
  C0 <- 790.00000  # G3 - Initial circumference
  V0 <- 6.01450    # G5 - First valid reading in mm
  
  # Your actual dendrometer readings in micrometers
  Vi_micrometers <- c(6014.50, 5996.10, 5989.20, 5911.00, 5851.20)
  
  # Expected Excel results
  expected_Ci <- c(790.000, 789.968, 789.955, 789.819, 789.708)
  
  cat("=== RUNNING WITH YOUR ACTUAL DATA (CORRECTED) ===\n")
  cat("C0 (initial circumference):", C0, "mm\n")
  cat("V0 (first valid reading):", V0, "mm\n")
  cat("Dendrometer readings (micrometers):", paste(Vi_micrometers, collapse = ", "), "\n")
  cat("Expected Excel Ci values:", paste(expected_Ci, collapse = ", "), "\n\n")
  
  # Test the function
  result <- convert_dendrometer_to_micrometers(Vi_micrometers, C0, V0, use_excel_method = TRUE)
  
  cat("=== FUNCTION RESULTS ===\n")
  print(result)
  
  # Compare with expected values
  cat("\n=== COMPARISON WITH EXPECTED EXCEL VALUES ===\n")
  comparison <- data.frame(
    Measurement = 1:length(Vi_micrometers),
    Vi_micrometers = Vi_micrometers,
    Expected_Ci = expected_Ci,
    Calculated_Ci = round(result$circumference_mm, 3),
    Difference = round(result$circumference_mm - expected_Ci, 6)
  )
  print(comparison)
  
  # Check if differences are small (within rounding error)
  max_diff <- max(abs(comparison$Difference))
  cat("\nMaximum difference:", max_diff, "mm\n")
  cat("Within acceptable range (<0.001 mm):", max_diff < 0.001, "\n")
  
  # Manual step-by-step for first few calculations
  cat("\n=== STEP-BY-STEP VERIFICATION ===\n")
  R0 <- C0 / (pi * 2)
  Dini <- 128
  
  # Calculate Cwire
  sqrt_term_cwire <- sqrt((Dini + R0 - V0)^2 - R0^2)
  acos_term_cwire <- acos(R0 / (Dini + R0 - V0))
  Cwire <- C0 + 2 * sqrt_term_cwire - 2 * R0 * acos_term_cwire
  
  cat("Cwire calculated:", round(Cwire, 5), "mm\n")
  
  # First calculation (i=1): D4 = R0, C5 = Vi[1]/1000
  Vi_1 <- Vi_micrometers[1] / 1000
  D4_1 <- R0  # Previous Ri = R0 for first calculation
  
  sqrt_term_1 <- sqrt((Dini + D4_1 - Vi_1)^2 - D4_1^2)
  acos_term_1 <- acos(D4_1 / (Dini + D4_1 - Vi_1))
  Ri_1 <- (Cwire - 2 * sqrt_term_1 + 2 * D4_1 * acos_term_1) / pi / 2
  Ci_1 <- Ri_1 * pi * 2
  
  cat("\nFirst calculation:\n")
  cat("  Vi (C5):", Vi_1, "mm\n")
  cat("  D4 (previous Ri):", D4_1, "mm\n")
  cat("  Ri calculated:", round(Ri_1, 5), "mm\n")
  cat("  Ci calculated:", round(Ci_1, 3), "mm\n")
  cat("  Expected Ci:", expected_Ci[1], "mm\n")
  cat("  Difference:", round(Ci_1 - expected_Ci[1], 6), "mm\n")
  
  # Second calculation (i=2): D4 = Ri_1, C5 = Vi[2]/1000
  Vi_2 <- Vi_micrometers[2] / 1000
  D4_2 <- Ri_1  # Previous Ri = Ri_1 from first calculation
  
  sqrt_term_2 <- sqrt((Dini + D4_2 - Vi_2)^2 - D4_2^2)
  acos_term_2 <- acos(D4_2 / (Dini + D4_2 - Vi_2))
  Ri_2 <- (Cwire - 2 * sqrt_term_2 + 2 * D4_2 * acos_term_2) / pi / 2
  Ci_2 <- Ri_2 * pi * 2
  
  cat("\nSecond calculation:\n")
  cat("  Vi (C5):", Vi_2, "mm\n")
  cat("  D4 (previous Ri):", round(D4_2, 5), "mm\n")
  cat("  Ri calculated:", round(Ri_2, 5), "mm\n")
  cat("  Ci calculated:", round(Ci_2, 3), "mm\n")
  cat("  Expected Ci:", expected_Ci[2], "mm\n")
  cat("  Difference:", round(Ci_2 - expected_Ci[2], 6), "mm\n")
  
  return(result)
}

# Run the corrected test
cat("Running corrected test with your data...\n")
corrected_result <- run_your_data_corrected()
```
