---
title: "dendroanalyst"
format: pdf
editor: visual
---

Getting weird pink text along right pane of plots need to identify:

help("\_{must remake R-ex/\*.R}\_")

# Code to run at launch

```{r}
#install.packages("dendRoAnalyst")
library(dplyr)
library(readr)
library(viridis)
library(tidyr)
library(lubridate)
library(here)
library(zoo)
library(tidyverse)
library(dendRoAnalyst)
custom_colors <- c("2014" = "magenta","2015" = "maroon","2016" = "red", "2017" = "orange", "2018" = "yellow", "2019" = "green", "2020" = "lightblue", "2021" = "blue", "2022" = "purple", "2023" = "pink","2024"="darkgreen")

HD_colors <- c("2018" = "lightblue3", "2019" = "lightblue2", "2020" = "lightblue1", "2021" = "black", "2022" = "pink", "2023" = "pink2","2024"="pink3")

#keep getting errors with logical and character vectors mixing with remarks= so let's patch the code locally with getAnywhere(daily.data)
daily.data.patched<-function (df, TreeNum) 
{
    if (TreeNum > ncol(df) - 1) {
        stop("The tree number is not valid for provided dataset.")
    }
    TIME <- DATE <- dm <- time <- Max <- Min <- Time_max <- Time_min <- NULL
    tn <- TreeNum + 1
    if (!inherits(df[[1]], "Date") && !inherits(df[[1]], "POSIXct")) {
        df[[1]] <- ymd_hms(df[[1]])
    }
    df <- tibble(df) %>% select(c(1, all_of(tn)))
    df <- df %>% rename(TIME = 1, dm = 2) %>% mutate(DATE = as.Date(TIME), 
        time = format(TIME, "%H:%M:%S"))
    result <- df %>% group_by(DATE) %>% summarize(Min = min(dm, 
        na.rm = T), Time_min = time[which.min(dm)], Max = max(dm, 
        na.rm = T), Time_max = time[which.max(dm)], mean = mean(dm, 
        na.rm = T), median = median(dm, na.rm = T), amplitude = Max - 
        Min, Remarks = ifelse(isTRUE(Time_max > Time_min), "*", ""))
    return(result)
}
```

# Function Information

Summaries in the right column in bold indicate extra relevant functions. Example code for all these are in separate code chunks in their corresponding headings which you can see in the outline to the right.

+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [clim.twd](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/clim.twd.html)                                                                                                                                                                                                                                                                                             | **Calculating relative growth change during no-rain periods. Requires dendrometer (e.g., nepa17) and rainfall (e.g., rain17) data.**       |
|                                                                                                                                                                                                                                                                                                                                                                              |                                                                                                                                            |
| Description from ?clim.twd:                                                                                                                                                                                                                                                                                                                                                  | E.g.,\                                                                                                                                     |
|                                                                                                                                                                                                                                                                                                                                                                              | relative_dry_growth\<-clim.twd(df=gf_nepa17, Clim=ktm_rain17, dailyValue='max', showPlot=TRUE) \[1\]\                                      |
| This function calculates the number and the location of climatically adverse periods within a climate time series. The user can define a duration and threshold of these conditions. The function also provides the relative radial/circumferencial change during each adverse period for the original or normalized data. See Raffelsbauer et al., (2019) for more details. | Output: "The adverse climatic periods are listed below:" start_doy, end_doy, Id,                                                           |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [daily.data](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/daily.data.html)                                                                                                                                                                                                                                                                                         | Calculation of **daily statistics for dendrometer data**                                                                                   |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dendro.resample](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dendro.resample.html)                                                                                                                                                                                                                                                                               | Resampling temporal resolution of dendrometer and climate data                                                                             |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dendro.truncate](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dendro.truncate.html)                                                                                                                                                                                                                                                                               | Truncation of the dendrometer data                                                                                                         |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.detrend](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.detrend.html)                                                                                                                                                                                                                                                                                         | Detrend the dendrometer data (**should we?**)                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.fit.gompertz](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.fit.gompertz.html)                                                                                                                                                                                                                                                                               | Fitting gompertz function on annual dendrometer data                                                                                       |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [dm.na.interpolation](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/dm.na.interpolation.html)                                                                                                                                                                                                                                                                       | Detection and **interpolation of missing values** in dendrometer data.                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [gf_nepa17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/gf_nepa17.html)                                                                                                                                                                                                                                                                                           | Dendrometer data of Kathmandu for 2017 with gap filled                                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [i.jump.locator](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/i.jump.locator.html)                                                                                                                                                                                                                                                                                 | Removing artefacts due to manual adjustments of dendrometers interactively                                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [jump.locator](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/jump.locator.html)                                                                                                                                                                                                                                                                                     | **Removing artefacts due to manual adjustments of dendrometers automatically for more than one dendrometers**                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [ktm_rain17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/ktm_rain17.html)                                                                                                                                                                                                                                                                                         | Daily rainfall data of Kathmandu for 2017.                                                                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [mean_detrended.dm](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/mean_detrended.dm.html)                                                                                                                                                                                                                                                                           | Calculate mean of detrended dendrometer data.                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [mov.cor.dm](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/mov.cor.dm.html)                                                                                                                                                                                                                                                                                         | **Calculating running correlation between dendrometer data and daily climate.**                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [nepa](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa.html)                                                                                                                                                                                                                                                                                                     | Dendrometer data from Kathmandu                                                                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [nepa17](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa17.html)                                                                                                                                                                                                                                                                                                 | Dendrometer data of Kathmandu for 2017                                                                                                     |
|                                                                                                                                                                                                                                                                                                                                                                              |                                                                                                                                            |
|                                                                                                                                                                                                                                                                                                                                                                              | from the paper: "The data was derived using an automatic logging band dendrometer (DRL26, EMS Brno) recording hourly circumference change" |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [pa2](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/nepa2.html)                                                                                                                                                                                                                                                                                                     | Dendrometer data from Kathmandu version 2                                                                                                  |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [network.interpolation](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/network.interpolation.html)                                                                                                                                                                                                                                                                   | Interpolation of NA values using the dendrometer network                                                                                   |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [phase.sc](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/phase.sc.html)                                                                                                                                                                                                                                                                                             | **Application of the stem-cycle approach to calculate different phases, their duration and to plot them.**                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [phase.zg](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/phase.zg.html)                                                                                                                                                                                                                                                                                             | **Application of the zero-growth approach to calculate different phases, their duration and to plot them.**                                |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_mov.cor](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_mov.cor.html)                                                                                                                                                                                                                                                                                     | **Plotting moving correlation with climate.**                                                                                              |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_SC_output](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_SC_output.html)                                                                                                                                                                                                                                                                                 | **Plotting output of stem cycle approach (i.e., output of phase.sc function).**                                                            |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [plot_ZG_output](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/plot_ZG_output.html)                                                                                                                                                                                                                                                                                 | **Plotting output of zero-growth approach (i.e., output of phase.zg function).**                                                           |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [read.dendrometer](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/read.dendrometer.html)                                                                                                                                                                                                                                                                             | **Reading dendrometer data and automatically converts datetime format to what is needed.**                                                 |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+
| [twd.maxima](http://127.0.0.1:42961/help/library/dendRoAnalyst/html/twd.maxima.html)                                                                                                                                                                                                                                                                                         | **Locating the maxima of TWD periods**                                                                                                     |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------+

-   The ‘phase.sc’ and ‘phase.zg’ only calculate the phases and phase statistics.

-   There are two separate functions added to plot the output of ‘phase.sc’ and ‘phase.zg’.

-   A new function, ‘dm.detrend’ has been added to detrend the dendrometer data using ‘gam’, ‘first difference’ and ‘gompertz’ function.

-   A new function ‘dm.fit.gompertz’ is introduced to fit Gompertz function. It also provide option to calculate the growth rate applying first derivative of the Gompertz function.

-   The function ‘spline.interpolation’ is now ‘dm.na.interpolation’, with two methods: ‘spline’ and ‘seasonal’ interpolation. The seasonal interpolation used ARIMA model to detect the daily pattern for three weeks (the week of gap occurrence + one week earlier and later) and interpolates the missing values.

-   A new function ‘read.dendrometer’ is included to read the dendrometer data from TXT, CSV and EXCEL files.

-   A new function ‘mean.detrended.dm’ calculates the averaged detrended dendrometer data with possibility getting first order autocorrelation free averaged series.

-   A new function ‘mov.cor.dm’ is added for calculating climate-growth relationships.

# Step 0: Dendro data: loading in and prepping

Example data for DendRoAnalyst has hourly dendrometer data and daily rainfall data, so we will be condensing our data down to match.

```{r}
#raw dendrometer data
load(here(file="DendroAnalyst/dscmet_422_all.RData")) #calling saved R object from ecomatik.rmd
#checking bad flags - no bad ones found so this code won't be rerun 
#unique_422_strings <- unique(
#  unlist(
#    sapply(
#      dscmet_422_all, function(x) x[is.character(x)])))

#getting hourly averages - starting with adding an hour column
dscmet_422_all$hour <- format(
  dscmet_422_all$Date, "%Y-%m-%d %H:00:00")
dscmet_422_all$Date<-as.POSIXct(dscmet_422_all$hour,format="%Y-%m-%d %H")
#getting daily averages - starting with adding a day column
dscmet_422_all$day <- format(
  dscmet_422_all$Date, "%Y-%m-%d")
dscmet_422_all$day<-as.POSIXct(dscmet_422_all$day,format="%Y-%m-%d")

#309 - OG DF - need to check if these values are accurate
d309_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg309 = mean(DEND309_AVG_4200_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg309))
colnames(d309_hourly)[colnames(d309_hourly) == 'hour'] <- 'Date'

#daily values 
d309_daily <- dscmet_422_all |> 
  group_by(day) |> 
  summarise(dailygrowth = mean(DEND309_AVG_4200_0_01,na.rm=TRUE))
colnames(d309_daily)[colnames(d309_daily) == 'day'] <- 'Date'

#310 - OG DF
d310_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg310 = mean(DEND310_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg310))  

#311 - OG DF
d311_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg311 = mean(DEND311_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg311)) 

#651 - OG WRC
d651_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg651 = mean(DEND651_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg651)) 


#654 - Secondary DF - filtering to start after -8mil weirdness
d654_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg654 = mean(DEND654_AVG_150_0_01,na.rm=TRUE))|>
  filter(hour >= as.POSIXct("2018-11-05 15:00:00"))|>
  filter(!is.nan(hourly_avg654)) 

#655 - Secondary DF - filtering to start after -8mil weirdness
d655_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg655 = mean(DEND655_AVG_150_0_01,na.rm=TRUE))|>
  filter(hour >= as.POSIXct("2018-11-05 15:00:00"))|>
  filter(!is.nan(hourly_avg655))

#656 - Secondary DF - filtering to start after -8mil weirdness
d656_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg656 = mean(DEND656_AVG_150_0_01,na.rm=TRUE))|>
  filter(hour >= as.POSIXct("2018-11-05 15:00:00"))|>
  filter(!is.nan(hourly_avg656)) 

#860 - Secondary DF - note only ~15% of original canopy remains alive
#also filtering to start after -8mil weirdness
d860_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg860 = mean(DEND860_AVG_150_0_01,na.rm=TRUE))|>
  filter(hour >= as.POSIXct("2019-05-28 09:00:00"))|>
  filter(!is.nan(hourly_avg860)) 

#646 OG WH
d646_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg646 = mean(DEND646_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg646)) 

#648 OG WH
d648_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg648 = mean(DEND648_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg648)) 

#652 OG WH
d652_hourly <- dscmet_422_all |> 
  group_by(hour) |> 
  summarise(hourly_avg652 = mean(DEND652_AVG_150_0_01,na.rm=TRUE))|>
  filter(!is.nan(hourly_avg652)) 
```

# Step 0: Rainfall data (from climate comparisons updated.rmd)

```{r}
load(file=here("P230.RData"))

#checking for bad flags - results in "values_to_remove_P230rain" so we won't rerun
#P230_rain<-P230_df|>select(c(Date,PRECIP_TOT_100_0_01,Flag_PRECIP_TOT_100_0_01))
#unique_rain_strings <- unique(
#  unlist(
#    sapply(
#      P230_rain, function(x) x[is.character(x)])))
values_to_remove_P230rain <- c("Q", "QQ")

# Filter data frame to swap back values with NA
filtered_P230_rain <- P230_df[!rowSums(sapply(P230_df, function(x) x %in% values_to_remove_P230rain | is.na(x))) > 0, ]
save(filtered_P230_rain,file = "filtered_P230_rain.RData")

#daily rain totals
filtered_P230_rain$day <- format(
  filtered_P230_rain$Date, "%Y-%m-%d")
filtered_P230_rain_daily <- filtered_P230_rain |> 
  group_by(day) |> 
  summarise(value = sum(PRECIP_TOT_100_0_01))
filtered_P230_rain_daily$IDs<-"precip_sum"
colnames(filtered_P230_rain_daily)[colnames(filtered_P230_rain_daily) == 'day'] <- 'Date'

#hourly totals 
filtered_P230_rain$hour <- format(
  filtered_P230_rain$Date, "%Y-%m-%d %H:00:00")
filtered_P230_rain_hourly <- filtered_P230_rain |> 
  group_by(hour) |> 
  summarise(precip_sum = sum(PRECIP_TOT_100_0_01,na.rm = TRUE))
filtered_P230_rain_hourly$time<-as.POSIXct(filtered_P230_rain_hourly$hour,format="%Y-%m-%d %H:00:00")
filtered_P230_rain_hourly<-filtered_P230_rain_hourly|>select(c(time,precip_sum))
filtered_P230_rain_hourly$IDs<-"precip_sum"
```

# Step 1: NA Interpolation

Skipped for now because the paper says that due to increasing uncertainty, this step is not recommended for gaps larger than 24 hours

## Interpolation option 1: dm.na.interpolation

```         
dm.na.interpolation(df, resolution, fill = FALSE, method = "spline") 
```

### Explanation of options

+--------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`         | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and following columns with dendrometer data for the same temporal resolution and time period. |
+--------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `resolution` | integer, indicating the resolution of dendrometer data in **minutes**.                                                                                                                 |
+--------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `fill`       | logical, if `TRUE` it fills the `NA` values using spline interpolation. Default is `FALSE`.                                                                                            |
+--------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `method`     | string, 'spline' for the spline interpolation or 'seasonal' for the seasonal interpolation                                                                                             |
+--------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Locating maxima TWD Periods with twd.maxima

`{r} # library(dendRoAnalyst) #> data(gf_nepa17) #> df1=gf_nepa17[2500:3500,] #> twd_max<-twd.maxima(df=df1, TreeNum=2, showPlot=FALSE) #> head(twd_max,10) #             start.time            end.time                Time         TWD #20  2017-04-15 11:00:00 2017-04-16 04:00:00 2017-04-15 22:00:00 0.028245289 #43  2017-04-16 08:00:00 2017-04-18 06:00:00 2017-04-16 21:00:00 0.070221431 #    duration_from_start twd.number #20                  540          1 #43                  660          2}`

## Interpolation option 2: using dendrometer network via network.interpolation

```         
network.interpolation(df, referenceDF, niMethod) 
```

### Explanation of options

+---------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`          | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and dendrometer data in the second column and onward. The data gaps must be filled with `NA` using the gap.interpolation function. |
+---------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `referenceDF` | dataframe with other dendrometers to be used as reference for the interpolation. The more dendrometers are included, the more robust will be the interpolation.                                                             |
+---------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `niMethod`    | string, either *'linear'* or *'proportional'* for interpolation method.                                                                                                                                                     |
+---------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

`{r} #Example from vignette #df1<-gf_nepa17 #> # Creating an artificial reference dataset. #> df2<-cbind(gf_nepa17,gf_nepa17[,2:3],gf_nepa17[,2:3]) #> # Creating gaps in dataset by replacing some of the reading with NA in dataset. #> df1[40:50,3]<-NA #> # Using proportional interpolation method. #> df1_NI<-network.interpolation(df=df1, referenceDF=df2, niMethod='proportional') #Warning in norm.inter(t, adj.alpha) : #  extreme order statistics used as endpoints #[1] "Interpolation in T3" #         Time Interpolated_value    LCI95    UCI95 #1  1483369200           48.31525 48.30600 48.32141 #2  1483372800           48.31522 48.31331 48.31585 #> head(df1_NI,10) #                  Time       T2       T3 #1  2017-01-01 00:00:00 62.26459 48.16801 #...10 2017-01-01 09:00:00 62.46552 48.21498}`

# Step 2: Remove artefacts from manual adjustments with jump.locator

```         
i.jump.locator(df, TreeNum, v) 
```

### Explanation of options for similar function i.jump.locator (only difference is that it adds TreeNum whereas jump.locator doesn't)

output is a new df that locates and adjusts the artefacts

+-----------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`      | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS` and the dendrometer data in following columns.                                                        |
+-----------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum` | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                                              |
+-----------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `v`       | numerical value which is considered as artefact. E.g. `v`=1 implies that if the difference to the consecutive data point is more than 1 or less than -1, it will be considered as an artefact. |
+-----------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

First look at the raw data outputs and see how the jumps are to help set the "v=" value. When you run the function it makes a table with the number of rows equal to the number of removed values and the columns equal to observation number, timestamp of the jump, and the value of the jump. If you name the jump locator object, that will add an object to your environment of the adjusted df with the jump values removed.

```{r}
jump_free_d309<-jump.locator(df=d309_hourly,v=200) 
#this takes out 41 values and makes a table with the jump value & time and makes a df with the removed values 
jump_free_d309$hourly_avg309_raw<-d309_hourly$hourly_avg309
jump_free_d309$hourly_avg309_0jump<-jump_free_d309$hourly_avg309
jump_free_d309$ts <- jump_free_d309$TIME
jump_free_d309$yr <- format(jump_free_d309$ts, "%Y")
jump_free_d309$yr <- as.factor(jump_free_d309$yr)
jump_free_d309$DOY <- yday(jump_free_d309$ts)

jump_free_d310<-jump.locator(df=d310_hourly,v=200)
jump_free_d310$hourly_avg310_raw <- d310_hourly$hourly_avg310
jump_free_d310$hourly_avg310_0jump <- jump_free_d310$hourly_avg310
jump_free_d310$ts <- jump_free_d310$TIME
jump_free_d310$yr <- format(jump_free_d310$ts, "%Y")
jump_free_d310$yr <- as.factor(jump_free_d310$yr)
jump_free_d310$DOY <- yday(jump_free_d310$ts)

jump_free_d311 <- jump.locator(df = d311_hourly, v = 200) 
jump_free_d311$hourly_avg311_raw <- d311_hourly$hourly_avg311
jump_free_d311$hourly_avg311_0jump <- jump_free_d311$hourly_avg311
jump_free_d311$ts <- jump_free_d311$TIME
jump_free_d311$yr <- format(jump_free_d311$ts, "%Y")
jump_free_d311$yr <- as.factor(jump_free_d311$yr)
jump_free_d311$DOY <- yday(jump_free_d311$ts)


jump_free_d646 <- jump.locator(df = d646_hourly, v = 200) 
jump_free_d646$hourly_avg646_raw <- d646_hourly$hourly_avg646
jump_free_d646$hourly_avg646_0jump <- jump_free_d646$hourly_avg646
jump_free_d646$ts <- jump_free_d646$TIME
jump_free_d646$yr <- format(jump_free_d646$ts, "%Y")
jump_free_d646$yr <- as.factor(jump_free_d646$yr)
jump_free_d646$DOY <- yday(jump_free_d646$ts)


jump_free_d648 <- jump.locator(df = d648_hourly, v = 200) 
jump_free_d648$hourly_avg648_raw <- d648_hourly$hourly_avg648
jump_free_d648$hourly_avg648_0jump <- jump_free_d648$hourly_avg648
jump_free_d648$ts <- jump_free_d648$TIME
jump_free_d648$yr <- format(jump_free_d648$ts, "%Y")
jump_free_d648$yr <- as.factor(jump_free_d648$yr)
jump_free_d648$DOY <- yday(jump_free_d648$ts)


jump_free_d652 <- jump.locator(df = d652_hourly, v = 200) 
jump_free_d652$hourly_avg652_raw <- d652_hourly$hourly_avg652
jump_free_d652$hourly_avg652_0jump <- jump_free_d652$hourly_avg652
jump_free_d652$ts <- jump_free_d652$TIME
jump_free_d652$yr <- format(jump_free_d652$ts, "%Y")
jump_free_d652$yr <- as.factor(jump_free_d652$yr)
jump_free_d652$DOY <- yday(jump_free_d652$ts)


jump_free_d654 <- jump.locator(df = d654_hourly, v = 200) 
jump_free_d654$hourly_avg654_raw <- d654_hourly$hourly_avg654
jump_free_d654$hourly_avg654_0jump <- jump_free_d654$hourly_avg654
jump_free_d654$ts <- jump_free_d654$TIME
jump_free_d654$yr <- format(jump_free_d654$ts, "%Y")
jump_free_d654$yr <- as.factor(jump_free_d654$yr)
jump_free_d654$DOY <- yday(jump_free_d654$ts)


jump_free_d655 <- jump.locator(df = d655_hourly, v = 200) 
jump_free_d655$hourly_avg655_raw <- d655_hourly$hourly_avg655
jump_free_d655$hourly_avg655_0jump <- jump_free_d655$hourly_avg655
jump_free_d655$ts <- jump_free_d655$TIME
jump_free_d655$yr <- format(jump_free_d655$ts, "%Y")
jump_free_d655$yr <- as.factor(jump_free_d655$yr)
jump_free_d655$DOY <- yday(jump_free_d655$ts)


jump_free_d656 <- jump.locator(df = d656_hourly, v = 200) 
jump_free_d656$hourly_avg656_raw <- d656_hourly$hourly_avg656
jump_free_d656$hourly_avg656_0jump <- jump_free_d656$hourly_avg656
jump_free_d656$ts <- jump_free_d656$TIME
jump_free_d656$yr <- format(jump_free_d656$ts, "%Y")
jump_free_d656$yr <- as.factor(jump_free_d656$yr)
jump_free_d656$DOY <- yday(jump_free_d656$ts)


jump_free_d860 <- jump.locator(df = d860_hourly, v = 200) 
jump_free_d860$hourly_avg860_raw <- d860_hourly$hourly_avg860
jump_free_d860$hourly_avg860_0jump <- jump_free_d860$hourly_avg860
jump_free_d860$ts <- jump_free_d860$TIME
jump_free_d860$yr <- format(jump_free_d860$ts, "%Y")
jump_free_d860$yr <- as.factor(jump_free_d860$yr)
jump_free_d860$DOY <- yday(jump_free_d860$ts)


jump_free_d651 <- jump.locator(df = d651_hourly, v = 200) 
jump_free_d651$hourly_avg651_raw <- d651_hourly$hourly_avg651
jump_free_d651$hourly_avg651_0jump <- jump_free_d651$hourly_avg651
jump_free_d651$ts <- jump_free_d651$TIME
jump_free_d651$yr <- format(jump_free_d651$ts, "%Y")
jump_free_d651$yr <- as.factor(jump_free_d651$yr)
jump_free_d651$DOY <- yday(jump_free_d651$ts)
```

### Plots to see the difference

#### Plot type 1: 1 curve per dendrometer

First define growing season for each year and Heat Dome line\
Second define raw and jump free ggplots\
Third make a figure with raw and jump free plots next to each other

```{r}
#Defining ranges for plot shading

#maybe we come back to adjust this after perhaps dentifying start of actual growing szn
growing_seasons_default <- data.frame(
  xmin = as.POSIXct(c("2018-04-01 00:00","2019-04-01 00:00","2020-04-01 00:00","2021-04-01 00:00","2022-04-01 00:00","2023-04-01 00:00","2024-04-01 00:00")),
  xmax = as.POSIXct(c("2018-10-01 00:00","2019-10-01 00:00","2020-10-01 00:00","2021-10-01 00:00","2022-10-01 00:00","2023-10-01 00:00","2024-10-01 00:00")),
  fill = "April1-Oct.1")
vline_HD<-as.POSIXct("2021-06-25 00:00")

#"\u25B2" = triangle

#309
raw309_1curve <- ggplot(
  jump_free_d309,aes(x=TIME,y=hourly_avg309_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG DF (ID 309) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free309_1curve <- ggplot(
  jump_free_d309,aes(x=TIME,y=hourly_avg309_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw309_1curve,jump_free309_1curve)
two_1curve_plots309 <- 
  gridExtra::grid.arrange(raw309_1curve,jump_free309_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots309.png"), 
       plot = two_1curve_plots309)


#310
raw310_1curve <- ggplot(
  jump_free_d310,aes(x=TIME,y=hourly_avg310_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG DF (ID 310) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free310_1curve <- ggplot(
  jump_free_d310,aes(x=TIME,y=hourly_avg310_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw310_1curve,jump_free310_1curve)
two_1curve_plots310 <- 
  gridExtra::grid.arrange(raw310_1curve, jump_free310_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots310.png"), 
       plot = two_1curve_plots310)

#311
raw311_1curve <- ggplot(
  jump_free_d311,aes(x=TIME,y=hourly_avg311_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG DF (ID 311) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free311_1curve <- ggplot(
  jump_free_d311,aes(x=TIME,y=hourly_avg311_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw311_1curve,jump_free311_1curve)
two_1curve_plots311 <- 
  gridExtra::grid.arrange(raw311_1curve, jump_free311_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots311.png"), 
       plot = two_1curve_plots311)

#646
raw646_1curve <- ggplot(
  jump_free_d646,aes(x=TIME,y=hourly_avg646_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG WH (ID 646) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free646_1curve <- ggplot(
  jump_free_d646,aes(x=TIME,y=hourly_avg646_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw646_1curve,jump_free646_1curve)
two_1curve_plots646 <- 
  gridExtra::grid.arrange(raw646_1curve, jump_free646_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots646.png"), 
       plot = two_1curve_plots646)

#648
raw648_1curve <- ggplot(
  jump_free_d648,aes(x=TIME,y=hourly_avg648_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG WH (ID 648) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(
    limits = as.POSIXct(c("2023-01-01 00:00:00", NA)),
    date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free648_1curve <- ggplot(
  jump_free_d648,aes(x=TIME,y=hourly_avg648_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(
    limits = as.POSIXct(c("2023-01-01 00:00:00", NA)),
    date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw648_1curve,jump_free648_1curve)
two_1curve_plots648 <- 
  gridExtra::grid.arrange(raw648_1curve, jump_free648_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots648.png"), 
       plot = two_1curve_plots648)

#651
raw651_1curve <- ggplot(
  jump_free_d651,aes(x=TIME,y=hourly_avg651_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG WRC (ID 651) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(
    limits = as.POSIXct(c("2023-01-01 00:00:00", NA)),
    date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free651_1curve <- ggplot(
  jump_free_d651,aes(x=TIME,y=hourly_avg651_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(
    limits = as.POSIXct(c("2023-01-01 00:00:00", NA)),
    date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw651_1curve,jump_free651_1curve)
two_1curve_plots651 <- 
  gridExtra::grid.arrange(raw651_1curve, jump_free651_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots651.png"), 
       plot = two_1curve_plots651)

#652
raw652_1curve <- ggplot(
  jump_free_d652,aes(x=TIME,y=hourly_avg652_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="OG WH (ID 652) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(
    limits = as.POSIXct(c("2023-01-01 00:00:00", NA)),
    date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free652_1curve <- ggplot(
  jump_free_d652,aes(x=TIME,y=hourly_avg652_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(
    limits = as.POSIXct(c("2023-01-01 00:00:00", NA)),
    date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw652_1curve,jump_free652_1curve)
two_1curve_plots652 <-
  gridExtra::grid.arrange(raw652_1curve, jump_free652_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots652.png"), 
       plot = two_1curve_plots652)

#654 
raw654_1curve <- ggplot(
  jump_free_d654,aes(x=TIME,y=hourly_avg654_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="2° DF (ID 654) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
jump_free654_1curve <- ggplot(
  jump_free_d654,aes(x=TIME,y=hourly_avg654_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD), 
               linetype = "solid", color = "red")+ 
    geom_rect(data = growing_seasons_default, 
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax, 
                                      ymin = -Inf, ymax = Inf, 
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw654_1curve,jump_free654_1curve)
two_1curve_plots654 <- 
  gridExtra::grid.arrange(raw654_1curve, jump_free654_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots654.png"), 
       plot = two_1curve_plots654)

#655
raw655_1curve <- ggplot(
  jump_free_d655,aes(x=TIME,y=hourly_avg655_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="2° DF (ID 655) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD),
               linetype = "solid", color = "red")+
    geom_rect(data = growing_seasons_default,
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax,
                                      ymin = -Inf, ymax = Inf,
                                      fill = fill),alpha = 0.1)
jump_free655_1curve <- ggplot(
  jump_free_d655,aes(x=TIME,y=hourly_avg655_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD),
               linetype = "solid", color = "red")+
    geom_rect(data = growing_seasons_default,
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax,
                                      ymin = -Inf, ymax = Inf,
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw655_1curve,jump_free655_1curve)
two_1curve_plots655 <- 
  gridExtra::grid.arrange(raw655_1curve, jump_free655_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots655.png"), 
       plot = two_1curve_plots655)

#656
raw656_1curve <- ggplot(
  jump_free_d656,aes(x=TIME,y=hourly_avg656_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="2° DF (ID 656) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD),
               linetype = "solid", color = "red")+
    geom_rect(data = growing_seasons_default,
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax,
                                      ymin = -Inf, ymax = Inf,
                                      fill = fill),alpha = 0.1)
jump_free656_1curve <- ggplot(
  jump_free_d656,aes(x=TIME,y=hourly_avg656_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD),
               linetype = "solid", color = "red")+
    geom_rect(data = growing_seasons_default,
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax,
                                      ymin = -Inf, ymax = Inf,
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw656_1curve,jump_free656_1curve)
two_1curve_plots656 <- 
  gridExtra::grid.arrange(raw656_1curve, jump_free656_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots656.png"), 
       plot = two_1curve_plots656)

#860
raw860_1curve <- ggplot(
  jump_free_d860,aes(x=TIME,y=hourly_avg860_raw))+
    geom_point(size=1.25,alpha=0.5)+
    labs(title="2° DF (ID 860) Stem Circumference Change (µm)",
         x="Date",y="Unadjusted\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD),
               linetype = "solid", color = "red")+
    geom_rect(data = growing_seasons_default,
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax,
                                      ymin = -Inf, ymax = Inf,
                                      fill = fill),alpha = 0.1)
jump_free860_1curve <- ggplot(
  jump_free_d860,aes(x=TIME,y=hourly_avg860_0jump))+
    geom_point(size=1.25,alpha=0.5)+
    labs(x="Date",y="Jump-Free\u25B2")+
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")+
    theme_minimal()+
    geom_vline(xintercept = as.numeric(vline_HD),
               linetype = "solid", color = "red")+
    geom_rect(data = growing_seasons_default,
              inherit.aes = FALSE,aes(xmin = xmin, xmax = xmax,
                                      ymin = -Inf, ymax = Inf,
                                      fill = fill),alpha = 0.1)
#combining plots and saving 
gridExtra::grid.arrange(raw860_1curve,jump_free860_1curve)
two_1curve_plots860 <- 
  gridExtra::grid.arrange(raw860_1curve, jump_free860_1curve)
ggsave(here("Plots/1-Curve Per Dendro/two_1curve_plots860.png"), 
       plot = two_1curve_plots860)

```

#### Plot type 2: each year starting at 0

##### 309

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d309$raw_309 <- jump_free_d309$hourly_avg309_raw 
jump_free_d309$jumpfree_309 <- jump_free_d309$hourly_avg309_0jump

wide309gro<-jump_free_d309|>
  select(c(raw_309,jumpfree_309,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_309,jumpfree_309),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide309gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide309gro$unadjusted_2018<-wide309gro$raw_309_2018 
wide309gro$nojump_2018<-wide309gro$jumpfree_309_2018 
#no adjustment for 2018 since first readings are in July

wide309gro$unadjusted_2019<-wide309gro$raw_309_2019-2016.833
wide309gro$nojump_2019<-wide309gro$jumpfree_309_2019-2016.833

wide309gro$unadjusted_2020 <- wide309gro$raw_309_2020-(-658.5833)
wide309gro$nojump_2020     <- wide309gro$jumpfree_309_2020-676.0000

wide309gro$unadjusted_2021 <- wide309gro$raw_309_2021-10387.00
wide309gro$nojump_2021      <- wide309gro$jumpfree_309_2021-(-489.5833)

wide309gro$unadjusted_2022 <- wide309gro$raw_309_2022-12954.83
wide309gro$nojump_2022  <- wide309gro$jumpfree_309_2022-1562.667

wide309gro$unadjusted_2023 <- wide309gro$raw_309_2023-16044.00
wide309gro$nojump_2023 <- wide309gro$jumpfree_309_2023-4843.771

wide309gro$unadjusted_2024 <- wide309gro$raw_309_2024-11627.33
wide309gro$nojump_2024   <- wide309gro$jumpfree_309_2024-7569.021


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long309gro<- wide309gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long309gro_unadjusted<-long309gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long309gro_nojump<-long309gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))

#ggplot where each year is a panel in one figure (not using)
#ggplot(long309gro, aes(x = DOY, y = growth, color = adjustment)) +
#  geom_line(alpha = 0.8) +
#  facet_wrap(~ yr) +
#  theme_minimal() +
#  labs(y = "Stem Circumference (µm)", color = "Data Type")

#ggplot with each year a separate line: unadjusted
unadjusted309_factor_yr <- 
  ggplot(long309gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG DF (ID 309) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted309_factor_yr.png"), 
       plot = unadjusted309_factor_yr)


#ggplot with each year a separate line: jump-free
nojump309_factor_yr<-
  ggplot(long309gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG DF (ID 309) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump309_factor_yr.png"), 
       plot = nojump309_factor_yr)

unadjusted309_factor_yr
nojump309_factor_yr
```

##### 310

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d310$raw_310 <- jump_free_d310$hourly_avg310_raw 
jump_free_d310$jumpfree_310 <- jump_free_d310$hourly_avg310_0jump

wide310gro<-jump_free_d310|>
  select(c(raw_310,jumpfree_310,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_310,jumpfree_310),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide310gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide310gro$unadjusted_2018<-wide310gro$raw_310_2018 
wide310gro$nojump_2018<-wide310gro$jumpfree_310_2018 

wide310gro$unadjusted_2019<-wide310gro$raw_310_2019-3965.000
wide310gro$nojump_2019<-wide310gro$jumpfree_310_2019-3118.333

wide310gro$unadjusted_2020 <- wide310gro$raw_310_2020-11829.92
wide310gro$nojump_2020     <- wide310gro$jumpfree_310_2020-10651.00

wide310gro$unadjusted_2021 <- wide310gro$raw_310_2021-4932.750
wide310gro$nojump_2021      <- wide310gro$jumpfree_310_2021-18123.08

wide310gro$unadjusted_2022 <- wide310gro$raw_310_2022-5927.250
wide310gro$nojump_2022  <- wide310gro$jumpfree_310_2022-24102.08

wide310gro$unadjusted_2023 <- wide310gro$raw_310_2023-6099.917
wide310gro$nojump_2023 <- wide310gro$jumpfree_310_2023-29155.43

wide310gro$unadjusted_2024 <- wide310gro$raw_310_2024-2261.083
wide310gro$nojump_2024   <- wide310gro$jumpfree_310_2024-34458.52


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long310gro<- wide310gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long310gro_unadjusted<-long310gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long310gro_nojump<-long310gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))

#ggplot with each year a separate line: unadjusted
unadjusted310_factor_yr <- 
  ggplot(long310gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG DF (ID 310) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted310_factor_yr.png"), 
       plot = unadjusted310_factor_yr)


#ggplot with each year a separate line: jump-free
nojump310_factor_yr<-
  ggplot(long310gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG DF (ID 310) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump310_factor_yr.png"), 
       plot = nojump310_factor_yr)

unadjusted310_factor_yr
nojump310_factor_yr
```

##### 311

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d311$raw_311 <- jump_free_d311$hourly_avg311_raw 
jump_free_d311$jumpfree_311 <- jump_free_d311$hourly_avg311_0jump

wide311gro<-jump_free_d311|>
  select(c(raw_311,jumpfree_311,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_311,jumpfree_311),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide311gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide311gro$unadjusted_2018<-wide311gro$raw_311_2018 
wide311gro$nojump_2018<-wide311gro$jumpfree_311_2018 

wide311gro$unadjusted_2019<-wide311gro$raw_311_2019-6911.750
wide311gro$nojump_2019<-wide311gro$jumpfree_311_2019-6378.500

wide311gro$unadjusted_2020 <- wide311gro$raw_311_2020-22133.25
wide311gro$nojump_2020     <- wide311gro$jumpfree_311_2020-18293.96

wide311gro$unadjusted_2021 <- wide311gro$raw_311_2021-5447.667
wide311gro$nojump_2021      <- wide311gro$jumpfree_311_2021-37181.96

wide311gro$unadjusted_2022 <- wide311gro$raw_311_2022-3337.000
wide311gro$nojump_2022  <- wide311gro$jumpfree_311_2022-52970.79

wide311gro$unadjusted_2023 <- wide311gro$raw_311_2023-2924.083
wide311gro$nojump_2023 <- wide311gro$jumpfree_311_2023-67335.13

wide311gro$unadjusted_2024 <- wide311gro$raw_311_2024-4723.000
wide311gro$nojump_2024   <- wide311gro$jumpfree_311_2024-79690.38


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long311gro<- wide311gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long311gro_unadjusted<-long311gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long311gro_nojump<-long311gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))


#ggplot with each year a separate line: unadjusted
unadjusted311_factor_yr <- 
  ggplot(long311gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG DF (ID 311) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted311_factor_yr.png"), 
       plot = unadjusted311_factor_yr)


#ggplot with each year a separate line: jump-free
nojump311_factor_yr<-
  ggplot(long311gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG DF (ID 311) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump311_factor_yr.png"), 
       plot = nojump311_factor_yr)

unadjusted311_factor_yr
nojump311_factor_yr
```

##### 646

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d646$raw_646 <- jump_free_d646$hourly_avg646_raw 
jump_free_d646$jumpfree_646 <- jump_free_d646$hourly_avg646_0jump

wide646gro<-jump_free_d646|>
  select(c(raw_646,jumpfree_646,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_646,jumpfree_646),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide646gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide646gro$unadjusted_2018<-wide646gro$raw_646_2018 
wide646gro$nojump_2018<-wide646gro$jumpfree_646_2018 

wide646gro$unadjusted_2019<-wide646gro$raw_646_2019-3600.333
wide646gro$nojump_2019<-wide646gro$jumpfree_646_2019-3600.333

wide646gro$unadjusted_2020 <- wide646gro$raw_646_2020-8365.083
wide646gro$nojump_2020     <- wide646gro$jumpfree_646_2020-8365.333

wide646gro$unadjusted_2021 <- wide646gro$raw_646_2021-14004.67
wide646gro$nojump_2021      <- wide646gro$jumpfree_646_2021-12486.08

wide646gro$unadjusted_2022 <- wide646gro$raw_646_2022-5758.917
wide646gro$nojump_2022  <- wide646gro$jumpfree_646_2022-18121.58

wide646gro$unadjusted_2023 <- wide646gro$raw_646_2023-7089.000
wide646gro$nojump_2023 <- wide646gro$jumpfree_646_2023-22514.58

wide646gro$unadjusted_2024 <- wide646gro$raw_646_2024-11417.58
wide646gro$nojump_2024   <- wide646gro$jumpfree_646_2024-26843.17


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long646gro<- wide646gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long646gro_unadjusted<-long646gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long646gro_nojump<-long646gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))

#ggplot with each year a separate line: unadjusted
unadjusted646_factor_yr <- 
  ggplot(long646gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG WH (ID 646) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted646_factor_yr.png"), 
       plot = unadjusted646_factor_yr)


#ggplot with each year a separate line: jump-free
nojump646_factor_yr<-
  ggplot(long646gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="OG WH (ID 646) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump646_factor_yr.png"), 
       plot = nojump646_factor_yr)

unadjusted646_factor_yr
nojump646_factor_yr
```

##### 648

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d648$raw_648 <- jump_free_d648$hourly_avg648_raw 
jump_free_d648$jumpfree_648 <- jump_free_d648$hourly_avg648_0jump

wide648gro<-jump_free_d648|>
  select(c(raw_648,jumpfree_648,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_648,jumpfree_648),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide648gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide648gro$unadjusted_2023 <- wide648gro$raw_648_2023-5097.000
wide648gro$nojump_2023 <- wide648gro$jumpfree_648_2023-5097.000
#first reading is April 3 but we normalize it anyway

wide648gro$unadjusted_2024 <- wide648gro$raw_648_2024-8609.833
wide648gro$nojump_2024   <- wide648gro$jumpfree_648_2024-8846.500


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long648gro<- wide648gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long648gro_unadjusted<-long648gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long648gro_nojump<-long648gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))

#ggplot with each year a separate line: unadjusted
unadjusted648_factor_yr <- 
  ggplot(long648gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+
  theme_minimal()+
  labs(title="OG WH (ID 648) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted648_factor_yr.png"), 
       plot = unadjusted648_factor_yr)


#ggplot with each year a separate line: jump-free
nojump648_factor_yr<-
  ggplot(long648gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  theme_minimal()+
  labs(title="OG WH (ID 648) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump648_factor_yr.png"), 
       plot = nojump648_factor_yr)

unadjusted648_factor_yr
nojump648_factor_yr
```

##### 651

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d651$raw_651 <- jump_free_d651$hourly_avg651_raw 
jump_free_d651$jumpfree_651 <- jump_free_d651$hourly_avg651_0jump

wide651gro<-jump_free_d651|>
  select(c(raw_651,jumpfree_651,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_651,jumpfree_651),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide651gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide651gro$unadjusted_2023 <- wide651gro$raw_651_2023-1680.455
wide651gro$nojump_2023 <- wide651gro$jumpfree_651_2023-1680.455
#first reading is April 3 but we normalize it anyway

wide651gro$unadjusted_2024 <- wide651gro$raw_651_2024-6491.167
wide651gro$nojump_2024   <- wide651gro$jumpfree_651_2024-6491.167


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long651gro<- wide651gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long651gro_unadjusted<-long651gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long651gro_nojump<-long651gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))


#ggplot with each year a separate line: unadjusted
unadjusted651_factor_yr <- 
  ggplot(long651gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  theme_minimal()+
  labs(title="OG WRC (ID 651) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) 
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted651_factor_yr.png"), 
       plot = unadjusted651_factor_yr)


#ggplot with each year a separate line: jump-free
nojump651_factor_yr<-
  ggplot(long651gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  theme_minimal()+
  labs(title="OG WRC (ID 651) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump651_factor_yr.png"), 
       plot = nojump651_factor_yr)

unadjusted651_factor_yr
nojump651_factor_yr
```

##### 652

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d652$raw_652 <- jump_free_d652$hourly_avg652_raw 
jump_free_d652$jumpfree_652 <- jump_free_d652$hourly_avg652_0jump

wide652gro<-jump_free_d652|>
  select(c(raw_652,jumpfree_652,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_652,jumpfree_652),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide652gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide652gro$unadjusted_2023 <- wide652gro$raw_652_2023-6909.000
wide652gro$nojump_2023 <- wide652gro$jumpfree_652_2023-6909.000
#first reading is April 3 but we normalize it anyway

wide652gro$unadjusted_2024 <- wide652gro$raw_652_2024-9722.083
wide652gro$nojump_2024   <- wide652gro$jumpfree_652_2024-9722.083


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long652gro<- wide652gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long652gro_unadjusted<-long652gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long652gro_nojump<-long652gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))


#ggplot with each year a separate line: unadjusted
unadjusted652_factor_yr <- 
  ggplot(long652gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  theme_minimal()+
  labs(title="OG WH (ID 652) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) 
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted652_factor_yr.png"), 
       plot = unadjusted652_factor_yr)


#ggplot with each year a separate line: jump-free
nojump652_factor_yr<-
  ggplot(long652gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  theme_minimal()+
  labs(title="OG WH (ID 652) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump652_factor_yr.png"), 
       plot = nojump652_factor_yr)

unadjusted652_factor_yr
nojump652_factor_yr
```

##### 654

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d654$raw_654 <- jump_free_d654$hourly_avg654_raw 
jump_free_d654$jumpfree_654 <- jump_free_d654$hourly_avg654_0jump

wide654gro<-jump_free_d654|>
  select(c(raw_654,jumpfree_654,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_654,jumpfree_654),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide654gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide654gro$unadjusted_2019<-wide654gro$raw_654_2019-1701.667
wide654gro$nojump_2019<-wide654gro$jumpfree_654_2019-1701.667

wide654gro$unadjusted_2020 <- wide654gro$raw_654_2020-8065.917
wide654gro$nojump_2020     <- wide654gro$jumpfree_654_2020-8065.917

wide654gro$unadjusted_2021 <- wide654gro$raw_654_2021-3945.000
wide654gro$nojump_2021      <- wide654gro$jumpfree_654_2021-17100.69

wide654gro$unadjusted_2022 <- wide654gro$raw_654_2022-2531.000
wide654gro$nojump_2022  <- wide654gro$jumpfree_654_2022-27357.36

wide654gro$unadjusted_2023 <- wide654gro$raw_654_2023-5525.167
wide654gro$nojump_2023 <- wide654gro$jumpfree_654_2023-39042.49

wide654gro$unadjusted_2024 <- wide654gro$raw_654_2024-4555.250
wide654gro$nojump_2024   <- wide654gro$jumpfree_654_2024-48766.91


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long654gro<- wide654gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long654gro_unadjusted<-long654gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long654gro_nojump<-long654gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))

#ggplot with each year a separate line: unadjusted
unadjusted654_factor_yr <- 
  ggplot(long654gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 654) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted654_factor_yr.png"), 
       plot = unadjusted654_factor_yr)


#ggplot with each year a separate line: jump-free
nojump654_factor_yr<-
  ggplot(long654gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 654) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump654_factor_yr.png"), 
       plot = nojump654_factor_yr)

unadjusted654_factor_yr
nojump654_factor_yr
```

##### 655

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d655$raw_655 <- jump_free_d655$hourly_avg655_raw 
jump_free_d655$jumpfree_655 <- jump_free_d655$hourly_avg655_0jump

wide655gro<-jump_free_d655|>
  select(c(raw_655,jumpfree_655,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_655,jumpfree_655),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide655gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide655gro$unadjusted_2019<-wide655gro$raw_655_2019-3178.333
wide655gro$nojump_2019<-wide655gro$jumpfree_655_2019-3178.333

wide655gro$unadjusted_2020 <- wide655gro$raw_655_2020-10151.67
wide655gro$nojump_2020     <- wide655gro$jumpfree_655_2020-10151.67

wide655gro$unadjusted_2021 <- wide655gro$raw_655_2021-5935.000
wide655gro$nojump_2021      <- wide655gro$jumpfree_655_2021-20665.89

wide655gro$unadjusted_2022 <- wide655gro$raw_655_2022-2174.833
wide655gro$nojump_2022  <- wide655gro$jumpfree_655_2022-32578.31

wide655gro$unadjusted_2023 <- wide655gro$raw_655_2023-2467.000
wide655gro$nojump_2023 <- wide655gro$jumpfree_655_2023-40936.77

wide655gro$unadjusted_2024 <- wide655gro$raw_655_2024-2475.750
wide655gro$nojump_2024   <- wide655gro$jumpfree_655_2024-47955.19


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long655gro<- wide655gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long655gro_unadjusted<-long655gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long655gro_nojump<-long655gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))


#ggplot with each year a separate line: unadjusted
unadjusted655_factor_yr <- 
  ggplot(long655gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 655) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted655_factor_yr.png"), 
       plot = unadjusted655_factor_yr)


#ggplot with each year a separate line: jump-free
nojump655_factor_yr<-
  ggplot(long655gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 655) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump655_factor_yr.png"), 
       plot = nojump655_factor_yr)

unadjusted655_factor_yr
nojump655_factor_yr
```

##### 656

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d656$raw_656 <- jump_free_d656$hourly_avg656_raw 
jump_free_d656$jumpfree_656 <- jump_free_d656$hourly_avg656_0jump

wide656gro<-jump_free_d656|>
  select(c(raw_656,jumpfree_656,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_656,jumpfree_656),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide656gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide656gro$unadjusted_2019<-wide656gro$raw_656_2019-3566.000
wide656gro$nojump_2019<-wide656gro$jumpfree_656_2019-3566.000

wide656gro$unadjusted_2020 <- wide656gro$raw_656_2020-9632.000
wide656gro$nojump_2020     <- wide656gro$jumpfree_656_2020-8592.667

wide656gro$unadjusted_2021 <- wide656gro$raw_656_2021-4416.250
wide656gro$nojump_2021      <- wide656gro$jumpfree_656_2021-17759.08

wide656gro$unadjusted_2022 <- wide656gro$raw_656_2022-2868.750
wide656gro$nojump_2022  <- wide656gro$jumpfree_656_2022-27458.50

wide656gro$unadjusted_2023 <- wide656gro$raw_656_2023-4333.167
wide656gro$nojump_2023 <- wide656gro$jumpfree_656_2023-34599.63

wide656gro$unadjusted_2024 <- wide656gro$raw_656_2024-1968.000
wide656gro$nojump_2024   <- wide656gro$jumpfree_656_2024-40281.13


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long656gro<- wide656gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long656gro_unadjusted<-long656gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long656gro_nojump<-long656gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))


#ggplot with each year a separate line: unadjusted
unadjusted656_factor_yr <- 
  ggplot(long656gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 656) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted656_factor_yr.png"), 
       plot = unadjusted656_factor_yr)


#ggplot with each year a separate line: jump-free
nojump656_factor_yr<-
  ggplot(long656gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 656) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump656_factor_yr.png"), 
       plot = nojump656_factor_yr)

unadjusted656_factor_yr
nojump656_factor_yr
```

##### 860

```{r}
#First seelct desired columns from jump_free df and pivot so every year has its own columm
jump_free_d860$raw_860 <- jump_free_d860$hourly_avg860_raw 
jump_free_d860$jumpfree_860 <- jump_free_d860$hourly_avg860_0jump

wide860gro<-jump_free_d860|>
  select(c(raw_860,jumpfree_860,yr,ts,DOY))|>
  pivot_wider(
    names_from = yr, 
    values_from = c(raw_860,jumpfree_860),
    names_sep="_")|>
  filter(DOY >= 91, DOY <= 274)

#normalizing columns by getting each year to start at 0 relative to the start of the X axis, e.g., wide860gro normalized in 2019 would have whatever value at DOY 91 = 0 by searching the wide df for "2019-04-01" and recording the value associated with "2019-04-01 00:00:00"

wide860gro$unadjusted_2019<-wide860gro$raw_860_2019
wide860gro$nojump_2019<-wide860gro$jumpfree_860_2019
#First reading May 29 so not normalized

wide860gro$unadjusted_2020 <- wide860gro$raw_860_2020-4410.167
wide860gro$nojump_2020     <- wide860gro$jumpfree_860_2020-4410.167

wide860gro$unadjusted_2021 <- wide860gro$raw_860_2021-11749.50
wide860gro$nojump_2021      <- wide860gro$jumpfree_860_2021-10942.56

wide860gro$unadjusted_2022 <- wide860gro$raw_860_2022-3391.000
wide860gro$nojump_2022  <- wide860gro$jumpfree_860_2022-17178.31

wide860gro$unadjusted_2023 <- wide860gro$raw_860_2023-5053.750
wide860gro$nojump_2023 <- wide860gro$jumpfree_860_2023-24277.02

wide860gro$unadjusted_2024 <- wide860gro$raw_860_2024-10681.33
wide860gro$nojump_2024   <- wide860gro$jumpfree_860_2024-29904.60


#combining back into 1 long df - note cols = code, which distingushes raw and jump-free data 
long860gro<- wide860gro |>  pivot_longer(
   cols = starts_with(c("unadjusted_","nojump_")),  
   names_to = c("adjustment"),
    values_to = "growth")|>
  separate(adjustment, into = c("adjustment", "yr"), sep = "_")

long860gro_unadjusted<-long860gro|>filter(
      adjustment=="unadjusted")|>select(
    c(ts,DOY,yr,growth))

long860gro_nojump<-long860gro|>filter(
      adjustment=="nojump")|>select(
    c(ts,DOY,yr,growth))

#ggplot with each year a separate line: unadjusted
unadjusted860_factor_yr <- 
  ggplot(long860gro_unadjusted, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 860) Raw Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Raw Normalized-1 Annual Line/unadjusted860_factor_yr.png"), 
       plot = unadjusted860_factor_yr)


#ggplot with each year a separate line: jump-free
nojump860_factor_yr<-
  ggplot(long860gro_nojump, aes(
  x = DOY, y = growth, color = as.factor(yr)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = 176, linetype = "dashed", color = "black")+
  theme_minimal()+
  labs(title="2° DF (ID 860) Jump-Free Circumference Change: April1-Oct 1",
      x="Day of Year",
      y="Stem Circumference Change (µm)",
      color="Legend") + 
  scale_color_manual(values = HD_colors) + 
  guides(color=guide_legend(override.aes = list(
            alpha = 1, size = 5))) +
  annotate("text",x=180,y=-500,label="Onset of Heat Dome",color="black")
ggsave(here("Plots/Jump-Free Normalized-1 Annual Line/nojump860_factor_yr.png"), 
       plot = nojump860_factor_yr)

unadjusted860_factor_yr
nojump860_factor_yr
```

# Step 2.1 (optional): Adjusting temporal resolution via dendro.resample

```         
dendro.resample(df, by, value) 
```

### Explanation of options

+------------+--------------------------------------------------------------------------------------------+
| `df`       | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`.  |
+------------+--------------------------------------------------------------------------------------------+
| `by`       | either *H, D, W* or *M* to resample data into hourly, daily, weekly or monthly resolution. |
+------------+--------------------------------------------------------------------------------------------+
| `value`    | either *max, min*, *mean* or *sum* for the resampling value.                               |
+------------+--------------------------------------------------------------------------------------------+

```{r}
# To resample monthly with maximum value
resample_M<-dendro.resample(df=nepa17[,1:2], by='M', value='max')
head(resample_M,10)
#example output
#         Time       T2
#1  2017-01-01 63.90855
#2  2017-02-01 64.88705
```

# Step 2.2 (optional): Extracting specific values with dendro.truncate

```         
dendro.truncate(df, CalYear, DOY) 
```

### Explanation of options

+-----------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`      | dataframe with the first column named date and time in the format `yyyy-mm-dd HH:MM:SS`.                                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
+-----------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `CalYear` | desired year of calculation.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
+-----------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `DOY`     | If we provide an array instead of a single value for `CalYear` and a single value for `DOY`, it truncates data from the `DOY` of the first `CalYear` to the same `DOY` of the second `CalYear`. Conversely, if we provide one value for `CalYear` and an array of two elements for `DOY` truncates the data form first `DOY` to second `DOY` within the same `CalYear`. Finally, if we provide an array with two values for both `DOY` and `CalYear`, it truncates data from the first `DOY` of the first `CalYear` to the second `DOY` of second `CalYear`. |
+-----------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#Example extracting data from doy 20 to 50 in 2017.
#trunc1<-dendro.truncate(df=nepa, CalYear=2017, DOY=c(20,50))
#head(trunc1,10)
#                 Time       T2       T3
#1  2017-01-20 00:00:00 64.92424 48.87772
#...10 2017-01-20 09:00:00 64.96733 48.88450
```

# Step 3.1: Phase differentiation

```         
phase.sc(df, TreeNum, smoothing = NULL) 
```

### Explanation of options for stem cycle approach

three phases in this approach: 1) Shrinkage, when the dendrometer reading is less than previous reading, 2) Expansion, when current reading is more than previous reading and 3) Increment, when current reading is higher than the previous maximum. Output can be plotted with code like this:

**plot_SC_output( SC_output, DOY, Year, cols = c("#fee8c8", "#fdbb84", "#e34a33"), phNames = c("Shrinkage", "Expansion", "Increment") )**

+-----------+----------------------------------------------------------------------------------------------------------------+
| `DOY`     | array with initial and final day for plotting. E.g. *c(a,b)*, where a = initial date and b = final date.       |
+-----------+----------------------------------------------------------------------------------------------------------------+
| `cols`    | array of three elements: colour for each phases.                                                               |
+-----------+----------------------------------------------------------------------------------------------------------------+
| `phNames` | array with three elements for three different phases. Default is **"Shrinkage", "Expansion" and "Increment"**. |
+-----------+----------------------------------------------------------------------------------------------------------------+

### Phase.sc inputs:

+-------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`        | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`. It should contain data with constant temporal resolution for best results.                                                                                                                    |
+-------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum`   | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                                                                                                                                       |
+-------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `smoothing` | numerical value from 1 to 12 which indicates the length of the smoothing spline, i.e. 1 = 1 hour and 12 = 12 hours. Default is `NULL` for no smoothing.The function [`smooth.Pspline`](http://127.0.0.1:17265/help/library/pspline/help/smooth.Pspline) is used for smoothing the data. |
+-------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Outputs a list of two dataframes. The first dataframe *SC_cycle* with cyclic phases along with various statistics and the second dataframe *SC_phase* with assigned phases for each data point.The dataframe *SC_cycle* contains the beginning, end, duration, magnitude and rate of each phase. The dataframe *SC_phase* contains time and corresponding phases during that time.

#### Testing Stem-cycle approach: which method to use (settled on jump-free default=0 smoothing)

```{r}
#309 raw - no smoothing
sc_d309_S0<-phase.sc(df=d309_hourly,TreeNum = 1)
sc_d309_plot_S0<-plot_SC_output(
  sc_d309_S0,DOY = c(169,183),Year = 2021,cols = c("red","green","blue"))

#309 raw - max smoothing
sc_d309_S12<-phase.sc(df=d309_hourly,TreeNum = 1,smoothing = 12)
sc_d309_plot_S12<-plot_SC_output(
  sc_d309_S12,DOY = c(169,183),Year = 2021,cols = c("red","green","blue"))
sc_d309_plot_S0
sc_d309_plot_S12

#309 jump-free - no smoothing
sc_d309_jf_S0<-phase.sc(df=jump_free_d309,TreeNum = 1)
sc_d309_jf_plot_S0<-plot_SC_output(
  sc_d309_jf_S0,DOY = c(169,183),Year = 2021,cols = c("red","green","blue"))

#309 jump-free - max smoothing
sc_d309_jf_S12<-phase.sc(df=jump_free_d309,TreeNum = 1,smoothing = 12)
sc_d309_jf_plot_S12<-plot_SC_output(
  sc_d309_jf_S12,DOY = c(169,183),Year = 2021,cols = c("red","green","blue"))

sc_d309_jf_plot_S0
sc_d309_jf_plot_S12
```

#### Plot for Stem-cycle approach: (using jump-free default=0 smoothing)

```{r}
#309 
sc_d309<-phase.sc(df=jump_free_d309,TreeNum = 1)
sc_d309_2021plot<-plot_SC_output(
  sc_d309,DOY = c(169,183),Year = 2021,cols = c("red","green","blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d309_2021plot.png"), 
       plot = sc_d309_2021plot)

#310 
sc_d310 <- phase.sc(df = jump_free_d310, TreeNum = 1)
sc_d310_2021plot_2wks <- plot_SC_output(
  sc_d310, DOY = c(162, 190), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d310_2021plot_2wks.png"), 
       plot = sc_d310_2021plot_2wks)

#311 - june 11 - july 9
sc_d311 <- phase.sc(df = jump_free_d311, TreeNum = 1)
sc_d311_2021plot_2wks <- plot_SC_output(
  sc_d311, DOY = c(162, 190), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d311_2021plot_2wks.png"), 
       plot = sc_d311_2021plot_2wks)

#646 
sc_d646 <- phase.sc(df = jump_free_d646, TreeNum = 1)
sc_d646_2021plot <- plot_SC_output(
  sc_d646, DOY = c(169, 183), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d646_2021plot.png"), 
       plot = sc_d646_2021plot)

#648 
sc_d648 <- phase.sc(df = jump_free_d648, TreeNum = 1)
sc_d648_2023plot <- plot_SC_output(
  sc_d648, DOY = c(169, 183), Year = 2023, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d648_2023plot.png"), 
       plot = sc_d648_2023plot)

#651 
sc_d651 <- phase.sc(df = jump_free_d651, TreeNum = 1)
sc_d651_2023plot <- plot_SC_output(
  sc_d651, DOY = c(169, 183), Year = 2023, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d651_2023plot.png"), 
       plot = sc_d651_2023plot)

#652 
sc_d652 <- phase.sc(df = jump_free_d652, TreeNum = 1)
sc_d652_2023plot <- plot_SC_output(
  sc_d652, DOY = c(169, 183), Year = 2023, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d652_2023plot.png"), 
       plot = sc_d652_2023plot)

#654
sc_d654 <- phase.sc(df = jump_free_d654, TreeNum = 1)
sc_d654_2021plot <- plot_SC_output(
  sc_d654, DOY = c(169, 183), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d654_2021plot.png"), 
       plot = sc_d654_2021plot)

#655 
sc_d655 <- phase.sc(df = jump_free_d655, TreeNum = 1)
sc_d655_2021plot <- plot_SC_output(
  sc_d655, DOY = c(169, 183), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d655_2021plot.png"), 
       plot = sc_d655_2021plot)

#656
sc_d656 <- phase.sc(df = jump_free_d656, TreeNum = 1)
sc_d656_2021plot <- plot_SC_output(
  sc_d656, DOY = c(169, 183), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d656_2021plot.png"), 
       plot = sc_d656_2021plot)

#860
sc_d860 <- phase.sc(df = jump_free_d860, TreeNum = 1)
sc_d860_2021plot_2wks <- plot_SC_output(
  sc_d860, DOY = c(162, 190), Year = 2021, cols = c("red", "green", "blue"))
ggsave(here("Plots/Stem-Cycle Plots/sc_d860_2021plot_2wks.png"), 
       plot = sc_d860_2021plot_2wks)

sc_d309_2021plot
sc_d310_2021plot
sc_d311_2021plot
sc_d646_2021plot
sc_d648_2023plot
sc_d651_2023plot
sc_d652_2023plot
sc_d654_2021plot
sc_d655_2021plot
sc_d656_2021plot
sc_d860_2021plot
```

#### Stem-Cycle approach analyses from table output above

```{r}
sc_d309_df<-sc_d309[["SC_cycle"]]

#phase start time vs magnitude, diff color for phase
sc_d309_MagPlot<-
  ggplot(sc_d309_df, aes(
  x = Start, y = Magnitude, color = as.factor(Phases)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = vline_HD, linetype = "dashed", color = "black")+
  theme_minimal()
sc_d309_MagPlot

#phase start time vs magnitude, increment phase only
sc_d309_increment<-sc_d309_df|>filter(Phases=="3")
sc_d309_increment_MagPlot<-
  ggplot(sc_d309_df, aes(
  x = Start, y = Magnitude))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = vline_HD, linetype = "dashed", color = "black")+
  theme_minimal()
sc_d309_increment_MagPlot

#phases vs magnitude 
ggplot(sc_d309_df, aes(
  x = Phases, y = Magnitude, color = as.factor(Phases)))+ 
  geom_point(alpha=0.5)


sc_d310_df<-sc_d310[["SC_cycle"]]|>filter(Phases=="3")

#phase start time vs rate, increment phase only
sc_d310_RatePlot3<-
  ggplot(sc_d310_df, aes(
  x = Start, y = rate))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = vline_HD, linetype = "dashed", color = "black")+
  theme_minimal()
sc_d310_RatePlot3


zg_d310_df<-zg_d310[["ZG_cycle"]]
#phase start time vs magnitude, diff color for phase
zg_d310_MagPlot<-
  ggplot(zg_d310_df, aes(
  x = Start, y = Magnitude, color = as.factor(Phases)))+ 
  geom_point(alpha=0.5)+ 
  geom_vline(xintercept = vline_HD, linetype = "dashed", color = "black")+
  theme_minimal()
zg_d310_MagPlot
```

### Inputs for zero-growth approach

+-----------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `df`      | dataframe with first column containing date and time in the format `yyyy-mm-dd HH:MM:SS`. It should contain data with constant temporal resolution for best results. |
+-----------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `TreeNum` | numerical value indicating the tree to be analysed. E.g. '1' refers to the first dendrometer data column in *df*.                                                    |
+-----------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------+

Outputs a list of two dataframes. The first dataframe *ZG_cycle* contains the cyclic phases along with various statistics and the second dataframe *ZG_phase* with assigned phases for each data point.

Output can be plotted with th following code:

```         
plot_ZG_output(ZG_output, DOY, Year) 
```

### Explanation of options for zero-growth approach

+-------------+----------------------------------------------------------------------------------------------------------+
| `ZG_output` | list the output of `phase.zg` function.                                                                  |
+-------------+----------------------------------------------------------------------------------------------------------+
| `DOY`       | array with initial and final day for plotting. E.g. *c(a,b)*, where a = initial date and b = final date. |
+-------------+----------------------------------------------------------------------------------------------------------+
| `Year`      | array for indicating year for plotting.                                                                  |
+-------------+----------------------------------------------------------------------------------------------------------+

Two subplots with GRO and TWD respectively.

Since DOY 176 is HD and these plots prefer smaller time scales, we are going to do +/- 1 week (DOY 169-183)

#### Zero-growth approach: (using jump-free hourly data)

```{r}
#309 - plot looks weird
zg_d309<-phase.zg(df=jump_free_d309,TreeNum = 1)
zg_d309_2021plot<-plot_ZG_output(zg_d309,DOY = c(169,183),Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d309_2021plot.png"), 
       plot = zg_d309_2021plot)

#310
zg_d310<-phase.zg(df=jump_free_d310,TreeNum = 1)
zg_d310_2021plot_2wks<-plot_ZG_output(zg_d310,DOY = c(162,190),Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d310_2021plot_2wks.png"), 
       plot = zg_d310_2021plot_2wks)

#311
zg_d311 <- phase.zg(df = jump_free_d311, TreeNum = 1)
zg_d311_2021plot_2wks <- plot_ZG_output(zg_d311, DOY = c(162,190), Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d311_2021plot_2wks.png"), 
       plot = zg_d311_2021plot_2wks)

#646
zg_d646 <- phase.zg(df = jump_free_d646, TreeNum = 1)
zg_d646_2021plot <- plot_ZG_output(zg_d646, DOY = c(169,183), Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d646_2021plot.png"), 
       plot = zg_d646_2021plot)

#648 - installed 2023
zg_d648 <- phase.zg(df = jump_free_d648, TreeNum = 1)
zg_d648_2023plot <- plot_ZG_output(zg_d648, DOY = c(169,183), Year = 2023)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d648_2023plot.png"), 
       plot = zg_d648_2023plot)

#651 - installed 2023
zg_d651 <- phase.zg(df = jump_free_d651, TreeNum = 1)
zg_d651_2023plot <- plot_ZG_output(zg_d651, DOY = c(169,183), Year = 2023)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d651_2023plot.png"), 
       plot = zg_d651_2023plot)

#652 - installed 2023
zg_d652 <- phase.zg(df = jump_free_d652, TreeNum = 1)
zg_d652_2023plot <- plot_ZG_output(zg_d652, DOY = c(169,183), Year = 2023)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d652_2023plot.png"), 
       plot = zg_d652_2023plot)

#654
zg_d654 <- phase.zg(df = jump_free_d654, TreeNum = 1)
zg_d654_2021plot <- plot_ZG_output(zg_d654, DOY = c(169,183), Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d654_2021plot.png"), 
       plot = zg_d654_2021plot)

#655
zg_d655 <- phase.zg(df = jump_free_d655, TreeNum = 1)
zg_d655_2021plot <- plot_ZG_output(zg_d655, DOY = c(169,183), Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d655_2021plot.png"), 
       plot = zg_d655_2021plot)

#656
zg_d656 <- phase.zg(df = jump_free_d656, TreeNum = 1)
zg_d656_2021plot <- plot_ZG_output(zg_d656, DOY = c(169,183), Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d656_2021plot.png"), 
       plot = zg_d656_2021plot)

#860
zg_d860 <- phase.zg(df = jump_free_d860, TreeNum = 1)
zg_d860_2021plot_2wks <- plot_ZG_output(zg_d860, DOY = c(162,190), Year = 2021)
#saving plot
ggsave(here("Plots/Zero-Growth Plots/zg_d860_2021plot_2wks.png"), 
       plot = zg_d860_2021plot_2wks)

zg_d309_2021plot
zg_d310_2021plot
zg_d311_2021plot
zg_d646_2021plot
zg_d648_2023plot
zg_d651_2023plot
zg_d652_2023plot
zg_d654_2021plot
zg_d655_2021plot
zg_d656_2021plot
zg_d860_2021plot
```

# Step 3.2: Daily stats: Calculating and making amplitude plots (max-min) from daily.data using jump-free datay

See "Code to run at launch" section for details on daily.data vs daily.data.patched

```         
daily.data(df, TreeNum) 
```

This function calculates various statistics of dendrometer data on a daily basis. The daily statistics includes the daily maximum and minimum with their corresponding times and daily amplitude (difference between daily maximum and minimum). See King et al. (2013) for details. Output is a dataframe with the daily statistics of the dendrometer data and Remarks = "\*" if Time_max \> Time_min otherwise "".

## Amplitude over time

```{r}
# 309
#adding new column to clearly define Remarks for plotting 
daily_stats309<-daily.data(df=jump_free_d309, TreeNum=1)|>
   mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))|>
#clarifying sign of amplitude so we don't have to plot by Remark
  mutate(amplitude_signed = case_when(
    Remarks == "*" ~ amplitude,
    Remarks == ""  ~ -amplitude))

|>
#adding approximate season halves with summer solstice (june 20=DOY171) separating the two in line with D'Orangeville's 50% of growth separator being around this time (June 19) - chronological halfway point would be July 1=DOY182, however (non leap year scale)
    mutate(season_half = case_when(
    doy >= 91 & doy <= 171 ~ "First Half",
    doy >= 172 & doy <= 274 ~ "Second Half"))
  
  
#setting up for DOY 
daily_stats309$yr <- format(
  daily_stats309$DATE, "%Y")
daily_stats309$yr<-as.factor(daily_stats309$yr)
daily_stats309$DOY<-yday(daily_stats309$DATE)

  mutate(season_half = case_when(
    doy >= 100 & doy <= 200 ~ "First Half",
    doy >= 201 & doy <= 300 ~ "Second Half",
    TRUE ~ NA_character_
  ))

#amplitude plot all years and distinguishing remarks
ggplot(
  daily_stats309,aes(x=as.POSIXct(DATE),y=amplitude_signed,color=Remarks_Description))+
  geom_point()+  
  geom_vline(xintercept = as.numeric(vline_HD), 
             linetype = "solid", color = "black")+
  labs(title="OG DF (ID 309) Daily Circ. Change",
      x="Date",y="Daily Stem Circumference Change (µm)",
      color="Time of Max. Circ. Value") + 
  guides(color=guide_legend(override.aes = list(alpha = 1, size = 5)))+
  theme_minimal()+
  annotate("text",
           x=as.POSIXct(vline_HD),y=-50,label="Onset of Heat Dome",color="black")



# 310
#adding new column to clearly define Remarks for plotting 
daily_stats310 <- daily.data(df = jump_free_d310, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))|>
#add column for daily max change
    mutate(max_diff = Max - lag(Max))|>
#clarifying sign of amplitude so we don't have to plot by Remark
  mutate(amplitude_signed = case_when(
    Remarks == "*" ~ amplitude,
    Remarks == ""  ~ -amplitude))

#setting up for DOY annual analysis if desired
daily_stats310$yr <- format(daily_stats310$DATE, "%Y")
daily_stats310$yr <- as.factor(daily_stats310$yr)
daily_stats310$DOY <- yday(daily_stats310$DATE)

ggplot(daily_stats310,aes(x=amplitude_signed))+geom_histogram()+theme_minimal()+
  labs(title = "OG DF (ID 310) Histogram of Circ. Amp",
       x = "Daily Growth Ampltiude", y = NULL)

#plotting changes in daily max
ggplot(daily_stats310, aes(x = as.POSIXct(DATE), y = max_diff)) +
  geom_line() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "OG DF (ID 310) Changes in Daily Max. Circ.",
       x = "Date", y = "(Max Day N) - (Max Day N-1)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = 300, label = "Onset of Heat Dome", color = "black") +   
  xlim(as.POSIXct("2021-06-11"), as.POSIXct("2021-07-09"))

#amplitude plot all years and distinguishing remarks
ggplot(daily_stats310, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "OG DF (ID 310) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")



# 311
daily_stats311 <- daily.data(df = jump_free_d311, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))|>
#clarifying sign of amplitude so we don't have to plot by Remark
  mutate(amplitude_signed = case_when(
    Remarks == "*" ~ amplitude,
    Remarks == ""  ~ -amplitude))

daily_stats311$yr <- format(daily_stats311$DATE, "%Y")
daily_stats311$yr <- as.factor(daily_stats311$yr)
daily_stats311$DOY <- yday(daily_stats311$DATE)

ggplot(daily_stats311, aes(x = as.POSIXct(DATE), y = amplitude_signed, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "OG DF (ID 311) Daily Circ. Amplitude",
       x = "Date", y = "Daily Stem Circumference Amplitude (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")



# 646
daily_stats646 <- daily.data(df = jump_free_d646, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats646$yr <- format(daily_stats646$DATE, "%Y")
daily_stats646$yr <- as.factor(daily_stats646$yr)
daily_stats646$DOY <- yday(daily_stats646$DATE)

ggplot(daily_stats646, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "OG WH (ID 646) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")



# 648
daily_stats648 <- daily.data(df = jump_free_d648, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats648$yr <- format(daily_stats648$DATE, "%Y")
daily_stats648$yr <- as.factor(daily_stats648$yr)
daily_stats648$DOY <- yday(daily_stats648$DATE)

ggplot(daily_stats648, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +
  labs(title = "OG WH (ID 648) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal()



# 651
daily_stats651 <- daily.data(df = jump_free_d651, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats651$yr <- format(daily_stats651$DATE, "%Y")
daily_stats651$yr <- as.factor(daily_stats651$yr)
daily_stats651$DOY <- yday(daily_stats651$DATE)

ggplot(daily_stats651, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  labs(title = "OG WRC (ID 651) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal()



# 652
daily_stats652 <- daily.data(df = jump_free_d652, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats652$yr <- format(daily_stats652$DATE, "%Y")
daily_stats652$yr <- as.factor(daily_stats652$yr)
daily_stats652$DOY <- yday(daily_stats652$DATE)

ggplot(daily_stats652, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  labs(title = "OG WH (ID 652) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() 



# 654
daily_stats654 <- daily.data(df = jump_free_d654, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats654$yr <- format(daily_stats654$DATE, "%Y")
daily_stats654$yr <- as.factor(daily_stats654$yr)
daily_stats654$DOY <- yday(daily_stats654$DATE)

ggplot(daily_stats654, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "2° DF (ID 654) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")



# 655
daily_stats655 <- daily.data(df = jump_free_d655, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats655$yr <- format(daily_stats655$DATE, "%Y")
daily_stats655$yr <- as.factor(daily_stats655$yr)
daily_stats655$DOY <- yday(daily_stats655$DATE)

ggplot(daily_stats655, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "2° DF (ID 655) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")




# 656
daily_stats656 <- daily.data(df = jump_free_d656, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))

daily_stats656$yr <- format(daily_stats656$DATE, "%Y")
daily_stats656$yr <- as.factor(daily_stats656$yr)
daily_stats656$DOY <- yday(daily_stats656$DATE)

ggplot(daily_stats656, aes(x = as.POSIXct(DATE), y = amplitude, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "2° DF (ID 656) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")



# 860
daily_stats860 <- daily.data(df = jump_free_d860, TreeNum = 1) |>
  mutate(Remarks_Description = case_when(
    Remarks == ""  ~ "Time_max < Time_min",
    Remarks == "*" ~ "Time_max > Time_min"))|>
#add column for daily max change
    mutate(max_diff = Max - lag(Max))

#setting up for DOY annual analysis if desired
daily_stats860$yr <- format(daily_stats860$DATE, "%Y")
daily_stats860$yr <- as.factor(daily_stats860$yr)
daily_stats860$DOY <- yday(daily_stats860$DATE)

daily_stats860<-daily_stats860|>
#clarifying sign of amplitude so we don't have to plot by Remark
  mutate(amplitude_signed = case_when(
    Remarks == "*" ~ amplitude,
    Remarks == ""  ~ -amplitude))


ggplot(daily_stats860,aes(x=amplitude_signed))+geom_histogram()+theme_minimal()+
  labs(title = "2° DF (ID 860) Histogram of Circ. Amp",
       x = "Daily Growth Ampltiude", y = NULL)

ggplot(daily_stats860,aes(x=amplitude))+geom_histogram()

#plotting changes in daily max
ggplot(daily_stats860, aes(x = as.POSIXct(DATE), y = max_diff)) +
  geom_line() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "2° DF (ID 860) Changes in Daily Max. Circ.",
       x = "Date", y = "(Max Day N) - (Max Day N-1)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal() +
  annotate("text", x = as.POSIXct(vline_HD), y = 300, label = "Onset of Heat Dome", color = "black") +   
  xlim(as.POSIXct("2021-06-11"), as.POSIXct("2021-07-09"))

```

## Daily maximum circumferential change

### 309

```{r}
#daily_stats860|>MAX DAY 2 MINUS MAX DAY 1

ggplot(daily_stats311, aes(x = as.POSIXct(DATE), y = Max, color = Remarks_Description)) +
  geom_point() +  
  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "black") +
  labs(title = "2° DF (ID 311) Daily Circ. Change",
       x = "Date", y = "Daily Stem Circumference Change (µm)",
       color = "Time of Max. Circ. Value") + 
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5))) +
  theme_minimal()  +
  annotate("text", x = as.POSIXct(vline_HD), y = -50, label = "Onset of Heat Dome", color = "black")

```

# Step 3.3: Relative Dry Growth During Climate Adverse events with clim.twd

```         
clim.twd(df,   Clim,   dailyValue = "max",   thresholdClim = "<10",   thresholdDays = ">5",   showPlot = TRUE ) 
```

```{r}
library(dendRoAnalyst)
data(gf_nepa17)
data(ktm_rain17)
relative_dry_growth<-clim.twd(df=gf_nepa17, Clim=ktm_rain17, dailyValue='max', showPlot=TRUE)
1

```

### Explanation of options

+-----------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `Clim`          | dataframe with the first column containing `Date in yyyy-mm-dd` and second column containing corresponding climate data.                                                                                                                                |
+-----------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `dailyValue`    | either *'max', 'min', 'mean'*, or *'sum'* for selecting the daily resampled value. Default is *'max'*. See [`dendro.resample`](http://127.0.0.1:17265/help/library/dendRoAnalyst/help/dendro.resample) for details.                                     |
+-----------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `thresholdClim` | string, the theshold for the respective climatic parameter. E.g. if climatic data is precipitation then days, where precipitation is below or equal to this value, are considered as adverse climate. Dafault is '\<10'.                                |
+-----------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `thresholdDays` | string, the minimum number of consecutive adverse days to be considered for analysis. For example, `thresholdDays`=2 means the relative radial/circumferential change is calculated for adverse periods lasting for more than 2 days. Default is '\>5'. |
+-----------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#raw309
 library(dendRoAnalyst)
relative_dry_growth309<-clim.twd(
  df=d309_daily, 
  Clim=filtered_P230_rain_daily, 
  dailyValue='mean', 
  thresholdClim = "<2.5",
  thresholdDays = ">10",
  showPlot=TRUE)

#jump-free309
relative_dry_growth_jf309<-clim.twd(
  df=jump_free_d309, 
  Clim=filtered_P230_rain_daily, 
  dailyValue='mean', 
  thresholdClim = "<2.5",
  thresholdDays = ">10",
  showPlot=TRUE)
```

# One curve per dendro: raw vs jump-free

#### Stat Comparisons:

```{r}
#dataframe with just the AVG dendrometer value
JustDendros<-data_prep524|>select(c(
 Date,DEND309_AVG_4200_0_01,DEND310_AVG_150_0_01,DEND311_AVG_150_0_01,DEND646_AVG_150_0_01,DEND648_AVG_150_0_01,DEND651_AVG_150_0_01,DEND652_AVG_150_0_01,DEND654_AVG_150_0_01,DEND655_AVG_150_0_01,DEND656_AVG_150_0_01,DEND860_AVG_150_0_01))

JustDendros_long<-

# Assuming df_long is your long format DataFrame
#lookup = pd.DataFrame({
#    'names_to': ['D1', 'D2', 'D3', 'D4', 'D5', 'D6', 'D7', 'D8', 'D9', 'D10'],
#    'species': ['Oak', 'Pine', 'Maple', 'Oak', 'Spruce', 'Fir', 'Birch', 'Elm', #'Cedar', 'Ash'],
#    'age_class': ['Young', 'Mature', 'Old', 'Young', 'Mature', 'Old', 'Young', #'Mature', 'Old', 'Young']
#})

#df_long = df_long.merge(lookup, on='names_to', how='left')


#df_long <- df %>%
#  pivot_longer(cols = -time, names_to = "variable", values_to = "value")


#p <- ggplot(df_long, aes(x = time, y = value, color = variable, linetype = #variable)) +
#  geom_line() +
#  geom_point(aes(shape = variable)) +
#  theme_minimal()


```

```{r}
#maybe we come back to adjust this after perhaps dentifying start of actual growing szn
growing_seasons_default <- data.frame(
  xmin = as.POSIXct(c("2018-04-01 00:00","2019-04-01 00:00","2020-04-01 00:00","2021-04-01 00:00","2022-04-01 00:00","2023-04-01 00:00","2024-04-01 00:00")),
  xmax = as.POSIXct(c("2018-10-01 00:00","2019-10-01 00:00","2020-10-01 00:00","2021-10-01 00:00","2022-10-01 00:00","2023-10-01 00:00","2024-10-01 00:00")),
  fill = "Approx. Growing Szn")
vline_HD<-as.POSIXct("2021-06-25 00:00")

ggplot(
  jump_free_d309,aes(x=TIME,y=value))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 309) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)


jumpfree_310_1curve<-ggplot(
  jump_free_d310,aes(x=TIME,y=hourly_avg310))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 310) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)
d310_hourly$hour_num<-as.numeric(as.character(d310_hourly$hour))

raw_310_1curve<-ggplot(
  d310_hourly,aes(x=hour_num,y=hourly_avg310))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 310) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)
gridExtra::grid.arrange(jumpfree_310_1curve,raw_310_1curve)

ggplot(
  jump_free_d311,aes(x=TIME,y=hourly_avg311))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 311) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d646,aes(x=TIME,y=hourly_avg646))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG WH (ID 646) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d648,aes(x=TIME,y=hourly_avg648))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG WH (ID 648) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d652,aes(x=TIME,y=hourly_avg652))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG WH (ID 652) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d654,aes(x=TIME,y=hourly_avg654))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="Secondary DF (ID 654) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d655,aes(x=TIME,y=hourly_avg655))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="Secondary DF (ID 655) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d656,aes(x=TIME,y=hourly_avg656))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="Secondary DF (ID 656) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)

ggplot(
  jump_free_d860,aes(x=TIME,y=hourly_avg860))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="Secondary DF (ID 860) Adjusted Stem Circumference Change",
      x="Date",y="Stem Circumference Change (µm)") +theme_minimal()+  geom_vline(xintercept = as.numeric(vline_HD), linetype = "solid", color = "red")+ geom_rect(data = growing_seasons_default, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            alpha = 0.1)
```

Raw plots were made in "ecomatik.rmd"

```{r}
#Apr1-Oct1 D309
jump_free_d309$yr<-format(jump_free_d309$TIME, "%Y")
jump_free_d309$yr<-as.factor(jump_free_d309$yr)
jump_free_d309$DOY<-yday(jump_free_d309$TIME)
ggplot(
  jump_free_d309,aes(x=DOY,y=value,color=factor(yr)))+geom_point(
    size=1.25,alpha=0.5)+labs(
      title="OG DF (ID 309) Adjusted Stem Circumference Change",
      x="Day of Year",y="Stem Circumference Change (µm)",
      color="Legend") + scale_color_manual(
        values = custom_colors) + guides(
          color=guide_legend(override.aes = list(
            alpha = 1, size = 5)))+ geom_vline(
              xintercept = 177, linetype = "dashed", color = "black")+theme_minimal()+ylim(-2000,25500)#+xlim(91,274)
```

# Modeling annual growth with a gompertz function

```         
dm.fit.gompertz(df, CalYear, TreeNum, f_derivative = F) 
gomp_fitted<-dm.fit.gompertz(df=gf_nepa17, TreeNum = 1, CalYear=2017)
```

### Explanation of options

+----------------+------------------------------------------------------------------------------------------------------------------------+
| `CalYear`      | numeric for year of calculation. If df has more than one year, assigning CalYear truncates the data of only that year. |
+----------------+------------------------------------------------------------------------------------------------------------------------+
| `f_derivative` | logical if yes returns first derivative of gompertz curve.                                                             |
+----------------+------------------------------------------------------------------------------------------------------------------------+

```{r}
jump_free_d309_noNA<-jump_free_d309|>filter(!is.na(jump_free_d309$value))
gompertz309<-dm.fit.gompertz(df=jump_free_d309_noNA,TreeNum=1,CalYear = 2021)
ggplot(gompertz309,aes(x=TIME,y=value))+geom_point()
```

# Running correlation between dendrometer and climate data.

Calculates running correlation between dendrometer series and provided climate variables. Users can choose methods such as pearson, kendall and spearman, see `?cor.test` for further information.

```         
mov.cor.dm(df, Clim, TreeNum, win_size, cor_method = "pearson") 
out_corr<-mov.cor.dm(df=gf_nepa17, Clim=ktm_rain17, TreeNum=1, win_size=21) 
```

### Explanation of options

+--------------+------------------------------------------------------------------------------------------------------------------+
| `win_size`   | numerical, the running days windows of which the correlation is to be calculated. The minimum value is 18.       |
+--------------+------------------------------------------------------------------------------------------------------------------+
| `cor_method` | string, method to be applied during correlation calculation. One of the following: pearson, kendall and spearman |
+--------------+------------------------------------------------------------------------------------------------------------------+

### Output can be plotted with the following:

plot_mov.cor(mov.cor.output, sig.only = T, ci = 0.95, clim_vars = "all")\
Example code:\
out_corr\<-mov.cor.dm(df=gf_nepa17, Clim=ktm_rain17, TreeNum=1, win_size=21) plot_mov.cor(mov.cor.output=out_corr, sig.only=TRUE, ci=0.95)

### Arguments

+------------------+----------------------------------------------------------------+
| `mov.cor.output` | list the output of `mov.cor.dm` function.                      |
+------------------+----------------------------------------------------------------+
| `sig.only`       | logical TRUE to plot only significant correlation.             |
+------------------+----------------------------------------------------------------+
| `ci`             | numeric confidence interval                                    |
+------------------+----------------------------------------------------------------+
| `clim_vars`      | array of climate variables or "all" for all climate variables. |
+------------------+----------------------------------------------------------------+

```{r}

```
