---
title: "Dend 309 full"
author: "Gabby John"
date: "`r Sys.Date()`"
output: pdf
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installing for the first time

`{#{packages <- (c("devtools","zoo","chron","dplyr","viridis", "RCurl", "DT","lubridate"))} #install.packages(setdiff(packages, rownames(installed.packages()))) #devtools::install_github("treenet/treenetproc")`

```{r}
library(treenetproc, `force = TRUE`)
library(zoo)
library(chron)
library(viridis)
library(dplyr)
library(ggplot2)
library(lubridate)
```

## step 0: identify helper functions

```{r}
left <- function(string, char){substr(string, 1,char)} 
right <- function (string, char){substr(string,nchar(string)-(char-1),nchar(string))}
col_sel<-c("cyan","darkorange","violet","blue","red","green","gold")
```

## step 1: read in dendrometer data and plot raw data
R will not open your CSV file if its title starts with a number and will instead give an error message. You can change the title so it starts with a letter character instead. 
Make sure that the NA strings reader matches the NA codes in the original CSV file, in this case "NaN". 

```{r}
dend309tsfull <- read.csv("~/research/dend309/309 summer work/309tsfull.csv", na.strings="NaN", stringsAsFactors=TRUE)
head(dend309tsfull)
```

## Step 2: fix timestamps (ts) and connect with dendrometer ID number

```{r}
dend309<-dend309tsfull
datetime<-dend309$datetime
ID = '309'
ts = as.POSIXct(dend309['datetime'][,],format="%Y-%m-%d %H:%M:%S",tz="GMT") 
dend309_L0<-cbind(dend309["value"]+0,ts,series=ID)
```

## Step 3: set parameters to plot raw data

```{r}
#First grab years string
years <-left(dend309_L0[,"ts"],4)
#To plot, set number of plots with par()
par(mfrow=c(1,1))
#Then set margins 
par(mar = c(5, 5, 5, 5))
```

## Step 4: plotting raw data for loop plots data iteratively.

```{r}
for(y in 1:length(unique(years))){
#Now we can select years
sel<-dend309_L0[which(years==unique(years)[y]),]
#But we need to handle first year first
if(y==1){
   plot(difftime(as.POSIXct(sel$ts,format="%Y-%m-%d %H:%M:%S",tz="GMT"),
  as.POSIXct(paste0(unique(years)[y],"-01-01 00:00:00"),
                  format="%Y-%m-%d %H:%M:%S",tz="GMT"),
                  units = "days"),
         sel$value,
         ylab=expression("L0 ("*mu*"m)"),
         xlab="Day of year",type="l",
         col=viridis(length(unique(years)))[y],
         xlim=c(0,365),
         ylim=c(min(dend309_L0$value,na.rm=T),
                max(dend309_L0$value,na.rm=T)),
         main=unique(dend309_L0$series))
    
    legend("topleft",
           as.character(unique(years)[-4]),
           col=rainbow(length(unique(years))),
           bty="n",pch=16)
# Now we can add other years
  }else{
    lines(difftime(as.POSIXct(sel$ts,format="%Y-%m-%d %H:%M:%S",tz="GMT"),
                   as.POSIXct(paste0(unique(years)[y],"-01-01 00:00:00"),
                              format="%Y-%m-%d %H:%M:%S",tz="GMT"),
                   units = "days"),
          sel$value,
          col=rainbow(length(unique(years)))[y],pch=16)}
}
```

## Step 5: cleaning the dendrometery data

```{r}
#Level 1 processing (time align) method 1
?treenetproc::proc_L1 
dend309_L1<- proc_L1(data_L0 = dend309_L0, reso = 60 ) 
                  #input = "wide", date_format ="%Y-%m-%d %H:%M:%S") 
head(dend309_L1)
#Level 1 processing (time align) method 2

dend309_L1<- proc_L1(data_L0 = dend309_L0, reso = 60,
                  date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT")
head(dend309_L1)

#level 1 processing of temperature data
DSCMET421_mean56m <- read.csv("~/research/DSCMET421 climate data/DSCMET421_mean56m.csv", na.strings="NaN", stringsAsFactors=TRUE)
head(DSCMET421_mean56m)

temp_data_309_L0 <- cbind(DSCMET421_mean56m,series=ID)
colnames(temp_data_309_L0) <-c('ts','value','series')

#Now we can time-align temperature data with proc_L1
temp_data_309_L1 <- proc_L1(data_L0 = temp_data_309_L0, reso = 60, 
                            date_format ="%Y-%m-%d %H:%M:%S", tz = "GMT") 
head(temp_data_309_L1)
```

Now we can move on to proc_dendro_L2, which integrates dendrometer readings with temperature, detects errors, corrects them and plots gro and twd by date.more info here: ?treenetproc::proc_dendro_L2

first plot full dataset for dendrometer. Plot will be exported to working directory as pdf

```{r}
#first plot full dataset for dendrometer. NOTE: Plot will be exported to working directory as pdf
dend309_L2 <- proc_dendro_L2(dendro_L1 = dend309_L1, temp_L1 = temp_data_309_L1, tol_out = 2, tol_jump = 3, plot = TRUE, frost_thr = 1, plot_period = "full", plot_export = TRUE, interpol = 3.5*60, frag_len = NULL, plot_name = "dend309summer_full", tz="GMT") 
graphics.off()
```

Looking at this we can see that the tolerance for outliers and tolerance for jumps were set lower (more stringent) than default value. We can plot monthly instead and look at exported pdf for greater detail.

```{r}
dend309_L2_monthly <- proc_dendro_L2(dendro_L1 = dend309_L1, temp_L1 = temp_data_309_L1, tol_out = 2, tol_jump = 3, frost_thr = 1, plot = TRUE, interpol = 3.5*60, plot_period = "monthly", plot_export = TRUE, plot_name = "dend309summer_monthly", tz="GMT") 
graphics.off()
```

We can manually correct some areas where automated data cleaning was not up to snuff.

```{r}
#info on function here: ?corr_dendro_L2

#check errors in pdf and with this subset 
dend309_L2[which((dend309_L2$flags)!="NA"),]

#highlight corrections made on the dendrometer data:
(dend309_L2[which(is.na(dend309_L2$flags)==F),])
```

##Step 5.1: Tree water deficit (TWD) as an indicator of drought stress.
```{r}
# plot minimum daily twd against day of year
par(mfrow=c(1,1))
par(mar = c(5, 5, 5, 5))

plot(1,
     1,
     ylim=c(0,max(dend309_L2$twd,na.rm=T)),
     xlim=c(100,300),
     ylab=expression("twd ("*mu*"m)"),
     xlab="Day of year",
     col="white")

for(y in c(1:length(unique(left(dend309_L2$ts,4))))){
 
   # selected year
  sel<-dend309_L2[which(left(dend309_L2$ts,4)==unique(left(dend309_L2$ts,4))[y]),]
 
   # calc twd
  twd<-suppressWarnings(aggregate(sel$twd,list(as.Date(sel$ts)),min,na.rm=T))
  twd$doy<-as.numeric(strftime(as.Date(twd$Group.1), format = "%j"))

  # clean
  twd[which(twd$x=="Inf"),"x"]<-NA

  lines(twd$doy,twd$x,col=col_sel[y],lwd=1.5)
  twd[which(is.na(twd$x)==T),"x"]<-0

}

legend("topleft",
       c(unique(left(dend309_L2$ts,4))),
       col=col_sel,
       bty="n",pch=16)

# Add growing season extent:
grow_seas_L2 <- grow_seas(dendro_L2 = dend309_L2,
                          agg_yearly=TRUE,
                          tol_seas = 0.1)


abline(v=c(mean(grow_seas_L2$gro_start),
           mean(grow_seas_L2$gro_end)),
       lty=2)
text(mean(c(mean(grow_seas_L2$gro_start),
            mean(grow_seas_L2$gro_end))),
     max(dend309_L2$twd,na.rm=T),
     "Growing season")
```



## Step 6: plot!

```{r}
ggplot((subset(dend309_L2, frost == "FALSE")), aes(x=ts))+ geom_line(aes(y=value), color = "grey70")+ geom_line(aes(y=gro_yr), color = "seagreen")+ geom_line(aes(y=twd), color = "red")
```

What if we want to visualize cumulative growth instead of growth for each year?

```{r}
#create column for separate years 
dend309_L2$yr <- substr(dend309_L2$ts,1,4)
#Turn these columns into separate dataframes for each year
dend309_2023_growth <- dend309_L2[which(dend309_L2$yr == "2023"),] 
dend309_2022_growth <- dend309_L2[which(dend309_L2$yr == "2022"),]
dend309_2021_growth <- dend309_L2[which(dend309_L2$yr == "2021"),]
dend309_2020_growth <- dend309_L2[which(dend309_L2$yr == "2020"),]
dend309_2019_growth <- dend309_L2[which(dend309_L2$yr == "2019"),]
dend309_2018_growth <- dend309_L2[which(dend309_L2$yr == "2018"),]

#"for 2023, add that years growth (gro_yr) to 2022's maximum growth for 2022, make a copy of gro_yr". 
dend309_2023_growth$gro_tot <- dend309_2023_growth$gro_yr+0
dend309_2022_growth$gro_tot <- dend309_2022_growth$gro_yr+0
dend309_2021_growth$gro_tot <- dend309_2021_growth$gro_yr+0
dend309_2020_growth$gro_tot <- dend309_2020_growth$gro_yr+5498
dend309_2019_growth$gro_tot <- dend309_2019_growth$gro_yr+5225
dend309_2018_growth$gro_tot <- dend309_2018_growth$gro_yr

#recombine both years data frames into one. 
dend309_L3<-rbind(dend309_2018_growth,dend309_2019_growth,dend309_2020_growth,dend309_2021_growth,dend309_2022_growth, dend309_2023_growth)

#Now plot!
ggplot((subset(dend309_L3, frost =="FALSE")), aes(x=ts))+ geom_line(aes(y=value), color = "grey70")+ geom_line(aes(y=gro_tot), color = "seagreen")+ geom_line(aes(y=twd), color = "red")
```

What if we just want ro look at the growing season?

```{r}
# we can calculate growing season start and end:
?treenetproc::grow_seas

# aggregate to growing season by year
grow_seas_L2 <- grow_seas(dendro_L2 = dend309_L2,agg_yearly=TRUE,tol_seas = 0.1)
knitr::kable(grow_seas_L2,
             caption = "Sample output data of the function `grow_seas`.")
```


```{r}
ggplot((subset(dend309_L3, ts < as.POSIXct("2024-03-01 00:00") & ts > as.POSIXct("2023-03-01 00:00"))), aes(x=ts))+ geom_line(aes(y=value), color = "grey70")+ geom_line(aes(y=gro_tot), color = "seagreen")+ geom_line(aes(y=twd), color = "red")+xlab("Day of Year (2023)")+ggtitle("Cumulative Growing Season Growth ID 309")
```


```{r}
#The function phase_stats() will compute rates of expansion(growth) and shrinkage (twd) and aggregate these on a daily basis. 
#More info: ?phase_stats

dend309_phase <- phase_stats(dend309_L2, plot_phase = TRUE, plot_export = TRUE, agg_daily = TRUE, tz = 'GMT' )

#These growth and shrink rates can be compared with temperature, VPD and soil moisture.
#We will likely restrict this analysis to the growing season.
#some preliminary analysis/exploration using this

ggplot(subset(dend309_phase, doy > 100 & doy < 250), aes(x=doy))+ geom_point(aes(y = shrink_amp, colour = "Shrinkage" ))+ geom_point(aes(y = exp_amp, colour ="Expansion"))+ ylab("Amplitude")

ggplot(subset(dend309_phase, doy > 100 & doy < 250), aes(x=doy))+ geom_point(aes(y = shrink_slope, colour = "Shrinkage" ))+ geom_point(aes(y = exp_slope, colour ="Expansion"))+ ylab("Slope")

ggplot(subset(dend309_phase, doy > 100 & doy < 250), aes(x=shrink_amp, y = exp_amp))+ geom_point()+ ylab(" Expansion amplitude")+ xlab(" Shrinkage amplitude")

ggplot(subset(dend309_phase, doy > 100 & doy < 250), aes(x=shrink_slope, y = exp_slope))+ geom_point()+ ylab(" Expansion slope")+ xlab(" Shrinkage slope")
```

## Step 6: Export

Now that data is cleaned, we can export for analysis with weather data.

```{r}
#for plotting 
write.csv(file = "dend309full_cleaned.csv", dend309_L3)

#daily growth stats 
write.csv(file = "dendro_309full_phase_stats.csv", dend309_phase)
```

# 
